### 2024-11-22 submission.Rmd
submission

---
title: "Using HMMs for pre-snap coverage detection"
output: html_document
date: "`r Sys.Date()`"
---

```{r setup, echo = FALSE, warning=FALSE, message=FALSE, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

The distinction between man- and zone defense is an important strategic decision of a defensive coordinator in each play in American football. While experienced coaches and quarterbacks can often identify these defensive strategies visually, the growing availability of tracking data presents another opportunity to automate this process. This project aims to leverage hidden Markov models (HMMs) to accurately classify defensive strategies --- man or zone coverage --- based on player movement data. By modeling hidden states that represent the offensive player to be guarded, this approach provides a data-driven framework for unraveling the complexity of defensive patterns and enables real-time tactical insights for coaching and analysis.

```{r cars}
summary(cars)
```

<center>
<video width="600" height="400" controls>
  <source src="play_mov.mp4" type="video/mp4">
</video>
</center>

### 2024-11-22 submission.Rmd
html

@@
 ---
+
+#### <span style="color: grey;"> Submission to metric track </span>
 
@@
 
-The distinction between man- and zone defense is an important strategic decision of a defensive coordinator in each play in American football.
-While experienced coaches and quarterbacks can often identify these defensive strategies visually, the growing availability of tracking data presents another opportunity to automate this process.
-This project aims to leverage hidden Markov models (HMMs) to accurately classify defensive strategies --- man or zone coverage --- based on player movement data.
+Choosing between man- and zone coverage is one of the most important strategic decisions of a defensive coordinator prior to each play in American football.
+While experienced coaches and quarterbacks can often identify these defensive strategies visually, the growing availability of tracking data presents another opportunity to infer the underlying tactic.
+This project aims to leverage hidden Markov models (HMMs) to detect defensive strategies --- man or zone coverage --- based on pre-snap player movement data. 
 By modeling hidden states that represent the offensive player to be guarded, this approach provides a data-driven framework for unraveling the complexity of defensive patterns and enables real-time tactical insights for coaching and analysis.
 
-```{r cars}
-summary(cars)
-```
+## Data
+
+This video displays a touchdown from the Kansas City Chiefs against the Arizona Cardinals in Week 1 of the 2022 NFL season.
+We can see that, pre-snap, Mecole Hardman (KC #17) travels from the left side to the right side, followed by the defender Marco Wilson (AZ #20), i.e.\ a clear indication for man-coverage.
 
@@
 <video width="600" height="400" controls>
-  <source src="play_mov.mp4" type="video/mp4">
+  <source src="kc_az_td.mp4" type="video/mp4">
 </video>
 </center>
+
+## Analysis
+
+Our analysis comprises different steps:
+
+#### 1. Pre-Motion Analysis
+We train a model to predict whether the defense plays a man- or zone coverage scheme. The 
+model uses the previously described features, blablabla
+
+#### 2. HMM Analysis
+We model the movements of defensive players during the phase of pre-snap motion within a hidden Markov framework, in which the underlying states represent the offensive players to be guarded (see Franks et al., 2015 for a similar approach in basketball).
+
+#### 3. Post-Motion Analysis
+We re-train the pre-motion model to predict whether the defense plays a man- or zone coverage scheme, however, in this step, we incorporate the results from the HMM analysis.
+
+#### 4. Motion evaluation
+By comparing the predictive performance of our pre-motion model and our post-motion model we can determine the effectiveness of motion to detect the correct defensive scheme.
+
+## Code
+
+All code for data preprocessing, model training, prediction and player evaluation can be found [here](https://github.com/janoleko/BDB-2025/).
+
+## References
+*Franks A, Miller A, Bornn L, Goldsberry K (2015). Characterizing the Spatial Structure of Defensive Skill in Professional Basketball. The Annals of Applied Statistics, 9(1), DOI:10.1214/14-AOAS799
+
+*Zucchini W, MacDonald I, Langrock R (2016). Hidden Markov Models for Time Series - An Introduction Using R. CRC Press

### 2024-11-22 submission.Rmd
text

@@
 We model the movements of defensive players during the phase of pre-snap motion within a hidden Markov framework, in which the underlying states represent the offensive players to be guarded (see Franks et al., 2015 for a similar approach in basketball).
+In contrast to Groom et al., 2024, who enforce a state to proxy zone coverage during corner kicks in soccer, we cannot do that as the classical zones in American football will only be covered post-snap.
+To remidiate this problem, we aim to derive the decision of man- or zone coverage from the number of switches for individual players.
+In particular, a low number of switches when offensive players are in motion indicates man coverage whereas a higher number indicates zone coverage.
 

### 2024-11-22 submission.Rmd
.

--- 2024-11-22 submission.Rmd
@@
 We train a model to predict whether the defense plays a man- or zone coverage scheme. The 
-model uses the previously described features, blablabla
+model uses the previously described features, blablabla and as observations information from those plays without pre-snap motion and information until the first motion for plays with motion.
 
@@
 We model the movements of defensive players during the phase of pre-snap motion within a hidden Markov framework, in which the underlying states represent the offensive players to be guarded (see Franks et al., 2015 for a similar approach in basketball).
+In contrast to Groom et al., 2024, who enforce a state to proxy zone coverage during corner kicks in soccer, we cannot proceed similarly as the classical coverage zones in American football will only be covered post-snap.
+To remediate this, we derive the decision of man- or zone coverage from the number of switches for individual players.
+In particular, a low number of switches when offensive players are in motion indicates man coverage whereas a higher number indicates zone coverage.
 
@@
 
+*Groom S, Morris D, Anderson L, Wang S (2024).
+Modeling Defensive Dynamics in Football: A Hidden Markov Model-Based Approach for Man-Marking and Zonal Defending Corner Analysis.
+The 2nd International Workshop on Intelligent Technologies for Precision Sports Science
+
 *Zucchini W, MacDonald I, Langrock R (2016). Hidden Markov Models for Time Series - An Introduction Using R. CRC Press

### 2024-11-22 submission.Rmd
.

@@
 
-Choosing between man- and zone coverage is one of the most important strategic decisions of a defensive coordinator prior to each play in American football.
+Choosing between man- and zone coverage is one of the most important strategic decisions a defensive coordinator has to take prior to each play in American football.
 While experienced coaches and quarterbacks can often identify these defensive strategies visually, the growing availability of tracking data presents another opportunity to infer the underlying tactic.
@@
 
-#### 1. Pre-Motion Analysis
+#### 1. Pre-motion analysis
 We train a model to predict whether the defense plays a man- or zone coverage scheme. The 
@@
 
-#### 2. HMM Analysis
+#### 2. HMM analysis
 We model the movements of defensive players during the phase of pre-snap motion within a hidden Markov framework, in which the underlying states represent the offensive players to be guarded (see Franks et al., 2015 for a similar approach in basketball).
@@
 
-#### 3. Post-Motion Analysis
+#### 3. Post-motion analysis
 We re-train the pre-motion model to predict whether the defense plays a man- or zone coverage scheme, however, in this step, we incorporate the results from the HMM analysis.

### 2024-11-26 coverage_model_real_data.R
.

@@ -5,7 +5,8 @@
 
 library(dplyr)
 library(tidyverse)
-
+#devtools::install_github("janoleko/LaMa")
+library(LaMa)
 
 ## data looks like
 
@@ -26,8 +27,7 @@ data = cbind(data, X)
 
 ## likelihood function
 
-#devtools::install_github("janoleko/LaMa")
-library(LaMa)
+
            
 nll = function(par){
   getAll(par, dat)
@@ -98,7 +98,8 @@ tracking_data = read.csv("tracking_week_1.csv")
 players = read.csv("players.csv")
 plays = read.csv("plays.csv")
 
-# Spiel 2, playId 224 als Beispiel benutzen (Lavante David geht mit Pollard mit und kreuzt CeeDee Lamb auf dem Weg dorthin)
+# Spiel 2022091110, playId 291 als Beispiel benutzen
+# Marco Wilson geht mit Mecole Hardman die ganze Zeit in Man Coverage mit
 
 tracking = tracking_data %>% 
   filter(gameId == unique(tracking_data$gameId)[2])
@@ -106,15 +107,22 @@ tracking = tracking_data %>%
 tracking = tracking_data %>% filter(gameId == 2022091110)
 
 rm(tracking_data)
-tracking_presnap = tracking %>% 
+
+# Filter out for data before the snap, i.e. the time of the motion
+# Connect with player and players df to receive information
+# Filter out the ball, offensive and defensive line and the QB
+tracking_presnap1 = tracking %>% 
   filter(frameType == "BEFORE_SNAP") %>% 
   left_join(., players %>% dplyr::select(nflId, position), by = "nflId") %>% 
-  left_join(., plays %>% dplyr::select(playId, gameId, possessionTeam, pff_manZone), 
+  left_join(., plays %>% dplyr::select(playId, gameId, absoluteYardlineNumber, possessionTeam, pff_manZone), 
             by = c("playId", "gameId")) %>% 
   mutate(off_def = ifelse(club == possessionTeam, 1, 0)) %>% 
-  filter(!(position %in% c("T", "G", "C", NA, "QB", "NT", "DT", "DE"))) %>% 
   mutate(y = y - 53.3/2) %>% # center y-coordinate
   filter(playId == 291)
+
+# Positional adjustment
+tracking_presnap = tracking_presnap1 %>% 
+  filter(!(position %in% c("T", "G", "C", NA, "QB", "NT", "DT", "DE")))
 
 # Offensivspieler-Daten extrahieren
[…]

### 2024-11-26 plots.R

@@ -11,10 +11,13 @@ plays = read.csv("plays.csv")
 # Tracking DATA Week 1 Dallas vs Tampa Bay
 # Tracking DATA Week 1 KC vs AZ
 
-tracking_kc_az = tracking_data %>% 
-  filter(gameId == 2022091110) %>% 
-  filter(playId == 291) %>% filter(frameId >= frameId[which(event == "line_set")[1]]) %>% 
-  filter(frameType == "BEFORE_SNAP")
+tracking_kc_az = tracking %>%
+  filter(gameId == 2022091110) %>%
+  filter(playId == 291) %>% filter(frameId >= frameId[which(event == "line_set")[1]]) #%>%
+  #filter(frameType == "BEFORE_SNAP") #%>% 
+  # left_join(., players %>% dplyr::select(nflId, position), by = "nflId") %>% 
+  # left_join(., plays %>% dplyr::select(playId, gameId, absoluteYardlineNumber, possessionTeam, pff_manZone), 
+  #           by = c("playId", "gameId"))
 
 # Line of Scrimmage
 los = plays %>% 
@@ -30,6 +33,12 @@ field_width <- 53.3
 team_colors = c("blue4", "black", "red")
 
 # probs are coming from the fitted model, must be taken from there
+new_colnames = c("defender_Id", att_ids)
+probs <- lapply(probs, function(df) {
+  colnames(df) <- new_colnames
+  return(df)
+})
+
 data1 = list()
 defender_data = NULL
 for (i in 1:length(probs)) {
@@ -38,7 +47,6 @@ for (i in 1:length(probs)) {
     cbind(.,probs[[i]])
   defender_data = bind_rows(defender_data, data1[[i]])
 }
-
 
 # Animation: Spielerbewegungen
 for (t in unique(tracking_kc_az$time)) {

### 2024-11-27 coverage_model_real_data_v2.R
New versions, v3 läuft noch nicht durch

@@ -10,6 +10,13 @@ library(LaMa)
 
 # With real data ----------------------------------------------------------
 
+# Version control
+# coverage_model_real_data = baseline model
+# coverage_model_real_data_v2 = 1) position ordered such that we can estimate different
+#                                  transition probs (attention when using split)
+#                               2) preprocessing prepared for a whole match 
+#                                  (not only one play)
+
 setwd("~/Sciebo/BDB 2025")
 tracking_data = read.csv("tracking_week_1.csv")
 players = read.csv("players.csv")
@@ -29,37 +36,42 @@ rm(tracking_data)
 
 # Ideen: 
 
-# Sortieren der Defender nach der y-Achse (damit immer die Wahrscheinlichkeit,
-# dass Spieler, die nah beieinander sind, eher die Angreifer übergeben, höher ist)
+# Es gibt auch das Event man_in_motion, nachdem wir starten können
 
-# Es gibt auch das Event man_in_motion, nach dem
-
-# Filter out for data before the snap, i.e. the time of the motion
-# Connect with player and players df to receive information
-# Filter out the ball, offensive and defensive line and the QB
 tracking_presnap1 = tracking %>% 
   mutate(y = y - 53.3/2) %>% # center y-coordinate
-  #split(.,tracking$displayName) %>% 
-  #map(~mutate(., initial_y = y[which(event == "line_set")[1]])) %>% 
-  #bind_rows() %>% 
-  filter(frameType == "BEFORE_SNAP") %>% 
+  mutate(displayName_fac = factor(displayName, levels = unique(displayName))) %>% 
+  mutate(playId_fac = factor(playId, levels = unique(playId))) %>% 
+  group_by(displayName_fac, playId_fac) %>% #use group split to stay with the order
+  group_split() %>% 
+  map(~mutate(., initial_y = y[which(event == "line_set")[1]])) %>% 
+  bind_rows() %>% 
+  group_by(playId_fac) %>% 
+  group_split() %>% 
+  map(~filter(.,frameId >= frameId[which(event == "line_set")[1]])) %>% 
+  bind_rows() %>% 
   # Take out the time where teams are in the huddle, i.e. before first line up
[…]

### 2024-11-27 coverage_model_real_data_v3.R

@@ -10,6 +10,13 @@ library(LaMa)
 
 # With real data ----------------------------------------------------------
 
+# Version control
+# coverage_model_real_data = baseline model
+# coverage_model_real_data_v2 = 1) position ordered such that we can estimate different
+#                                  transition probs (attention when using split)
+#                               2) preprocessing prepared for a whole match 
+#                                  (not only one play)
+
 setwd("~/Sciebo/BDB 2025")
 tracking_data = read.csv("tracking_week_1.csv")
 players = read.csv("players.csv")
@@ -21,7 +28,7 @@ plays = read.csv("plays.csv")
 tracking = tracking_data %>% 
   filter(gameId == unique(tracking_data$gameId)[1])
 
-tracking = tracking_data %>% filter(gameId == 2022091110) %>% filter(playId == 291) 
+tracking = tracking_data %>% filter(gameId == 2022091110) #%>% filter(playId == 291) 
 
 #write.csv(tracking, file = "tracking_data_week_1_game_kcaz.csv")
 
@@ -29,37 +36,42 @@ rm(tracking_data)
 
 # Ideen: 
 
-# Sortieren der Defender nach der y-Achse (damit immer die Wahrscheinlichkeit,
-# dass Spieler, die nah beieinander sind, eher die Angreifer übergeben, höher ist)
+# Es gibt auch das Event man_in_motion, nachdem wir starten können
 
-# Es gibt auch das Event man_in_motion, nach dem
-
-# Filter out for data before the snap, i.e. the time of the motion
-# Connect with player and players df to receive information
-# Filter out the ball, offensive and defensive line and the QB
 tracking_presnap1 = tracking %>% 
   mutate(y = y - 53.3/2) %>% # center y-coordinate
-  #split(.,tracking$displayName) %>% 
-  #map(~mutate(., initial_y = y[which(event == "line_set")[1]])) %>% 
-  #bind_rows() %>% 
-  filter(frameType == "BEFORE_SNAP") %>% 
+  mutate(displayName_fac = factor(displayName, levels = unique(displayName))) %>% 
+  mutate(playId_fac = factor(playId, levels = unique(playId))) %>% 
[…]

### 2024-11-28 submission.Rmd
.

@@
 This project aims to leverage hidden Markov models (HMMs) to detect defensive strategies --- man or zone coverage --- based on pre-snap player movement data. 
-By modeling hidden states that represent the offensive player to be guarded, this approach provides a data-driven framework for unraveling the complexity of defensive patterns and enables real-time tactical insights for coaching and analysis.
+By modeling hidden states that proxy the offensive player to be guarded, this approach provides a data-driven framework for unraveling the complexity of defensive patterns and enables real-time tactical insights for coaching and analysis.
 
-## Data
-
-This video displays a touchdown from the Kansas City Chiefs against the Arizona Cardinals in Week 1 of the 2022 NFL season.
-We can see that, pre-snap, Mecole Hardman (KC #17) travels from the left side to the right side, followed by the defender Marco Wilson (AZ #20), i.e.\ a clear indication for man-coverage.
+The following video displays a touchdown from the Kansas City Chiefs against the Arizona Cardinals in Week 1 of the 2022 NFL season.
+We can see that, pre-snap, Mecole Hardman (KC #17) is in motion.
+He is immediately followed by the defender Marco Wilson (AZ #20), which is a clear indication for man-coverage.
 
@@
 
+## Data
+Analyzing tracking data from nine weeks of the NFL 2022 season, we aim to forecast the defensive scheme (man- or zone defense).
+For this, we use the corresponding data from PFF that analysed every play and assigned the categories \textit{man}, \textit{zone} and \textit{others} representing the different schemes.
+As it is not properly described what \textit{others} means, we omit every play that is associated with this value.
+Then, we end up with XY plays in total, from which the defense played Y in zone and X in man coverage.
+
+Within these plays, we concentrate on the tracking data after the line has been set (because we are not interested in how players come out of the huddle) and before the ball has been snapped by the Center.
+For the HMM analysis, we further concentrate on those plays with pre-snap motion (ZZ plays).
+
+##### Feature engineering
+To accurately forecast the defensive scheme (man- or zone defense) for every play, we need to create various features derived from the tracking data.
+In particular, we conducted the following feature engineering steps:
+
+\textbf{UMSCHREIBEN} \textit{For all players and the ball carrier we use the features already contained in the tracking data, namely x- and y-coordinates, speed, acceleration, distance covered, orientation and direction.
+Except for the ball carrier we further compute the euclidean distance, x-distance and y-distance to the ball carrier.
+For defensive players only, we additionally compute the absolute difference of the defender’s direction and the angle of the shortest segment between the defender and the ball carrier.
+Subsequently, we order all players (in each frame) with respect to their euclidean distance to the ball carrier and standardize all features.
+More details regarding data preprocessing can be found in the Appendix.}
+
 ## Analysis
@@
 We train a model to predict whether the defense plays a man- or zone coverage scheme. The 
-model uses the previously described features, blablabla and as observations information from those plays without pre-snap motion and information until the first motion for plays with motion.
+model uses the previously described features, blablabla.
 
@@
 We model the movements of defensive players during the phase of pre-snap motion within a hidden Markov framework, in which the underlying states represent the offensive players to be guarded (see Franks et al., 2015 for a similar approach in basketball).
-In contrast to Groom et al., 2024, who enforce a state to proxy zone coverage during corner kicks in soccer, we cannot proceed similarly as the classical coverage zones in American football will only be covered post-snap.
+In contrast to Groom et al., 2024, who enforce a state to proxy zone coverage during corner kicks in soccer, we cannot proceed similarly as the classical coverage zones in American football will only be covered by the defenders post-snap.
+
+\textbf{den folgenden Teil könnte man auch in den Appendix packen}
+
+A hidden Markov models consists of an observed time series $\{\boldsymbol{y}_t\}_{t=1}^T$ --- here, the y-coordinates of the defensive players and an unobserved first-order Markov chain $\{ g_t\}_{t=1}^T$, with $g_t \in \{1,\ldots,N\}$ which proxies the offensives players to be guarded at every time point $t$.
+The Markov chain is fully described by an initial distribution $\boldsymbol{\delta}=\bigl( \Pr(g_1=1), \ldots, \Pr(g_1=N) \bigr)$ and a transition probability matrix (t.p.m.) $\boldsymbol{\Gamma} = (\gamma_{ij}),$ with $\gamma_{ij} = \Pr(g_t = j| g_{t-1} = i), \ i,j = 1, \ldots, N$.
+The connection of both stochastic processes arises from the assumption that the distribution of the observations $\boldsymbol{y}_t$ are fully determined by the currently active state, i.e.\
+\begin{equation*}
+f(\boldsymbol{y}_t|g_1, \ldots, g_T, \boldsymbol{y}_1, \ldots, \boldsymbol{y}_{t-1},\boldsymbol{y}_{t+1},\ldots,\boldsymbol{y}_T) = f(\boldsymbol{y}_t|g_t).
+\end{equation*}
+In general, $f$ can be any density or probability mass function depending on the type of data.
+We are following the approaches of Franks et al.
+(2015) and Groom (2024) and opt for the Normal distribution.
+
 To remediate this, we derive the decision of man- or zone coverage from the number of switches for individual players.
@@
 By comparing the predictive performance of our pre-motion model and our post-motion model we can determine the effectiveness of motion to detect the correct defensive scheme.
+
+## Results
+
+Hier die Animation rein mit den Verbindungen von den decodierten States
+
+## Discussion
 
@@
 *Zucchini W, MacDonald I, Langrock R (2016). Hidden Markov Models for Time Series - An Introduction Using R. CRC Press
+
+## Appendix

### 2024-12-09 submission.Rmd
New Delta

@@
 
-Choosing between man- and zone coverage is one of the most important strategic decisions a defensive coordinator has to take prior to each play in American football.
-While experienced coaches and quarterbacks can often identify these defensive strategies visually, the growing availability of tracking data presents another opportunity to infer the underlying tactic.
+Choosing between man- and zone coverage is one of the most important strategic decisions a defensive coordinator has to take before each offensive play in American football.
+While experienced coaches and quarterbacks can often identify these defensive strategies visually, the growing availability of tracking data presents another opportunity to infer the tactics by the defence.
 This project aims to leverage hidden Markov models (HMMs) to detect defensive strategies --- man or zone coverage --- based on pre-snap player movement data. 
@@
 Subsequently, we order all players (in each frame) with respect to their euclidean distance to the ball carrier and standardize all features.
-More details regarding data preprocessing can be found in the Appendix.}
+More details regarding data pre-processing can be found in the Appendix.}
 
@@
 #### 1. Pre-motion analysis
-We train a model to predict whether the defense plays a man- or zone coverage scheme. The 
-model uses the previously described features, blablabla.
+We train a model to predict whether the defense plays a man- or zone coverage scheme. In particular, .....
+
+The model uses the previously described features, blablabla.
 
 #### 2. HMM analysis
-We model the movements of defensive players during the phase of pre-snap motion within a hidden Markov framework, in which the underlying states represent the offensive players to be guarded (see Franks et al., 2015 for a similar approach in basketball).
-In contrast to Groom et al., 2024, who enforce a state to proxy zone coverage during corner kicks in soccer, we cannot proceed similarly as the classical coverage zones in American football will only be covered by the defenders post-snap.
+We model the movements of defensive players during the phase of pre-snap motion within a hidden Markov framework, in which the underlying states represent the offensive players to be guarded (see Franks et al.
+2015 for a similar approach in basketball).
+In contrast to Groom et al.
+(2024), who enforce a state to proxy zone coverage during corner kicks in soccer, we cannot proceed similarly as the classical coverage zones in American football will only be covered by the defenders post-snap.
 
@@
 \end{equation*}
-In general, $f$ can be any density or probability mass function depending on the type of data.
-We are following the approaches of Franks et al.
-(2015) and Groom (2024) and opt for the Normal distribution.
+In general, $f$ can be any density or probability mass function depending on the type of data. Following the approaches of Franks et al. (2015) and Groom (2024), we opt for a Gaussian distribution.
 
@@
 #### 3. Post-motion analysis
-We re-train the pre-motion model to predict whether the defense plays a man- or zone coverage scheme, however, in this step, we incorporate the results from the HMM analysis.
+We re-train the pre-motion model to predict whether the defense plays a man- or zone coverage scheme, however, in this step, we incorporate results from the HMM analysis as further covariates.
 
@@
 By comparing the predictive performance of our pre-motion model and our post-motion model we can determine the effectiveness of motion to detect the correct defensive scheme.
+Moreover, we can assess which teams predominantly apply pre-snap motions to increase the likelihood of correctly identifying the applied defensive strategy.
 
@@
 
-All code for data preprocessing, model training, prediction and player evaluation can be found [here](https://github.com/janoleko/BDB-2025/).
+All code for data pre-processing, model training, prediction and player evaluation can be found [here](https://github.com/janoleko/BDB-2025/).
 

### 2024-12-09 coverage_model_real_data.R

@@ -10,7 +10,7 @@ library(LaMa)
 
 # With real data ----------------------------------------------------------
 
-setwd("~/Sciebo/BDB 2025")
+setwd("C:/Users/michels/sciebo/BDB 2025")
 tracking_data = read.csv("tracking_week_1.csv")
 players = read.csv("players.csv")
 plays = read.csv("plays.csv")
@@ -103,6 +103,45 @@ Delta = t(
          })
 )
 
+rnames_old = rownames(Delta)
+# only defenders that are close to offensive players
+Delta2 = t(
+  sapply(split(data, data$nflId),
+         function(x){
+           delta = numeric(n_att)
+           delta[which.min(abs(x$y[1] - x[1, which(str_detect(names(x),"player"))[1]-1 + 1:n_att]))] = min(abs(x$y[1] - x[1, which(str_detect(names(x),"player"))[1]-1 + 1:n_att]))
+           delta
+         })
+)
+
+# Funktion zum Anwenden der Regel pro Zeile
+process_row <- function(row) {
+  non_zero_indices <- which(row > 0)
+  if (length(non_zero_indices) == 1) {
+    result <- rep(0, length(row))
+    result[non_zero_indices] <- row[non_zero_indices]
+  } else if (length(non_zero_indices) > 1) {
+    result <- rep(0, length(row))
+    smallest_index <- non_zero_indices[which.min(row[non_zero_indices])]
+    result[smallest_index] <- row[smallest_index]
+  } else {
+    result <- row
+  }
+  return(result)
+}
+
+# Anwenden auf den gesamten Data Frame
+processed_df <- as.data.frame((apply(Delta2, 2, process_row)))
+
+# Spaltennamen beibehalten
+colnames(processed_df) <- colnames(Delta2)
+row.names(processed_df) = rnames_old
+
+Delta = processed_df[which(rowSums(processed_df)>0),]
+rnames_new = rownames(Delta)
+
+data = data %>% filter(nflId %in% rnames_new)
+
 dat = list(y_pos = data$y,
            X = as.matrix(data[, which(str_detect(names(data),"player"))[1]-1 + 1:n_att]),
            ID = data$nflId,
@@ -156,7 +195,7 @@ probs = cbind(ID = trackID, probs)
 probs = split(as.data.frame(probs), probs[,1])
 probs = lapply(probs, as.matrix)
 
-def = 8
+def = 5
 probs[[def]][1,1]
 start = 1
 plot(probs[[def]][start,-1], type = "h", ylim = c(0,1))

### 2024-12-09 submission.Rmd
Amazon Prime Modell

@@
 This project aims to leverage hidden Markov models (HMMs) to detect defensive strategies --- man or zone coverage --- based on pre-snap player movement data. 
-By modeling hidden states that proxy the offensive player to be guarded, this approach provides a data-driven framework for unraveling the complexity of defensive patterns and enables real-time tactical insights for coaching and analysis.
-
-The following video displays a touchdown from the Kansas City Chiefs against the Arizona Cardinals in Week 1 of the 2022 NFL season.
-We can see that, pre-snap, Mecole Hardman (KC #17) is in motion.
-He is immediately followed by the defender Marco Wilson (AZ #20), which is a clear indication for man-coverage.
+By modeling hidden states that represent the offensive player being guarded, this approach builds on previous attempts to predict zone- or man coverage (see screenshot below from the Pittsburgh Steelers vs.
+Cleveland Browns match), which focused primarily on specific plays without motion.
+In contrast, we are now able to include player movement and exploit the additional information available.
+In this way, we provide a data-driven framework for unravelling the complexity of defensive patterns, enabling real-time tactical insights for coaching and analysis.
 
 <center>
-<video width="600" height="400" controls>
-  <source src="kc_az_td.mp4" type="video/mp4">
-</video>
+  <img src="prime_coverage_prediction.jpg" width="600" height="400" alt="Coverage Prediction">
 </center>
@@
 (2024), who enforce a state to proxy zone coverage during corner kicks in soccer, we cannot proceed similarly as the classical coverage zones in American football will only be covered by the defenders post-snap.
+
+The following video displays a touchdown from the Kansas City Chiefs against the Arizona Cardinals in Week 1 of the 2022 NFL season.
+We can see that, pre-snap, Mecole Hardman (KC #17) is in motion.
+He is immediately followed by the defender Marco Wilson (AZ #20), which is a clear indication for man-coverage.
+
+<center>
+<video width="600" height="400" controls>
+  <source src="kc_az_td.mp4" type="video/mp4">
+</video>
+</center>
 

### 2024-12-09 submission.Rmd
disc

--- 2024-12-09 submission.Rmd
@@
 Choosing between man- and zone coverage is one of the most important strategic decisions a defensive coordinator has to take before each offensive play in American football.
-While experienced coaches and quarterbacks can often identify these defensive strategies visually, the growing availability of tracking data presents another opportunity to infer the tactics by the defence.
-This project aims to leverage hidden Markov models (HMMs) to detect defensive strategies --- man or zone coverage --- based on pre-snap player movement data. 
-By modeling hidden states that proxy the offensive player to be guarded, this approach provides a data-driven framework for unraveling the complexity of defensive patterns and enables real-time tactical insights for coaching and analysis.
-
-The following video displays a touchdown from the Kansas City Chiefs against the Arizona Cardinals in Week 1 of the 2022 NFL season.
-We can see that, pre-snap, Mecole Hardman (KC #17) is in motion.
-He is immediately followed by the defender Marco Wilson (AZ #20), which is a clear indication for man-coverage.
+While experienced offensive coordinators and quarterbacks can often identify these defensive strategies visually, the growing availability of tracking data presents another opportunity to infer these defensive tactics.
+Previous approaches attempted to predict zone- or man coverage.
+One example can be seen in the below screenshot: here, display the Amazon Prime broadcast of the week 11 match between the Pittsburgh Steelers and Cleveland Browns of the current season includes the man- or zone prediction of Amazon's NFL Next Gen Stats model.
+Although the full model and its specifies is not public, the live broadcast prediction mostly focus on specific plays without any pre-snap motion.
+In contrast, in this project, we include player movements after the line has been set, i.e.
+pre-snap motion, and exploit this additional information using hidden Markov models (HMMs).
+By modeling hidden states that represent the offensive player being guarded, we can infer the probabilities of state switches (i.e.
+whether defenders trail with offensive players or pass the responsibility on to their teammates).
+Implementing this information into an pre-motion detection model (such as the model by Amazon), we can show that the detection accuracy of the correct defensive strategy --- man or zone coverage --- is heavily increased.
+In this way, we provide a data-driven framework for assessing the efficiency of pre-snap motion, but also unravelling the complexity of defensive patterns, enabling real-time tactical insights for coaches.
 
 <center>
-<video width="600" height="400" controls>
-  <source src="kc_az_td.mp4" type="video/mp4">
-</video>
+  <img src="prime_coverage_prediction.jpg" width="600" height="400" alt="Coverage Prediction">
 </center>
@@
 
+The following video displays a touchdown from the Kansas City Chiefs against the Arizona Cardinals in Week 1 of the 2022 NFL season.
+We can see that, pre-snap, Mecole Hardman (KC #17) is in motion.
+He is immediately followed by the defender Marco Wilson (AZ #20), which is a clear indication for man-coverage.
+
+<center>
+<video width="600" height="400" controls>
+  <source src="kc_az_td.mp4" type="video/mp4">
+</video>
+</center>
+
 \textbf{den folgenden Teil könnte man auch in den Appendix packen}
@@
 
+A drawback in our approach is that we have seen that the prediction accuracy of the pre-motion model is not perfect, among other things, due to insufficient hyperparameter tuning.
+However, this project focus' was on the pre-snap motion and the appropriate translation of this information into a statistical framework, here, the hidden Markov model framework.
+Thus, our pre-motion model can easily be replaced by another model (such as by the NFL Next Gen Stats model) which in turn can be interconnected with our modeling extension, and, thus, exploiting the information of player movements.
+
 ## Code

### 2024-12-09 submission.Rmd
.

@@
 Previous approaches attempted to predict zone- or man coverage.
-One example can be seen in the below screenshot: here, display the Amazon Prime broadcast of the week 11 match between the Pittsburgh Steelers and Cleveland Browns of the current season includes the man- or zone prediction of Amazon's NFL Next Gen Stats model.
-Although the full model and its specifies is not public, the live broadcast prediction mostly focus on specific plays without any pre-snap motion.
-In contrast, in this project, we include player movements after the line has been set, i.e.
+One example can be seen in the below screenshot which displays a snapshot from the Amazon Prime broadcast of the week 12 match between the Pittsburgh Steelers and Cleveland Browns of the current NFL season, including a prediction by Amazon's NFL Next Gen Stats model whether man- or zone coverage will be played.
+Although the full model and its details are not public, the live broadcast prediction mostly focuses on specific plays without any pre-snap motion.
+In contrast, in this project, we include player movements before the snap, i.e.
 pre-snap motion, and exploit this additional information using hidden Markov models (HMMs).
-By modeling hidden states that represent the offensive player being guarded, we can infer the probabilities of state switches (i.e.
-whether defenders trail with offensive players or pass the responsibility on to their teammates).
-Implementing this information into an pre-motion detection model (such as the model by Amazon), we can show that the detection accuracy of the correct defensive strategy --- man or zone coverage --- is heavily increased.
-In this way, we provide a data-driven framework for assessing the efficiency of pre-snap motion, but also unravelling the complexity of defensive patterns, enabling real-time tactical insights for coaches.
+By modeling hidden states that represent the offensive player being guarded, we infer the probabilities of state persistence, i.e.
+a defensive player is trailing with the offensive opponent, and, thus, also state switches, i.e.
+an offensive players' coverage is passed on to another teammate.
+Implementing this information into an pre-motion detection model (similar to the one from Amazon), we can show that the detection accuracy of the correct defensive strategy --- man- or zone coverage --- is heavily increased.
+In this way, we provide a data-driven framework for assessing the efficiency of pre-snap motion to unveal the defensive tactic, but also unravel the complexity of defensive patterns, enabling real-time tactical insights for coaches.
 
@@
 ## Data
-Analyzing tracking data from nine weeks of the NFL 2022 season, we aim to forecast the defensive scheme (man- or zone defense).
-For this, we use the corresponding data from PFF that analysed every play and assigned the categories \textit{man}, \textit{zone} and \textit{others} representing the different schemes.
+We analyze tracking data from nine weeks of the NFL 2022 season, provided by the NFL Big Data Bowl 2024.
+Beside the tracking data, we also use the delivered play and player information and the corresponding data from PFF that analysed every play and assigned the categories \textit{man}, \textit{zone} and \textit{others} representing the different schemes.
 As it is not properly described what \textit{others} means, we omit every play that is associated with this value.
-Then, we end up with XY plays in total, from which the defense played Y in zone and X in man coverage.
+Then, we end up with XY offensive plays in total, from which the defense played Y in zone and X in man coverage.
 
@@
 
-#### 1. Pre-motion analysis
+#### 1. Pre-motion prediction
 We train a model to predict whether the defense plays a man- or zone coverage scheme. In particular, .....
@@
 
-#### 3. Post-motion analysis
+#### 3. Post-motion prediction
 We re-train the pre-motion model to predict whether the defense plays a man- or zone coverage scheme, however, in this step, we incorporate results from the HMM analysis as further covariates.
+In particular, we include the state-switching probabilities....
 
 #### 4. Motion evaluation
-By comparing the predictive performance of our pre-motion model and our post-motion model we can determine the effectiveness of motion to detect the correct defensive scheme.
-Moreover, we can assess which teams predominantly apply pre-snap motions to increase the likelihood of correctly identifying the applied defensive strategy.
+By comparing the predictive performance of our model without motion and our post-motion model we can determine the effectiveness of player movements before the snap to detect the correct defensive scheme.
+Moreover, we assess which teams predominantly apply pre-snap motions to increase the likelihood of correctly identifying the applied defensive strategy.
 

### 2024-12-09 plots.R
.

@@ -95,3 +95,64 @@ for (t in unique(tracking_kc_az$time)) {
   Sys.sleep(0.1) # Wartezeit in Sekunden
 }
 
+
+# Feste Farbskala (von hellgrau zu schwarz)
+color_palette <- colorRampPalette(c("lightgray", "black"))
+
+# Anzahl an Farbstufen definieren (z. B. 100)
+n_colors <- 100
+colors <- color_palette(n_colors)# Animation: Spielerbewegungen
+for (t in unique(tracking_kc_az$time)) {
+  # Daten für den aktuellen Zeitpunkt filtern
+  current_data <- tracking_kc_az[tracking_kc_az$time == t, ]
+  
+  # Spielfeld zeichnen
+  plot(
+    NA, 
+    xlim = c(min(tracking_kc_az$x)-1, max(tracking_kc_az$x)+1),#c(0, field_length), 
+    ylim = c(min(tracking_kc_az$y)-1, max(tracking_kc_az$y)+1),#c(0, field_width),
+    xlab = "Spielfeld-Länge", ylab = "Spielfeld-Breite",
+    main = paste("Spielerbewegungen - Zeit:", t), asp = 1
+  )
+  grid() # Raster für das Spielfeld
+  
+  # Spielerpositionen hinzufügen
+  points(current_data$x, current_data$y, pch = 19, 
+         col = team_colors[as.factor(current_data$club)], cex = 2)
+  abline(v = los)
+  #text(current_data$x, current_data$y, labels = current_data$nflId, pos = 3, col = "black")
+  
+  # Verbindungslinien basierend auf Wahrscheinlichkeiten hinzufügen
+  defender_probs <- defender_data[defender_data$time == t, ]#[[t %% length(defender_data) + 1]] # Beispiel: Wahrscheinlichkeiten für diesen Zeitpunkt
+  for (i in 1:nrow(defender_probs)) {
+    defender <- defender_probs$defender_Id[i]
+    # Maximale Wahrscheinlichkeit für diesen Verteidiger
+    #max_prob <- which.max(defender_probs[i, 20:24])
+    for (j in 1:n_att) {
+      attacker <- as.numeric(names(defender_probs)[j + 19])
+      current_data2 = current_data %>% filter(!is.na(nflId))
+      # Koordinaten von Verteidiger und Angreifer finden
+      defender_pos <- current_data[which(current_data2$nflId == defender), c("x", "y")]
[…]

### 2024-12-11 coverage_model_real_data_v3.R
.

@@ -44,13 +44,13 @@ tracking_presnap1 = tracking %>%
   mutate(playId_fac = factor(playId, levels = unique(playId))) %>% 
   group_by(displayName_fac, playId_fac) %>% #use group split to stay with the order
   group_split() %>% 
+  # Take out the time where teams are in the huddle, i.e. before first line up
   map(~mutate(., initial_y = y[which(event == "line_set")[1]])) %>% 
   bind_rows() %>% 
   group_by(playId_fac) %>% 
   group_split() %>% 
   map(~filter(.,frameId >= frameId[which(event == "line_set")[1]])) %>% 
   bind_rows() %>% 
-  # Take out the time where teams are in the huddle, i.e. before first line up
   filter(frameType == "BEFORE_SNAP") %>% # Filter out for data before the snap, i.e. the time of the motion
   # Connect with player and players df to receive information
   left_join(., players %>% dplyr::select(nflId, position), by = "nflId") %>% 
@@ -106,6 +106,18 @@ data <- def_data %>%
 ## deterministic initial distribution
 n_att = 5
 
+Delta_fun = function(data){
+  
+  t(
+  sapply((split(data, data$nflId))[as.character(unique(data$nflId))],
+         function(x){
+           delta = numeric(n_att)
+           delta[which.min(abs(x$y[1] - x[1, which(str_detect(names(x),"player"))[1]-1 + 1:n_att]))] = 1
+           delta
+         })
+)
+}
+
 #new
 playId = data$playId
 playerId = data$nflId
@@ -115,7 +127,7 @@ dat = list(playId = data %>% group_by(playId) %>% group_split() %>% map(~pull(.,
            y_pos = data %>% group_by(playId) %>% group_split() %>% map(~pull(.,y)),
            X = data.frame(playId = data$playId, 
                             X = as.matrix(data[, which(str_detect(names(data),"player"))[1]-1 + 1:n_att])) %>% 
-             group_by(playId) %>% group_split() %>% map(~select(.,-c(playId))),
+             group_by(playId) %>% group_split() %>% map(~select(.,-c(playId))) %>% lapply(as.matrix),
            n_att = n_att,
            Deltas = lapply(data %>% group_by(playId) %>% group_split(), function(x) Delta_fun(x))
 )
@@ -136,7 +148,8 @@ nll = function(par){
   # alpha * X + (1-alpha) * y_middle
[…]

### 2024-12-11 plots.R

@@ -97,7 +97,7 @@ for (t in unique(tracking_kc_az$time)) {
 
 
 # Feste Farbskala (von hellgrau zu schwarz)
-color_palette <- colorRampPalette(c("lightgray", "black"))
+color_palette <- colorRampPalette(c("white", "black"))
 
 # Anzahl an Farbstufen definieren (z. B. 100)
 n_colors <- 100
@@ -116,10 +116,6 @@ for (t in unique(tracking_kc_az$time)) {
   )
   grid() # Raster für das Spielfeld
   
-  # Spielerpositionen hinzufügen
-  points(current_data$x, current_data$y, pch = 19, 
-         col = team_colors[as.factor(current_data$club)], cex = 2)
-  abline(v = los)
   #text(current_data$x, current_data$y, labels = current_data$nflId, pos = 3, col = "black")
   
   # Verbindungslinien basierend auf Wahrscheinlichkeiten hinzufügen
@@ -148,6 +144,11 @@ for (t in unique(tracking_kc_az$time)) {
         )
       #}      
     }
+    
+    # Spielerpositionen hinzufügen
+    points(current_data$x, current_data$y, pch = 19, 
+           col = team_colors[as.factor(current_data$club)], cex = 2)
+    abline(v = los)
 
   }
   

### 2024-12-11 tracks.R

@@ -1,0 +1,341 @@
+
+#### BDB 2025 - Coverage model ####
+
+# Packages ----------------------------------------------------------------
+
+library(dplyr)
+library(tidyverse)
+#devtools::install_github("janoleko/LaMa")
+library(LaMa)
+
+# With real data ----------------------------------------------------------
+
+# Version control
+# coverage_model_real_data = baseline model
+# coverage_model_real_data_v2 = 1) position ordered such that we can estimate different
+#                                  transition probs (attention when using split)
+#                               2) preprocessing prepared for a whole match 
+#                                  (not only one play)
+
+setwd("C:/Users/michels/sciebo/BDB 2025")
+tracking_data = read.csv("tracking_week_1.csv")
+players = read.csv("players.csv")
+plays = read.csv("plays.csv")
+
+# Spiel 2022091110, playId 291 als Beispiel benutzen
+# Marco Wilson geht mit Mecole Hardman die ganze Zeit in Man Coverage mit
+
+tracking = tracking_data %>% 
+  filter(gameId %in% unique(tracking_data$gameId)[1:2]) 
+
+tracking = tracking %>% filter(playId %in% unique(tracking$playId)[1:2])
+#tracking = tracking_data %>% filter(gameId == 2022091110) #%>% filter(playId == 291) 
+
+#write.csv(tracking, file = "tracking_data_week_1_game_kcaz.csv")
+
+#rm(tracking_data)
+
+# Ideen: 
+
+# Es gibt auch das Event man_in_motion, nachdem wir starten können
+
+tracking_presnap1 = tracking %>% 
+  mutate(uniId = paste0(gameId, playId, nflId)) %>% 
+  mutate(uniId2 = paste0(gameId, playId)) %>% 
+  mutate(y = y - 53.3/2) %>% # center y-coordinate
+  mutate(displayName_fac = factor(displayName, levels = unique(displayName))) %>% 
+  mutate(playId_fac = factor(playId, levels = unique(playId))) %>% 
+  group_by(displayName_fac, playId_fac) %>% #use group split to stay with the order
+  group_split() %>% 
+  # Take out the time where teams are in the huddle, i.e. before first line up
[…]

### 2024-12-11 tracks.R
changes added

@@ -25,8 +25,8 @@ plays = read.csv("plays.csv")
 # Spiel 2022091110, playId 291 als Beispiel benutzen
 # Marco Wilson geht mit Mecole Hardman die ganze Zeit in Man Coverage mit
 
-tracking = tracking_data %>% 
-  filter(gameId %in% unique(tracking_data$gameId)[1:2]) 
+tracking = tracking_data #%>% 
+#filter(gameId %in% unique(tracking_data$gameId)[1:2]) 
 
 tracking = tracking %>% filter(playId %in% unique(tracking$playId)[1:2])
 #tracking = tracking_data %>% filter(gameId == 2022091110) #%>% filter(playId == 291) 
@@ -104,7 +104,7 @@ off_players <- tracking_presnap %>%
   mutate(off_num = row_number()) %>% # Nummeriere Offensivspieler innerhalb jeder Sekunde
   # arrange(initial_y) %>% 
   ungroup()
-  
+
 off_data = off_players %>% 
   select(gameId, playId, time, off_num, y) %>% #, x) %>%
   pivot_wider(
@@ -315,7 +315,7 @@ def_ids = def_data %>%
                                                  filter(nflId %in% att_ids) %>% pull(nflId)), ]
 
 def = 999
-def = "202209120044654618"
+def = "20220911136547877"
 start = 1
 plot(as.numeric(probs[[def]][start,-1]), type = "h", ylim = c(0,1))
 for(t in (start + 1) : nrow(probs[[def]])){
@@ -323,7 +323,59 @@ for(t in (start + 1) : nrow(probs[[def]])){
   Sys.sleep(0.1)
 }
 
-#Start Delta Reihenfolge passt noch nicht
+
+analyze_data <- function(df) {
+  # Spalten 2 bis 6 in numerische Werte umwandeln
+  numeric_values <- apply(df[, 2:6], 2, function(col) as.numeric(as.character(col)))
+  
+  # Fehlende Werte (NA) behandeln, falls Umwandlung fehlschlägt
+  if (anyNA(numeric_values)) {
+    warning("Es gibt NA-Werte nach der Umwandlung. Bitte prüfen!")
+  }
+  
+  # Den größten Wert pro Zeile bestimmen
+  max_values <- apply(numeric_values, 1, which.max)
+  
+  # Anzahl der Änderungen des größten Wertes berechnen
+  changes <- sum(c(NA, diff(max_values)) != 0, na.rm = TRUE)
+  
+  # Ergebnis als Liste zurückgeben
[…]

### 2024-12-12 tracks.R
wildcat prohibited

@@ -31,6 +31,11 @@ tracking = tracking_data #%>%
 tracking = tracking %>% filter(playId %in% unique(tracking$playId)[1:2])
 #tracking = tracking_data %>% filter(gameId == 2022091110) #%>% filter(playId == 291) 
 
+tracking_data$uniId2 = paste0(tracking_data$gameId, tracking_data$playId)
+
+unid2_motion = tracking_data %>% filter(event == "man_in_motion") %>% pull(uniId2)
+
+tracking = tracking_data %>% filter(uniId2 %in% unid2_motion)
 #write.csv(tracking, file = "tracking_data_week_1_game_kcaz.csv")
 
 #rm(tracking_data)
@@ -69,12 +74,13 @@ tracking_presnap1 = tracking %>%
 position_counts <- tracking_presnap1 %>%
   group_by(uniId2, position) %>%
   summarise(Count = n(), .groups = "drop") %>%
-  filter(position %in% c("T", "G")) %>%
+  filter(position %in% c("T", "G", "QB", "C")) %>%
   pivot_wider(names_from = position, values_from = Count, values_fill = 0)
 
 # 2. Filtere PlayIds mit gleichen Häufigkeiten von T und G
 valid_play_ids <- position_counts %>%
   filter(T == G) %>%
+  filter(QB == C) %>% 
   pull(uniId2)
 
 valid_play_ids2 = tracking_presnap1 %>% group_by(uniId2) %>% 
@@ -385,6 +391,7 @@ plays %>%
   View()
 select(playId)
 
+save(probs, file = "probs.RData")
 
 # Some Ideas
 

### 2024-12-12 tracks.R
.

@@ -19,16 +19,19 @@ library(LaMa)
 
 setwd("C:/Users/michels/sciebo/BDB 2025")
 tracking_data = read.csv("tracking_week_1.csv")
+tracking_data = bind_rows(tracking_data, read.csv("tracking_week_2.csv"))
+tracking_data = bind_rows(tracking_data, read.csv("tracking_week_3.csv"))
+tracking_data = bind_rows(tracking_data, read.csv("tracking_week_4.csv"))
 players = read.csv("players.csv")
 plays = read.csv("plays.csv")
 
 # Spiel 2022091110, playId 291 als Beispiel benutzen
 # Marco Wilson geht mit Mecole Hardman die ganze Zeit in Man Coverage mit
 
-tracking = tracking_data #%>% 
+#tracking = tracking_data #%>% 
 #filter(gameId %in% unique(tracking_data$gameId)[1:2]) 
 
-tracking = tracking %>% filter(playId %in% unique(tracking$playId)[1:2])
+#tracking = tracking %>% filter(playId %in% unique(tracking$playId)[1:2])
 #tracking = tracking_data %>% filter(gameId == 2022091110) #%>% filter(playId == 291) 
 
 tracking_data$uniId2 = paste0(tracking_data$gameId, tracking_data$playId)
@@ -391,7 +394,7 @@ plays %>%
   View()
 select(playId)
 
-save(probs, file = "probs.RData")
+save(probs, file = "probs_4weeks.RData")
 
 # Some Ideas
 

### 2024-12-17 plots.R
.

@@ -19,6 +19,14 @@ tracking_kc_az = tracking %>%
   # left_join(., plays %>% dplyr::select(playId, gameId, absoluteYardlineNumber, possessionTeam, pff_manZone), 
   #           by = c("playId", "gameId"))
 
+
+#20220911 031700
+#data %>% filter(uniId == "2022091103170046097")
+tracking_kc_az = tracking_data %>% filter(gameId == 2022091103, playId == 1700)%>% filter(frameId >= frameId[which(event == "line_set")[1]])
+los = plays %>% 
+  filter(gameId == 2022091103) %>% 
+  filter(playId == 1700) %>% pull(absoluteYardlineNumber)
+
 # Line of Scrimmage
 los = plays %>% 
   filter(gameId == 2022091110) %>% 

### 2024-12-17 tracks_weighted.R

@@ -1,0 +1,574 @@
+
+#### BDB 2025 - Coverage model ####
+
+# Packages ----------------------------------------------------------------
+
+library(dplyr)
+library(tidyverse)
+#devtools::install_github("janoleko/LaMa")
+library(LaMa)
+
+# With real data ----------------------------------------------------------
+
+# Version control
+# coverage_model_real_data = baseline model
+# coverage_model_real_data_v2 = 1) position ordered such that we can estimate different
+#                                  transition probs (attention when using split)
+#                               2) preprocessing prepared for a whole match 
+#                                  (not only one play)
+
+pre_process <- function(tracking_data){
+  
+  tracking_data$uniId2 = paste0(tracking_data$gameId, tracking_data$playId)
+  
+  unid2_motion = tracking_data %>% filter(event == "man_in_motion") %>% pull(uniId2)
+  
+  tracking = tracking_data %>% filter(uniId2 %in% unid2_motion)
+  
+  tracking_presnap1 = tracking %>% 
+    mutate(uniId = paste0(gameId, playId, nflId)) %>% 
+    mutate(uniId2 = paste0(gameId, playId)) %>% 
+    mutate(y = y - 53.3/2) %>% # center y-coordinate
+    mutate(x = x - 120/2) %>% 
+    mutate(displayName_fac = factor(displayName, levels = unique(displayName))) %>% 
+    mutate(playId_fac = factor(playId, levels = unique(playId))) %>% 
+    group_by(displayName_fac, playId_fac) %>% #use group split to stay with the order
+    group_split() %>% 
+    # Take out the time where teams are in the huddle, i.e. before first line up
+    map(~mutate(., initial_y = y[which(event == "line_set")[1]])) %>% 
+    bind_rows() %>% 
+    group_by(playId_fac) %>% 
+    group_split() %>% 
+    map(~filter(.,frameId >= frameId[which(event == "line_set")[1]])) %>% 
+    bind_rows() %>% 
+    filter(frameType == "BEFORE_SNAP") %>% # Filter out for data before the snap, i.e. the time of the motion
+    # Connect with player and players df to receive information
+    left_join(., players %>% dplyr::select(nflId, position), by = "nflId") %>% 
[…]

### 2024-12-18 tracks_weighted.R
.

@@ -140,6 +140,8 @@ pre_process <- function(tracking_data){
 n_att = 5
 
 setwd("C:/Users/michels/sciebo/BDB 2025")
+players = read.csv("players.csv")
+plays = read.csv("plays.csv")
 tracking_data = pre_process(read.csv("tracking_week_1.csv"))
 tracking_data = bind_rows(tracking_data, pre_process(read.csv("tracking_week_2.csv")))
 tracking_data = bind_rows(tracking_data, pre_process(read.csv("tracking_week_3.csv")))
@@ -151,8 +153,6 @@ tracking_data = bind_rows(tracking_data, pre_process(read.csv("tracking_week_8.c
 tracking_data = bind_rows(tracking_data, pre_process(read.csv("tracking_week_9.csv")))
 
 data = tracking_data
-players = read.csv("players.csv")
-plays = read.csv("plays.csv")
 
 # Spiel 2022091110, playId 291 als Beispiel benutzen
 # Marco Wilson geht mit Mecole Hardman die ganze Zeit in Man Coverage mit
@@ -184,7 +184,7 @@ rm(def_data, off_data, off_players, position_counts, position_counts2, tracking_
 #data = data %>% filter(abs(x - absoluteYardlineNumber) < 10)
 
 # Fit model ---------------------------------------------------------------
-data = data %>% filter(gameId == 2022091110) %>% filter(playId == 291) #filter(uniId2 == "202209110057")# | uniId2 == "202209120064")
+#data = data %>% filter(gameId == 2022091110) %>% filter(playId == 291) #filter(uniId2 == "202209110057")# | uniId2 == "202209120064")
 
 ## deterministic initial distribution
 
@@ -421,7 +421,7 @@ rnames_new = rownames(Delta3)
 
 data = data %>% filter(uniId %in% rnames_new)
 
-write.csv(data, file = "data_4_weeks_for_HMM.csv")
+write.csv(data, file = "data_all_weeks_for_HMM.csv")
 #Deltas = Deltas[which(rownames(Deltas) %in% rnames_new),]
 
 dat = list(y_pos = data$y,
@@ -567,7 +567,7 @@ ggplot(res_df %>% filter(pff_manZone != "Other"), aes(x = average_sd, fill = pff
   scale_y_continuous(labels = scales::percent) + # Prozentformat auf der y-Achse
   theme_minimal()
 
-#save(probs, file = "probs_4weeks.RData")
[…]

### 2024-12-19 tracks_weighted.R
.

@@ -267,8 +267,8 @@ Delta2 <- t(
              x_off <- x[1,i]
              
              # Gewichtete Distanzberechnung
-             wy <- 4  # Gewicht für die y-Distanz
-             wx <- 0.5  # Gewicht für die x-Distanz
+             wy <- 0.95  # Gewicht für die y-Distanz
+             wx <- 0.05  # Gewicht für die x-Distanz
              sqrt(wy * (y_off - y_def)^2 + wx * (x_off - x_def)^2)
            }) %>% unlist()
            
@@ -416,19 +416,20 @@ Delta3 = bind_rows(Delta3)
 
 rnames_new = rownames(Delta3)
 
+Delta4 = as.data.frame(matrix(0.2, nrow = length(unique(data$uniId)), ncol = n_att))
 #Delta4 = Delta4 %>% select(-rnames)
 #Delta4[Delta4 > 0] = 1
 
 data = data %>% filter(uniId %in% rnames_new)
 
-write.csv(data, file = "data_all_weeks_for_HMM.csv")
+#write.csv(data, file = "data_all_weeks_for_HMM.csv")
 #Deltas = Deltas[which(rownames(Deltas) %in% rnames_new),]
 
 dat = list(y_pos = data$y,
            X = as.matrix(data[, which(str_detect(names(data),"player"))[1]-1 + n_att + 1:n_att]),
            ID = data$uniId,
            n_att = n_att, 
-           Delta = as.matrix(Delta3))
+           Delta = as.matrix(Deltas))
 
 par = list(eta = rep(-4, n_att * (n_att - 1)),
            logsigma = log(1),
@@ -502,6 +503,12 @@ for(t in (start + 1) : nrow(probs[[def]])){
   Sys.sleep(0.1)
 }
 
+# Funktion zur Berechnung der Entropie
+calculate_entropy <- function(Z) {
+  Z <- Z[Z > 0]  # Entferne 0-Werte
+  entropy <- -sum(Z * log(Z))  # Berechne Entropie
+  return(entropy)
+}
 
 analyze_data <- function(df) {
   # Spalten 2 bis 6 in numerische Werte umwandeln
@@ -511,6 +518,10 @@ analyze_data <- function(df) {
   if (anyNA(numeric_values)) {
     warning("Es gibt NA-Werte nach der Umwandlung. Bitte prüfen!")
   }
+  #
+  # if(nrow(numeric_values) > 3){
[…]

### 2024-12-20 submission.Rmd
.

@@
 ---
-title: "Using HMMs for pre-snap coverage detection"
+title: "HMMotion: Using tracking data to predict coverage strategies in American Football"
+author: "Rouven Michels, Robert Bajons, Jan-Ole Koslik"
 output: html_document
@@
 
-Choosing between man- and zone coverage is one of the most important strategic decisions a defensive coordinator has to take before each offensive play in American football.
-While experienced offensive coordinators and quarterbacks can often identify these defensive strategies visually, the growing availability of tracking data presents another opportunity to infer these defensive tactics.
+Choosing between man- and zone coverage is one of the most important strategic decisions a defensive coordinator has to make prior to each offensive play in American football.
+While opposing offensive coordinators and quarterbacks try to identify these defensive strategies visually, the rise of tracking data provides another opportunity to infer the hidden defensive tactics.
+
 Previous approaches attempted to predict zone- or man coverage.
@@
 Although the full model and its details are not public, the live broadcast prediction mostly focuses on specific plays without any pre-snap motion.
-In contrast, in this project, we include player movements before the snap, i.e.
-pre-snap motion, and exploit this additional information using hidden Markov models (HMMs).
-By modeling hidden states that represent the offensive player being guarded, we infer the probabilities of state persistence, i.e.
-a defensive player is trailing with the offensive opponent, and, thus, also state switches, i.e.
-an offensive players' coverage is passed on to another teammate.
-Implementing this information into an pre-motion detection model (similar to the one from Amazon), we can show that the detection accuracy of the correct defensive strategy --- man- or zone coverage --- is heavily increased.
-In this way, we provide a data-driven framework for assessing the efficiency of pre-snap motion to unveal the defensive tactic, but also unravel the complexity of defensive patterns, enabling real-time tactical insights for coaches.
+We proceeed similarly, however, in this project, we do not stop there.
+In contrast, we also depict player movements before the snap, i.e.\ pre-snap motion, and exploit this additional information using hidden Markov models (HMMs).
+By modeling the trajectories of defenders depending on hidden states, which represent the offensive players, we infer probabilities for each offensive player being guarded by the specific defenders in every instance of a play that contains player movements.
+We map this high-dimensional time series information to a single number for each play by calculating the entropy, thus, ending up with a measure of the degree of uncertainty.
+Incorporating this information into the existing pre-motion model, we can show that the AUC and the detection accuracy are increased.
+In this way, we provide a data-driven framework that is able to assess the efficiency of pre-snap motion to reveal the defensive strategy, enabling real-time tactical insights for coaches.
 
@@
 We analyze tracking data from nine weeks of the NFL 2022 season, provided by the NFL Big Data Bowl 2024.
-Beside the tracking data, we also use the delivered play and player information and the corresponding data from PFF that analysed every play and assigned the categories \textit{man}, \textit{zone} and \textit{others} representing the different schemes.
+Beside the tracking data, we also use information on plays and players.
+We further considered the corresponding data from PFF that assigned the categories \textit{man}, \textit{zone} and \textit{others} representing the different schemes to each play.
 As it is not properly described what \textit{others} means, we omit every play that is associated with this value.
+Moreover, we omit plays with more than five offensive linemen and with two quarterbacks and those plays that did not contain any pre-snap motion.
 Then, we end up with XY offensive plays in total, from which the defense played Y in zone and X in man coverage.
@@
 Within these plays, we concentrate on the tracking data after the line has been set (because we are not interested in how players come out of the huddle) and before the ball has been snapped by the Center.
-For the HMM analysis, we further concentrate on those plays with pre-snap motion (ZZ plays).
 
@@
 To accurately forecast the defensive scheme (man- or zone defense) for every play, we need to create various features derived from the tracking data.
-In particular, we conducted the following feature engineering steps:
+In particular, we conducted the following feature engineering steps: ROBERT
 
@@
 #### 1. Pre-motion prediction
-We train a model to predict whether the defense plays a man- or zone coverage scheme. In particular, .....
+We train different model (LASSO, Random Forest, XGBoost) to predict whether the defense plays a man- or zone coverage scheme. In particular, .....
 
 The model uses the previously described features, blablabla.
+
+ROBERT
 
@@
 2015 for a similar approach in basketball).
-In contrast to Groom et al.
-(2024), who enforce a state to proxy zone coverage during corner kicks in soccer, we cannot proceed similarly as the classical coverage zones in American football will only be covered by the defenders post-snap.
+
+OLE
 
@@
 \end{equation*}
-In general, $f$ can be any density or probability mass function depending on the type of data. Following the approaches of Franks et al. (2015) and Groom (2024), we opt for a Gaussian distribution.
+In general, $f$ can be any density or probability mass function depending on the type of data. Following the approaches of Franks et al. (2015), we opt for a Gaussian distribution.
 
@@
 In particular, a low number of switches when offensive players are in motion indicates man coverage whereas a higher number indicates zone coverage.
+
+Entropie
 
@@
 
-#### 4. Motion evaluation
+## Results
+
 By comparing the predictive performance of our model without motion and our post-motion model we can determine the effectiveness of player movements before the snap to detect the correct defensive scheme.
@@
 
-## Results
+auc da unbalanced data
+Hier die Animation rein mit den Verbindungen von den decodierten States
 
-Hier die Animation rein mit den Verbindungen von den decodierten States
+robustheit checken mit simplen summarys nach der motion
+
+tests mit conditional independence (vllt in Anhang)
+
+team analysen
 
@@
 A drawback in our approach is that we have seen that the prediction accuracy of the pre-motion model is not perfect, among other things, due to insufficient hyperparameter tuning.
-However, this project focus' was on the pre-snap motion and the appropriate translation of this information into a statistical framework, here, the hidden Markov model framework.
+However, this project's focus was on the pre-snap motion, in particular the appropriate translation of this information into a statistical framework, here, the hidden Markov model.
 Thus, our pre-motion model can easily be replaced by another model (such as by the NFL Next Gen Stats model) which in turn can be interconnected with our modeling extension, and, thus, exploiting the information of player movements.
+
+wir brauchen mehr Daten weil Anzahl an Plays sind nicht besonders hoch
+
+covariates in state process and state dependendt process of the HMM
 
@@
 
-*Groom S, Morris D, Anderson L, Wang S (2024).
+#*Groom S, Morris D, Anderson L, Wang S (2024).
 Modeling Defensive Dynamics in Football: A Hidden Markov Model-Based Approach for Man-Marking and Zonal Defending Corner Analysis.

### 2024-12-20 tracks_weighted.R

@@ -494,6 +494,7 @@ def_ids = def_data %>%
 (players %>% filter(nflId %in% att_ids))[match(att_ids, players %>% 
                                                  filter(nflId %in% att_ids) %>% pull(nflId)), ]
 
+par(mfrow= c(1,1))
 def = 999
 def = "202209111029153565"
 start = 1

### 2024-12-20 submission.Rmd
.

@@
 
-Choosing between man- and zone coverage is one of the most important strategic decisions a defensive coordinator has to make prior to each offensive play in American football.
-While opposing offensive coordinators and quarterbacks try to identify these defensive strategies visually, the rise of tracking data provides another opportunity to infer the hidden defensive tactics.
+Predicting whether a defense is playing man or zone coverage has long been a focal point in football analytics.
+A notable example is Amazon's NFL Next Gen Stats model, which delivers predictions during live broadcasts (see a snapshot of the 2024 Week 12 matchup between the Pittsburgh Steelers and Cleveland Browns.
+These predictions, however, are often limited to analyzing specific plays without incorporating pre-snap motion which is a crucial element of modern offensive strategies.
 
-Previous approaches attempted to predict zone- or man coverage.
-One example can be seen in the below screenshot which displays a snapshot from the Amazon Prime broadcast of the week 12 match between the Pittsburgh Steelers and Cleveland Browns of the current NFL season, including a prediction by Amazon's NFL Next Gen Stats model whether man- or zone coverage will be played.
-Although the full model and its details are not public, the live broadcast prediction mostly focuses on specific plays without any pre-snap motion.
-We proceeed similarly, however, in this project, we do not stop there.
-In contrast, we also depict player movements before the snap, i.e.\ pre-snap motion, and exploit this additional information using hidden Markov models (HMMs).
-By modeling the trajectories of defenders depending on hidden states, which represent the offensive players, we infer probabilities for each offensive player being guarded by the specific defenders in every instance of a play that contains player movements.
-We map this high-dimensional time series information to a single number for each play by calculating the entropy, thus, ending up with a measure of the degree of uncertainty.
-Incorporating this information into the existing pre-motion model, we can show that the AUC and the detection accuracy are increased.
-In this way, we provide a data-driven framework that is able to assess the efficiency of pre-snap motion to reveal the defensive strategy, enabling real-time tactical insights for coaches.
+Our project takes this analysis a step further.
+While we similarly predict man or zone coverage when the teams are set, we further leverage the additional information of pre-snap player movements.
+Using hidden Markov models (HMMs), we model defenders' trajectories based on hidden states, representing the offensive players they may be guarding.
+This probabilistic approach allows us to infer dynamic coverage patterns during plays with pre-snap motion.
+
+To exploit this complex, high-dimensional data into actionable insights, we calculate the entropy of the state probabilities, producing a single measure of uncertainty for each play.
+Incorporating these as covariates into the existing pre-motion model significantly improves both the AUC and detection accuracy.
+In doing so, we present a data-driven framework that evaluates the effectiveness of pre-snap motion in uncovering defensive strategies, providing real-time tactical insights for coaches.
 

### 2024-12-20 submission.Rmd
.

@@
 ## Introduction
-
-Predicting whether a defense is playing man or zone coverage has long been a focal point in football analytics.
+Deciding between man and zone coverage is one of the most critical strategic choices a defensive coordinator must make before each offensive play in American football.
+While experienced offensive coordinators and quarterbacks often rely on visual cues to identify these defensive schemes, the increasing availability of player tracking data offers a new avenue to uncover and analyze these tactics.
 A notable example is Amazon's NFL Next Gen Stats model, which delivers predictions during live broadcasts (see a snapshot of the 2024 Week 12 matchup between the Pittsburgh Steelers and Cleveland Browns.
@@
 
-Our project takes this analysis a step further.
+Our project takes this model a step further.
 While we similarly predict man or zone coverage when the teams are set, we further leverage the additional information of pre-snap player movements.
-Using hidden Markov models (HMMs), we model defenders' trajectories based on hidden states, representing the offensive players they may be guarding.
+Using a hidden Markov model (HMM), we model defenders' trajectories based on hidden states, which represent the offensive players they may be guarding.
 This probabilistic approach allows us to infer dynamic coverage patterns during plays with pre-snap motion.
+To exploit this high-dimensional time-series data, we calculate the entropy of the state probabilities, thus, producing measures of uncertainty for each play.
+Incorporating these as covariates into the existing pre-motion model significantly improves both the AUC and detection accuracy.
 
-To exploit this complex, high-dimensional data into actionable insights, we calculate the entropy of the state probabilities, producing a single measure of uncertainty for each play.
-Incorporating these as covariates into the existing pre-motion model significantly improves both the AUC and detection accuracy.
 In doing so, we present a data-driven framework that evaluates the effectiveness of pre-snap motion in uncovering defensive strategies, providing real-time tactical insights for coaches.

### 2024-12-29 submission.Rmd
.
sub

@@
 #### <span style="color: grey;"> Submission to metric track </span>
+
+```{r colorize,echo=FALSE}
+#| echo: false
+# colorize text: use inline as `r colorize(text, color)` to print `text` in a given `color`
+# can also be used to color a color name, as in r colorize("red")`
+colorize <- function(text, color) {
+  if (missing(color)) color <- text
+  if (knitr::is_latex_output()) {
+    sprintf("\\textcolor{%s}{%s}", color, text)
+  } else if (knitr::is_html_output()) {
+    sprintf("<span style='color: %s;'>%s</span>", color, text)
+  } else text
+}
+```
 
@@
 These predictions, however, are often limited to analyzing specific plays without incorporating pre-snap motion which is a crucial element of modern offensive strategies.
+`r colorize("Ich weiß nicht, ob ich das so formulieren würde...
+Wir wissen nicht was deren Modell alles inkludiert.
+Es kann gut sein, dass sie auch tracking und daher auch premotion Daten nehmen.
+Oder glaubst Du ist das eine Expertise von einer Person mit domain knowledge und kein statistisches Modell?","red")`
 
@@
 <center>
-  <img src="prime_coverage_prediction.jpg" width="600" height="400" alt="Coverage Prediction">
+  <img src="prime_coverage_prediction.jpg" width="500" height="250" alt="Coverage Prediction">
 </center>
@@
 To accurately forecast the defensive scheme (man- or zone defense) for every play, we need to create various features derived from the tracking data.
-In particular, we conducted the following feature engineering steps: ROBERT
+In particular, we conducted the following feature engineering steps: We first consider all 11 players on each side of the field and compute features related to the convex hull of the positions of the players.
+In particular, for defense and offense, we compute the area spanned by the convex hull of all player such as well as the largest $y$ distance (i.e.
+the width of the hull) and the largest $x$ distance (i.e.
+the length of the hull).
+In addition, we select the 5 most relevant players on each side of the field (for offense: we omit the offensive line and the QB; for defense: `r colorize("TODO, Rouven","red")`) and use their standardized $x$ and $y$ coordinates as covariates.
+(`r colorize("Distances, orientation??
+Appenix??","red")`)
 

### 2024-12-30 submission.Rmd
.

@@
 A notable example is Amazon's NFL Next Gen Stats model, which delivers predictions during live broadcasts (see a snapshot of the 2024 Week 12 matchup between the Pittsburgh Steelers and Cleveland Browns.
-These predictions, however, are often limited to analyzing specific plays without incorporating pre-snap motion which is a crucial element of modern offensive strategies.
+These predictions, however, seem to be, from a viewership's assessment, limited to specific plays without incorporating pre-snap motion which is a crucial element of modern offensive strategies.
 `r colorize("Ich weiß nicht, ob ich das so formulieren würde...
@@
 Es kann gut sein, dass sie auch tracking und daher auch premotion Daten nehmen.
-Oder glaubst Du ist das eine Expertise von einer Person mit domain knowledge und kein statistisches Modell?","red")`
+Oder glaubst Du ist das eine Expertise von einer Person mit domain knowledge und kein statistisches Modell?
+Rouven: Ne glaube das ist ein stat.
+Modell aber habe mir ein paar Szenen angeschaut und da kam die Prediction eigentlich immer nur wenn sich keiner bewegte.","red")`
 
@@
 In particular, we conducted the following feature engineering steps: We first consider all 11 players on each side of the field and compute features related to the convex hull of the positions of the players.
-In particular, for defense and offense, we compute the area spanned by the convex hull of all player such as well as the largest $y$ distance (i.e.
-the width of the hull) and the largest $x$ distance (i.e.
+In particular, for defense and offense, we compute the area spanned by the convex hull of all player such as well as the largest $y$ distance (i.e.\ the width of the hull) and the largest $x$ distance (i.e.
 the length of the hull).
-In addition, we select the 5 most relevant players on each side of the field (for offense: we omit the offensive line and the QB; for defense: `r colorize("TODO, Rouven","red")`) and use their standardized $x$ and $y$ coordinates as covariates.
+In addition, we select the five most relevant players on each side of the field.
+For offense, we omit the offensive line and the QB, while, for defense, we omit nose tackles, defensive tackles and defensive ends, and select the five defenders that were the closest to the five attackers corresponding to a weighted euclidean distance, putting much more emphasis on the y-axis.
+Finally, we use their standardized $x$ and $y$ coordinates as covariates and order defensive and offensive players according to their $y$ coordinates..
 (`r colorize("Distances, orientation??
-Appenix??","red")`)
+Appenix??
+Rouven: würde sagen ja","red")`)
 
@@
 #### 1. Pre-motion prediction
-We train different model (LASSO, Random Forest, XGBoost) to predict whether the defense plays a man- or zone coverage scheme. In particular, .....
+We train different models (LASSO, Random Forest, XGBoost) to predict whether the defense plays a man- or zone coverage scheme. In particular, .....
 
@@
 
-#### 2. HMM analysis
-We model the movements of defensive players during the phase of pre-snap motion within a hidden Markov framework, in which the underlying states represent the offensive players to be guarded (see Franks et al.
+#### 2. Hidden Markov model
+We describe the movements of the five defensive players during the phase of pre-snap motion within a hidden Markov model (see the Appendix for a thorough description).
+In particular, we assume the defender's $y$-coordinates in each time step $t$ to be realizations of Gaussian distributions with means according to the $y$-coordinates of the offensive players to be guarded (these are the underlying states) and a standard deviation to be estimated (see Franks et al.
 2015 for a similar approach in basketball).
@@
 
-The following video displays a touchdown from the Kansas City Chiefs against the Arizona Cardinals in Week 1 of the 2022 NFL season.
+The results of the HMM are exemplified using the following video.
+It displays a touchdown from the Kansas City Chiefs against the Arizona Cardinals in Week 1 of the 2022 NFL season.
 We can see that, pre-snap, Mecole Hardman (KC #17) is in motion.
@@
 
-\textbf{den folgenden Teil könnte man auch in den Appendix packen}
+Using t
 
-A hidden Markov models consists of an observed time series $\{\boldsymbol{y}_t\}_{t=1}^T$ --- here, the y-coordinates of the defensive players and an unobserved first-order Markov chain $\{ g_t\}_{t=1}^T$, with $g_t \in \{1,\ldots,N\}$ which proxies the offensives players to be guarded at every time point $t$.
-The Markov chain is fully described by an initial distribution $\boldsymbol{\delta}=\bigl( \Pr(g_1=1), \ldots, \Pr(g_1=N) \bigr)$ and a transition probability matrix (t.p.m.) $\boldsymbol{\Gamma} = (\gamma_{ij}),$ with $\gamma_{ij} = \Pr(g_t = j| g_{t-1} = i), \ i,j = 1, \ldots, N$.
-The connection of both stochastic processes arises from the assumption that the distribution of the observations $\boldsymbol{y}_t$ are fully determined by the currently active state, i.e.\
-\begin{equation*}
-f(\boldsymbol{y}_t|g_1, \ldots, g_T, \boldsymbol{y}_1, \ldots, \boldsymbol{y}_{t-1},\boldsymbol{y}_{t+1},\ldots,\boldsymbol{y}_T) = f(\boldsymbol{y}_t|g_t).
-\end{equation*}
-In general, $f$ can be any density or probability mass function depending on the type of data. Following the approaches of Franks et al. (2015), we opt for a Gaussian distribution.
+#### 3. Post-motion prediction
+We re-train the pre-motion model to predict whether the defense plays a man- or zone coverage scheme, however, in this step, we incorporate results from the HMM analysis as further covariates.
+In particular, we include the state-switching probabilities....
 
@@
 Entropie
-
-#### 3. Post-motion prediction
-We re-train the pre-motion model to predict whether the defense plays a man- or zone coverage scheme, however, in this step, we incorporate results from the HMM analysis as further covariates.
-In particular, we include the state-switching probabilities....
 
@@
 
-A drawback in our approach is that we have seen that the prediction accuracy of the pre-motion model is not perfect, among other things, due to insufficient hyperparameter tuning.
-However, this project's focus was on the pre-snap motion, in particular the appropriate translation of this information into a statistical framework, here, the hidden Markov model.
-Thus, our pre-motion model can easily be replaced by another model (such as by the NFL Next Gen Stats model) which in turn can be interconnected with our modeling extension, and, thus, exploiting the information of player movements.
-
-wir brauchen mehr Daten weil Anzahl an Plays sind nicht besonders hoch
-
-covariates in state process and state dependendt process of the HMM
+One limitation of our approach lies in the imperfect prediction accuracy of the pre-motion model, primarily due to insufficient hyperparameter tuning and the relatively small number of plays involving motion.
+However, the primary focus of this project was on pre-snap motion, particularly on how to effectively translate this information into the hidden Markov model.
+Importantly, our pre-motion model is modular and can be seamlessly replaced by another model, such as the NFL Next Gen Stats model, which can then be integrated into our modeling extension to fully leverage the insights provided by pre-snap motion.
 
@@
 
-#*Groom S, Morris D, Anderson L, Wang S (2024).
-Modeling Defensive Dynamics in Football: A Hidden Markov Model-Based Approach for Man-Marking and Zonal Defending Corner Analysis.
-The 2nd International Workshop on Intelligent Technologies for Precision Sports Science
-
 *Zucchini W, MacDonald I, Langrock R (2016). Hidden Markov Models for Time Series - An Introduction Using R. CRC Press
@@
 ## Appendix
+
+#### Hidden Markov Model
+A hidden Markov models consists of an observed time series $\{\boldsymbol{y}_t\}_{t=1}^T$ --- here, the y-coordinates of the defensive players and an unobserved first-order Markov chain $\{ g_t\}_{t=1}^T$, with $g_t \in \{1,\ldots,N\}$ which proxies the offensives players to be guarded at every time point $t$.
+The Markov chain is fully described by an initial distribution $\boldsymbol{\delta}=\bigl( \Pr(g_1=1), \ldots, \Pr(g_1=N) \bigr)$ and a transition probability matrix (t.p.m.) $\boldsymbol{\Gamma} = (\gamma_{ij}),$ with $\gamma_{ij} = \Pr(g_t = j| g_{t-1} = i), \ i,j = 1, \ldots, N$.
+The connection of both stochastic processes arises from the assumption that the distribution of the observations $\boldsymbol{y}_t$ are fully determined by the currently active state, i.e.\
+\begin{equation*}
+f(\boldsymbol{y}_t|g_1, \ldots, g_T, \boldsymbol{y}_1, \ldots, \boldsymbol{y}_{t-1},\boldsymbol{y}_{t+1},\ldots,\boldsymbol{y}_T) = f(\boldsymbol{y}_t|g_t).
+\end{equation*}
+In general, $f$ can be any density or probability mass function depending on the type of data. Following the approaches of Franks et al. (2015), we opt for a Gaussian distribution.

### 2024-12-30 submission.Rmd
.

@@
 While experienced offensive coordinators and quarterbacks often rely on visual cues to identify these defensive schemes, the increasing availability of player tracking data offers a new avenue to uncover and analyze these tactics.
-A notable example is Amazon's NFL Next Gen Stats model, which delivers predictions during live broadcasts (see a snapshot of the 2024 Week 12 matchup between the Pittsburgh Steelers and Cleveland Browns.
-These predictions, however, seem to be, from a viewership's assessment, limited to specific plays without incorporating pre-snap motion which is a crucial element of modern offensive strategies.
-`r colorize("Ich weiß nicht, ob ich das so formulieren würde...
-Wir wissen nicht was deren Modell alles inkludiert.
-Es kann gut sein, dass sie auch tracking und daher auch premotion Daten nehmen.
-Oder glaubst Du ist das eine Expertise von einer Person mit domain knowledge und kein statistisches Modell?
-Rouven: Ne glaube das ist ein stat.
-Modell aber habe mir ein paar Szenen angeschaut und da kam die Prediction eigentlich immer nur wenn sich keiner bewegte.","red")`
+A notable example is Amazon's NFL Next Gen Stats model, which delivers coverage predictions during live broadcasts (see a snapshot of the 2024 Week 12 matchup between the Pittsburgh Steelers and Cleveland Browns).
+However, these models seem to be trained on plays without pre-snap motion, or at least to the situations before motions (see [Amazon](https://aws.amazon.com/de/blogs/machine-learning/identifying-defense-coverage-schemes-in-nfls-next-gen-stats/)), which is a crucial element of modern offensive strategies.
 
 Our project takes this model a step further.
-While we similarly predict man or zone coverage when the teams are set, we further leverage the additional information of pre-snap player movements.
+While we similarly predict man- or zone coverage when the teams are set, we further leverage the additional information of pre-snap player movements.
 Using a hidden Markov model (HMM), we model defenders' trajectories based on hidden states, which represent the offensive players they may be guarding.
-This probabilistic approach allows us to infer dynamic coverage patterns during plays with pre-snap motion.
-To exploit this high-dimensional time-series data, we calculate the entropy of the state probabilities, thus, producing measures of uncertainty for each play.
-Incorporating these as covariates into the existing pre-motion model significantly improves both the AUC and detection accuracy.
-
-In doing so, we present a data-driven framework that evaluates the effectiveness of pre-snap motion in uncovering defensive strategies, providing real-time tactical insights for coaches.
+Incorporating summary statistics of the probabilistic HMM results as covariates into the existing pre-motion model significantly improves both the AUC and detection accuracy and further allows for evaluating the effectiveness of pre-snap motion in uncovering defensive strategies, providing real-time tactical insights for coaches.
 
@@
 
-Within these plays, we concentrate on the tracking data after the line has been set (because we are not interested in how players come out of the huddle) and before the ball has been snapped by the Center.
-
 ##### Feature engineering
@@
 Rouven: würde sagen ja","red")`)
-
-\textbf{UMSCHREIBEN} \textit{For all players and the ball carrier we use the features already contained in the tracking data, namely x- and y-coordinates, speed, acceleration, distance covered, orientation and direction.
-Except for the ball carrier we further compute the euclidean distance, x-distance and y-distance to the ball carrier.
-For defensive players only, we additionally compute the absolute difference of the defender’s direction and the angle of the shortest segment between the defender and the ball carrier.
-Subsequently, we order all players (in each frame) with respect to their euclidean distance to the ball carrier and standardize all features.
-More details regarding data pre-processing can be found in the Appendix.}
 
@@
 
-Using t
+This probabilistic approach allows us to infer dynamic coverage patterns during plays with pre-snap motion.
+To exploit this high-dimensional time-series data, we calculate the entropy of the state probabilities, thus, producing measures of uncertainty for each play.
 
@@
 We re-train the pre-motion model to predict whether the defense plays a man- or zone coverage scheme, however, in this step, we incorporate results from the HMM analysis as further covariates.
-In particular, we include the state-switching probabilities....
+In particular, we use the aforementioned state probabilities of guarding the different offensive players in each second.
 

### 2024-12-31 submission.Rmd
sub R2

@@
 As it is not properly described what \textit{others} means, we omit every play that is associated with this value.
-Moreover, we omit plays with more than five offensive linemen and with two quarterbacks and those plays that did not contain any pre-snap motion.
-Then, we end up with XY offensive plays in total, from which the defense played Y in zone and X in man coverage.
+Moreover, we omit plays with more than five offensive linemen and with two quarterbacks `r colorize("and those plays that did not contain any pre-snap motion","orange")` (`r colorize("Ich würde deutlicher drauf eingehen, dass wir nur an motion plays interessiert sind!
+Vl ein Punkt machen und schreiben: Since we are specifically interested in analyzing pre-snap player movements, we omit plays that did not contain any pre-snap motion.","red")`).
+Then, we end up with 3985 offensive plays in total, from which the defense played 2973 in zone and 1012 in man coverage.
 
@@
 For offense, we omit the offensive line and the QB, while, for defense, we omit nose tackles, defensive tackles and defensive ends, and select the five defenders that were the closest to the five attackers corresponding to a weighted euclidean distance, putting much more emphasis on the y-axis.
-Finally, we use their standardized $x$ and $y$ coordinates as covariates and order defensive and offensive players according to their $y$ coordinates..
-(`r colorize("Distances, orientation??
-Appenix??
-Rouven: würde sagen ja","red")`)
+Finally, we use their standardized $x$ and $y$ coordinates as covariates and order defensive and offensive players according to their $y$ coordinates.
+Additionally, for each of the relevant defenders, we compute distances to the football and their orientation with respect to the quarterback (`r colorize("values taken at event lineset","orange")`).
+Finally, we extract relevant information from play-by-play data, such as quarter, down, yards to go, home and away score and the game clock.
+
+Detailed information can be found in the Appendix.
 
@@
 #### 1. Pre-motion prediction
-We train different models (LASSO, Random Forest, XGBoost) to predict whether the defense plays a man- or zone coverage scheme. In particular, .....
+We train different models to predict whether the defense plays a man- or zone coverage scheme. In particular, in the first model, we only use features available pre-motion.
 

### 2024-12-31 submission.Rmd
.

@@
 A notable example is Amazon's NFL Next Gen Stats model, which delivers coverage predictions during live broadcasts (see a snapshot of the 2024 Week 12 matchup between the Pittsburgh Steelers and Cleveland Browns).
-However, these models seem to be trained on plays without pre-snap motion, or at least to the situations before motions (see [Amazon](https://aws.amazon.com/de/blogs/machine-learning/identifying-defense-coverage-schemes-in-nfls-next-gen-stats/)), which is a crucial element of modern offensive strategies.
+However, these models seem to be trained on plays without pre-snap motion, or at least to the situations before motions (see [Amazon](https://aws.amazon.com/de/blogs/machine-learning/identifying-defense-coverage-schemes-in-nfls-next-gen-stats/)), while motion is a crucial element of modern offensive strategies.
 
@@
 As it is not properly described what \textit{others} means, we omit every play that is associated with this value.
-Moreover, we omit plays with more than five offensive linemen and with two quarterbacks `r colorize("and those plays that did not contain any pre-snap motion","orange")` (`r colorize("Ich würde deutlicher drauf eingehen, dass wir nur an motion plays interessiert sind!
-Vl ein Punkt machen und schreiben: Since we are specifically interested in analyzing pre-snap player movements, we omit plays that did not contain any pre-snap motion.","red")`).
-Then, we end up with 3985 offensive plays in total, from which the defense played 2973 in zone and 1012 in man coverage.
+Moreover, we omit plays with more than five offensive linemen and with two quarterbacks.
+Since we are specifically interested in analyzing pre-snap player movements, we omit plays that did not contain any pre-snap motion.
+Then, we end up with $3985$ offensive plays in total, from which the defense played $2973$ in zone and $1012$ in man coverage.
 
@@
 Finally, we use their standardized $x$ and $y$ coordinates as covariates and order defensive and offensive players according to their $y$ coordinates.
-Additionally, for each of the relevant defenders, we compute distances to the football and their orientation with respect to the quarterback (`r colorize("values taken at event lineset","orange")`).
+Additionally, for each of the relevant defenders, we compute distances to the football and their orientation with respect to the quarterback (`r colorize("values taken at event lineset.
+Rouven: Die Values haben wir doch rausgenommen aus dem finalen Modell oder?","orange")`).
 Finally, we extract relevant information from play-by-play data, such as quarter, down, yards to go, home and away score and the game clock.
@@
 </center>
+
+We locally decoded this sequence.
 

### 2024-12-31 submission.Rmd
.

@@
 
+This probabilistic approach allows us to infer dynamic coverage patterns during plays with pre-snap motion.
+
 OLE
@@
 
-We locally decoded this sequence.
+#### 3. Post-motion prediction
+We re-train the pre-motion model to predict whether the defense plays a man- or zone coverage scheme, but, in this step, we incorporate results from the HMM analysis as further covariates.
+However, a caveat of the aforemententioned state probabilities is that they constitute as high-dimensional time-series data which makes them hard to use them as covariates.
+An outlet of this problem is to employ suitable summary statistics (`r colorize("Robert hier spezifizieren welche du final genommen hast","orange")`).
+In particular, we calculate for each defender the most likely offensive player in each time point $t$ and count the number of state switches.
+Summing these numbers across defenders and, thus, plays, are the first, relatively simple, summary statistic.
+In addition to that, we calculate the mean entropy during a play:
 
-This probabilistic approach allows us to infer dynamic coverage patterns during plays with pre-snap motion.
+$$H(X) = - \sum_{i=1}^n P(x_i) \log_b P(x_i)$$
+Entropy is a measure of uncertainty or randomness in a probability distribution, here, the probability distribution of the defender for guarding the offensive players.
+Higher entropy indicates greater unpredictability, while lower entropy signifies more predictability.
+
+In particular, we use the aforementioned state probabilities of guarding the different offensive players in each second.
 To exploit this high-dimensional time-series data, we calculate the entropy of the state probabilities, thus, producing measures of uncertainty for each play.
-
-#### 3. Post-motion prediction
-We re-train the pre-motion model to predict whether the defense plays a man- or zone coverage scheme, however, in this step, we incorporate results from the HMM analysis as further covariates.
-In particular, we use the aforementioned state probabilities of guarding the different offensive players in each second.
-
-To remediate this, we derive the decision of man- or zone coverage from the number of switches for individual players.
-In particular, a low number of switches when offensive players are in motion indicates man coverage whereas a higher number indicates zone coverage.
-
-Entropie
 

### 2025-01-01 submission.Rmd
.

@@
 However, a caveat of the aforemententioned state probabilities is that they constitute as high-dimensional time-series data which makes them hard to use them as covariates.
-An outlet of this problem is to employ suitable summary statistics (`r colorize("Robert hier spezifizieren welche du final genommen hast","orange")`).
-In particular, we calculate for each defender the most likely offensive player in each time point $t$ and count the number of state switches.
-Summing these numbers across defenders and, thus, plays, are the first, relatively simple, summary statistic.
-In addition to that, we calculate the mean entropy during a play:
+An outlet of this problem is to employ suitable summary statistics (`r colorize("Robert hier spezifizieren welche du final genommen hast, ich gehe jetzt mal von summe an switches und entropy aus","orange")`).
+In particular, we calculate for each defender the most likely offensive player $k = 1, \ldots, 5$ to be guarded in each time point $t = 1, \ldots, T_n$.
+From this, we count the number of state switches and sum these numbers across defenders as a first, simple, summary statistics and further calculate across defenders $j = 1, \ldots, 5$ the mean entropy in offensive play $n = 1,\ldots, N$, i.e.:
 
-$$H(X) = - \sum_{i=1}^n P(x_i) \log_b P(x_i)$$
-Entropy is a measure of uncertainty or randomness in a probability distribution, here, the probability distribution of the defender for guarding the offensive players.
+$$H(n) = - \frac{1}{5}\sum_{j = 1}^5 \sum_{k=1}^{5} \left( \frac{1}{T_n} \sum_{t=1}^{T_n} \mathbb{1}\left(\arg\max_{i=1,\ldots,5} X_{t,i} = k\right) \cdot \log\left(\frac{1}{T_n} \sum_{t=1}^{T_n} \mathbb{1}\left(\arg\max_{i=1,\ldots,5} X_{t,i} = k\right)\right) \right).$$
+Entropy is a measure of uncertainty or randomness in a probability distribution, here, the empirical distribution of the defenders guarding the respective offensive players.
 Higher entropy indicates greater unpredictability, while lower entropy signifies more predictability.
-
-In particular, we use the aforementioned state probabilities of guarding the different offensive players in each second.
-To exploit this high-dimensional time-series data, we calculate the entropy of the state probabilities, thus, producing measures of uncertainty for each play.
 

### 2025-01-01 submission.Rmd
.

@@
 Our project takes this model a step further.
-While we similarly predict man- or zone coverage when the teams are set, we further leverage the additional information of pre-snap player movements.
+While we similarly predict man- or zone coverage when the teams are set before snaps, we further leverage the additional information of pre-snap player movements.
 Using a hidden Markov model (HMM), we model defenders' trajectories based on hidden states, which represent the offensive players they may be guarding.
-Incorporating summary statistics of the probabilistic HMM results as covariates into the existing pre-motion model significantly improves both the AUC and detection accuracy and further allows for evaluating the effectiveness of pre-snap motion in uncovering defensive strategies, providing real-time tactical insights for coaches.
+Incorporating summary statistics of the probabilistic HMM results as features into the existing pre-motion model significantly improves both the AUC and detection accuracy and further allows for evaluating the effectiveness of pre-snap motion in uncovering defensive strategies, providing real-time tactical insights for coaches.
 
@@
 For offense, we omit the offensive line and the QB, while, for defense, we omit nose tackles, defensive tackles and defensive ends, and select the five defenders that were the closest to the five attackers corresponding to a weighted euclidean distance, putting much more emphasis on the y-axis.
-Finally, we use their standardized $x$ and $y$ coordinates as covariates and order defensive and offensive players according to their $y$ coordinates.
+Finally, we use their standardized $x$ and $y$ coordinates as features and order defensive and offensive players according to their $y$ coordinates.
 Additionally, for each of the relevant defenders, we compute distances to the football and their orientation with respect to the quarterback (`r colorize("values taken at event lineset.
@@
 #### 3. Post-motion prediction
-We re-train the pre-motion model to predict whether the defense plays a man- or zone coverage scheme, but, in this step, we incorporate results from the HMM analysis as further covariates.
-However, a caveat of the aforemententioned state probabilities is that they constitute as high-dimensional time-series data which makes them hard to use them as covariates.
+We re-train the pre-motion model to predict whether the defense plays a man- or zone coverage scheme, but, in this step, we incorporate results from the HMM analysis as further features.
+However, a caveat of the aforemententioned state probabilities is that they constitute as high-dimensional time-series data which makes them hard to use them as features.
 An outlet of this problem is to employ suitable summary statistics (`r colorize("Robert hier spezifizieren welche du final genommen hast, ich gehe jetzt mal von summe an switches und entropy aus","orange")`).
@@
 
-One limitation of our approach lies in the imperfect prediction accuracy of the pre-motion model, primarily due to insufficient hyperparameter tuning and the relatively small number of plays involving motion.
+While we were clearly able to show that motion enables the defensive coverage scheme, a limitation of our approach lies in the imperfect prediction accuracy of the pre-motion model, primarily due to insufficient hyperparameter tuning and the relatively small number of plays involving motion.
 However, the primary focus of this project was on pre-snap motion, particularly on how to effectively translate this information into the hidden Markov model.
-Importantly, our pre-motion model is modular and can be seamlessly replaced by another model, such as the NFL Next Gen Stats model, which can then be integrated into our modeling extension to fully leverage the insights provided by pre-snap motion.
+Importantly, our pre-motion model is modular and can be seamlessly replaced by another model, such as the NFL Next Gen Stats model, which can then be integrated into our modeling extension to fully leverage the insights provided by our extention.
 

### 2025-01-02 HMM_plot.R
.

@@ -1,0 +1,57 @@
+library(dplyr)
+library(stringr)
+tracking_kc_az = read.csv("tracking_kc.csv")
+defender_data = read.csv("kc_defender_data.csv")
+
+# Spielfeld-Abmessungen (angepasst an Football)
+field_length <- 120
+field_width <- 53.3
+
+team_colors = c("blue4", "black", "red")
+los = 19
+# Animation: Spielerbewegungen
+for (t in unique(tracking_kc_az$time)) {
+  # Daten für den aktuellen Zeitpunkt filtern
+  current_data <- tracking_kc_az[tracking_kc_az$time == t, ]
+  
+  # Spielfeld zeichnen
+  plot(
+    NA, 
+    xlim = c(min(tracking_kc_az$x)-1, max(tracking_kc_az$x)+1),#c(0, field_length), 
+    ylim = c(min(tracking_kc_az$y)-1, max(tracking_kc_az$y)+1),#c(0, field_width),
+    xlab = "Spielfeld-Länge", ylab = "Spielfeld-Breite",
+    main = paste("Spielerbewegungen - Zeit:", t), asp = 1
+  )
+  grid() # Raster für das Spielfeld
+  
+  # Spielerpositionen hinzufügen
+  points(current_data$x, current_data$y, pch = 19, 
+         col = team_colors[as.factor(current_data$club)], cex = 2)
+  abline(v = los)
+  #text(current_data$x, current_data$y, labels = current_data$nflId, pos = 3, col = "black")
+  
+  # Verbindungslinien basierend auf Wahrscheinlichkeiten hinzufügen
+  defender_probs <- defender_data[defender_data$time == t, ]#[[t %% length(defender_data) + 1]] # Beispiel: Wahrscheinlichkeiten für diesen Zeitpunkt
+  for (i in 1:nrow(defender_probs)) {
+    defender <- defender_probs$nflId[i]
+    # Maximale Wahrscheinlichkeit für diesen Verteidiger
+    max_prob <- which.max(defender_probs[i, 21:25])
+    attacker <- as.numeric(str_sub(names(defender_probs)[max_prob + 20], 2, 6))
+    
+    current_data2 = current_data %>% filter(!is.na(nflId))
+    # Koordinaten von Verteidiger und Angreifer finden
+    defender_pos <- current_data[which(current_data2$nflId == defender), c("x", "y")]
[…]

### 2025-01-02 submission.Rmd

@@
 Additionally, for each of the relevant defenders, we compute distances to the football and their orientation with respect to the quarterback (`r colorize("values taken at event lineset.
-Rouven: Die Values haben wir doch rausgenommen aus dem finalen Modell oder?","orange")`).
+Rouven: Die Values haben wir doch rausgenommen aus dem finalen Modell oder?
+Robert: ich denke das sollten wir noch besprechen.
+Ich habe ohnehin mehrere Modelle gefitted...","orange")`).
 Finally, we extract relevant information from play-by-play data, such as quarter, down, yards to go, home and away score and the game clock.
@@
 #### 1. Pre-motion prediction
-We train different models to predict whether the defense plays a man- or zone coverage scheme. In particular, in the first model, we only use features available pre-motion.
+We train different models to predict whether the defense plays a man- or zone coverage scheme.
+In particular, we start with fitting a model only using features available pre-motion.
+Since the aim of the project is to show the effectiveness of pre-snap motion, this allows us to establish a baseline model for detecting zone vs man coverage without motion.
 
-The model uses the previously described features, blablabla.
-
-ROBERT
+Since we have a limited dataset available (only 3985 plays), we want to balance the complexity of the model, i.e.
+we control the number of features available.
+Using the previously described features, we obtain 67 variables.
+`r colorize("6 convex hull related features (3 for offense, 3 for defense), 20 player positions features (10 standardized $x$ and $y$ coordinates, 5 for offense, 5 for defense), 30 distance features (10 Euclidian distances, 10 $x$ distances, and 10 $y$ distances to the football), 10 orientation features (5 offense, 5 defense) and 6 play by play features (probably better suited for appendix...)","orange")`.
+Given the small data set this is rather high number of features and therefore we consider two model classes.
+On the one hand, we fit an elastic net model, which performs implicit feature selection and is able to handle multicollinearity.
+On the other hand we use an xgboost model, which is able to capture non linear effects (and interactions) and in general also handles collinearity, but needs careful hyperparameter tuning and generally performs better on bigger data sets.
 
@@
 However, a caveat of the aforemententioned state probabilities is that they constitute as high-dimensional time-series data which makes them hard to use them as features.
-An outlet of this problem is to employ suitable summary statistics (`r colorize("Robert hier spezifizieren welche du final genommen hast, ich gehe jetzt mal von summe an switches und entropy aus","orange")`).
+An outlet of this problem is to employ suitable summary statistics (`r colorize("Robert hier spezifizieren welche du final genommen hast, ich gehe jetzt mal von summe an switches und entropy aus; Robert: Im moment nehmen ich alle 4 (sum, average, nr changes, average ent)...
+Sollten wir nochmal besprechen.","orange")`).
 In particular, we calculate for each defender the most likely offensive player $k = 1, \ldots, 5$ to be guarded in each time point $t = 1, \ldots, T_n$.
@@
 Hier die Animation rein mit den Verbindungen von den decodierten States
+
+![](AUCs.png)
+![](loglosses.png)
+
+LEGENDE INFO:
+
+- Model af: alle features im model (pre motion mit distances und orientation, hmm features (average,sum,nr changes, average ent), naive postmotion, pbp)
+- Model af_ex_hmm: alle features außer hmm features (pre motion mit distances und orientation, naive postmotion, pbp)
+- Model af_npo: alle features außer naive postmotion features
+- Model af_npo_ex_hmm: alle features außer naive postmotion features und hmm features
+- Model xy_f: premotion features nur x,y coords der spieler (keine distances und orientation), sonst alle features
+- Model xy_f_ex_hmm: wie xy_f, nur ohne hmm features
+- Model xy_npo: wie xy_f, nur ohne naive postmotion features
+- Model xy_npo_ex_hmm: wie xy_npo, nur ohne hmm features
 

### 2025-01-03 submission.Rmd
updating HMM parts

@@
 #### 2. Hidden Markov model
-We describe the movements of the five defensive players during the phase of pre-snap motion within a hidden Markov model (see the Appendix for a thorough description).
-In particular, we assume the defender's $y$-coordinates in each time step $t$ to be realizations of Gaussian distributions with means according to the $y$-coordinates of the offensive players to be guarded (these are the underlying states) and a standard deviation to be estimated (see Franks et al.
+
+We describe the movements of the five defensive players during the phase of pre-snap motion by a hidden Markov model (see the Appendix for an in-depth description).
+In particular, we assume that each defender's $y$-coordinate at each time point $t$ is a realization from a particular Gaussian distribution with mean according to the $y$-coordinate of the offensive player that is currently guarded and an estimated standard deviation.
+HMMs are particularly useful in this setting as there is no direct knowledge of the guarding allocation.
+This is only partially observed based on the reaction of the defender to the offenders' movement.
+Naturally, the model thus treats the information on who is currently guarded as a latent state variable, and hence each observation as a realization from a mixture of Gaussian distributions (see Franks et al.
 2015 for a similar approach in basketball).
-
-This probabilistic approach allows us to infer dynamic coverage patterns during plays with pre-snap motion.
-
-OLE
+Ultimately, the goal is to use the fitted model for so-called *state-decoding*, i.e.
+inferring information on the guarding allocation based on the observations made.
+For this task, HMMs not only leverage information on the $y$-coordinates of the defender and all offenders potetially guarded at each time point, but also incorporate probabilistic information based on the previous and subsequent observations.
+For practical details on the model fitting and state decoding, also see the Appendix.
 
@@
 
+*Koslik J (2024). LaMa: Fast Numerical Maximum Likelihood Estimation for Latent Markov Models. R package version 2.0.2, <https://CRAN.R-project.org/package=LaMa>.
+
 *Zucchini W, MacDonald I, Langrock R (2016). Hidden Markov Models for Time Series - An Introduction Using R. CRC Press
@@
 
-#### Hidden Markov Model
-A hidden Markov models consists of an observed time series $\{\boldsymbol{y}_t\}_{t=1}^T$ --- here, the y-coordinates of the defensive players and an unobserved first-order Markov chain $\{ g_t\}_{t=1}^T$, with $g_t \in \{1,\ldots,N\}$ which proxies the offensives players to be guarded at every time point $t$.
-The Markov chain is fully described by an initial distribution $\boldsymbol{\delta}=\bigl( \Pr(g_1=1), \ldots, \Pr(g_1=N) \bigr)$ and a transition probability matrix (t.p.m.) $\boldsymbol{\Gamma} = (\gamma_{ij}),$ with $\gamma_{ij} = \Pr(g_t = j| g_{t-1} = i), \ i,j = 1, \ldots, N$.
-The connection of both stochastic processes arises from the assumption that the distribution of the observations $\boldsymbol{y}_t$ are fully determined by the currently active state, i.e.\
+#### Hidden Markov model
+
+A hidden Markov model consists of an observed time series $\{\boldsymbol{y}_t\}_{t=1}^T$ and an unobserved first-order Markov chain $\{ g_t\}_{t=1}^T$, with $g_t \in \{1,\ldots,N\}$.
+In this case, at every time point $t$, $\boldsymbol{y}_t$ is the y-coordinate of the defensive player and $g_t$ proxies the offensives player to be guarded, i.e.
+the guarding allocation.
+The Markov chain is fully described by an initial distribution $\boldsymbol{\delta}^{(1)} = \bigl( \Pr(g_1=1), \ldots, \Pr(g_1=N) \bigr)$ and a transition probability matrix (t.p.m.) $\boldsymbol{\Gamma} = (\gamma_{ij})$, with $\gamma_{ij} = \Pr(g_t = j \mid g_{t-1} = i), \ i,j = 1, \ldots, N$.
+The connection of both stochastic processes arises from the assumption that the distribution of the observations $\boldsymbol{y}_t$ are fully determined by the state that is currently active.
+More formally,
 \begin{equation*}
-f(\boldsymbol{y}_t|g_1, \ldots, g_T, \boldsymbol{y}_1, \ldots, \boldsymbol{y}_{t-1},\boldsymbol{y}_{t+1},\ldots,\boldsymbol{y}_T) = f(\boldsymbol{y}_t|g_t).
+f(\boldsymbol{y}_t \mid g_1, \ldots, g_T, \boldsymbol{y}_1, \ldots, \boldsymbol{y}_{t-1},\boldsymbol{y}_{t+1},\ldots,\boldsymbol{y}_T) = f(\boldsymbol{y}_t \mid g_t = j), \qquad j \in \{1, \ldots, N\},
 \end{equation*}
-In general, $f$ can be any density or probability mass function depending on the type of data. Following the approaches of Franks et al. (2015), we opt for a Gaussian distribution.
+which we denote by $f_j(\boldsymbol{y}_t)$ in short.
+In general, $f_j$ can be any density or probability mass function depending on the type of data and a typical choice is a parametric distribution with separate parameters for each latent state.
+Following the approaches of Franks et al.
+(2015), we opt for a Gaussian distribution with a mean that is fully determined by the current $y$-coordinate of defender $j \in \{1, \dots, N\}$ and a standard deviation that is fixed accross all states but estimated from the data.
+
+To fit the model, we use direct numerical likelihood maximization.
+The HMM likelihood for the motion of a specific defender in a specific play can be calculated based on the so-called *forward algorithm*.
+It effectively performs a summation over all possible latent state sequences in an efficient manner, rendering the computational complexity linear in the number of observations.
+Time series of different defenders within the same play, and of different plays are treated as independent, hence their likelihood contributions are summed to obtain the full likelihood of the training data.
+For practical implementation, we wrote a custom likelihood function in R, using the function `forward()` and other convenience functions from the R package `LaMa` (Koslik, 2024) to speed up computations.
+Furthermore, we used the R package `RTMB` to make this likelihood function compatible with automatic differentiation, making the numerical optimization process more efficient and robust.
+The parameters to be estimated are only the transition probability matrix $\boldsymbol{\Gamma}$ and the standard deviation of the Gaussian distribution. 
+The initial distribution of each defender for each play is fixed and computed based on ... ROUVEN
+
+Having fitted an HMM to the data, we can use the model to predict the underlying state sequence based on the observations.
+This process is called state decoding and two main approaches exist.
+So-called *local decoding* constructs the conditional distributions
+$$
+\Pr(g_t = j \mid \boldsymbol{y}_1, \dots, \boldsymbol{y}_T)
+$$
+while *global decoding* using the *Viterbi-algorithm* finds the state sequence that maximizes the joint probability of the state sequence given the observations.
+Local decoding retains more probabilistic information as it provides a categorical state distribution for each time point while global decoding is more suitable for visualization purposes as it provides a single state sequence that is most likely to have generated the observations.
+To obtain both the local state probabilities and the global state sequence, we used the functions `stateprobs()` and `viterbi()` that are also contained in the R package `LaMa`.

### 2025-01-03 HMM_plot.R
updating hmm plot

@@ -1,5 +1,6 @@
 library(dplyr)
 library(stringr)
+
 tracking_kc_az = read.csv("tracking_kc.csv")
 defender_data = read.csv("kc_defender_data.csv")
 
@@ -7,8 +8,23 @@ defender_data = read.csv("kc_defender_data.csv")
 field_length <- 120
 field_width <- 53.3
 
-team_colors = c("blue4", "black", "red")
+pitch_colors = c("#a4e86f", "#7eb356")
+team_colors = c("red", "orange", "white")#"blue4")
 los = 19
+
+clubs = levels(as.factor(tracking_kc_az$club))[levels(as.factor(tracking_kc_az$club)) != "football"]
+
+
+xlim = round(c(min(tracking_kc_az$x)-5, max(tracking_kc_az$x)+5), -1)
+# ylim = round(c(min(tracking_kc_az$y)-1, max(tracking_kc_az$y)+1), -1)
+ylim = c(0, field_width)
+asp_ratio =  diff(xlim) / diff(ylim)
+
+xlabs = c("<10", "<20", "<30", "<40", "50", "40>", "30>", "20>", "10>")
+
+par(pty = "m")  # Allow custom plot region
+par(pin = c(3, 3 / asp_ratio))  # Set fixed width and height (in inches)
+
 # Animation: Spielerbewegungen
 for (t in unique(tracking_kc_az$time)) {
   # Daten für den aktuellen Zeitpunkt filtern
@@ -17,17 +33,63 @@ for (t in unique(tracking_kc_az$time)) {
   # Spielfeld zeichnen
   plot(
     NA, 
-    xlim = c(min(tracking_kc_az$x)-1, max(tracking_kc_az$x)+1),#c(0, field_length), 
-    ylim = c(min(tracking_kc_az$y)-1, max(tracking_kc_az$y)+1),#c(0, field_width),
-    xlab = "Spielfeld-Länge", ylab = "Spielfeld-Breite",
-    main = paste("Spielerbewegungen - Zeit:", t), asp = 1
+    xlim = xlim, # c(min(tracking_kc_az$x)-1, max(tracking_kc_az$x)+1),#c(0, field_length), 
+    ylim = c(ylim[1]-5, ylim[2]), # c(min(tracking_kc_az$y)-1, max(tracking_kc_az$y)+1),#c(0, field_width),
+    # xlab = "Spielfeld-Länge", ylab = "Spielfeld-Breite",
[…]

### 2025-01-03 HMM_plot.R
udating coverage plot

@@ -1,5 +1,6 @@
 library(dplyr)
 library(stringr)
+library(animation)
 
 tracking_kc_az = read.csv("tracking_kc.csv")
 defender_data = read.csv("kc_defender_data.csv")
@@ -9,22 +10,29 @@ field_length <- 120
 field_width <- 53.3
 
 pitch_colors = c("#a4e86f", "#7eb356")
-team_colors = c("red", "orange", "white")#"blue4")
+team_colors = c("#97233F", "orange", "white")#"blue4")
 los = 19
 
 clubs = levels(as.factor(tracking_kc_az$club))[levels(as.factor(tracking_kc_az$club)) != "football"]
 
 
-xlim = round(c(min(tracking_kc_az$x)-5, max(tracking_kc_az$x)+5), -1)
+xlim = round(c(min(tracking_kc_az$x) - 8, 
+               max(tracking_kc_az$x)+5), - 1)
 # ylim = round(c(min(tracking_kc_az$y)-1, max(tracking_kc_az$y)+1), -1)
 ylim = c(0, field_width)
 asp_ratio =  diff(xlim) / diff(ylim)
 
-xlabs = c("<10", "<20", "<30", "<40", "50", "40>", "30>", "20>", "10>")
+xlabs = c("10", "20", "30", "40", "50", "40", "30", "20", "10")
 
 par(pty = "m")  # Allow custom plot region
 par(pin = c(3, 3 / asp_ratio))  # Set fixed width and height (in inches)
 
+
+ani.options(
+  res = 400
+)
+
+# saveGIF({
 # Animation: Spielerbewegungen
 for (t in unique(tracking_kc_az$time)) {
   # Daten für den aktuellen Zeitpunkt filtern
@@ -34,7 +42,7 @@ for (t in unique(tracking_kc_az$time)) {
   plot(
     NA, 
     xlim = xlim, # c(min(tracking_kc_az$x)-1, max(tracking_kc_az$x)+1),#c(0, field_length), 
-    ylim = c(ylim[1]-5, ylim[2]), # c(min(tracking_kc_az$y)-1, max(tracking_kc_az$y)+1),#c(0, field_width),
+    ylim = c(ylim[1]-8, ylim[2]), # c(min(tracking_kc_az$y)-1, max(tracking_kc_az$y)+1),#c(0, field_width),
     # xlab = "Spielfeld-Länge", ylab = "Spielfeld-Breite",
     xlab = "", ylab = "",
     # main = paste("Spielerbewegungen - Zeit:", t), 
[…]

### 2025-01-03 HMM_plot.R
updating plot

@@ -24,7 +24,8 @@ asp_ratio =  diff(xlim) / diff(ylim)
 
 xlabs = c("10", "20", "30", "40", "50", "40", "30", "20", "10")
 
-par(pty = "m")  # Allow custom plot region
+par(pty = "m",
+    mar = c(1,1,1,1))  # Allow custom plot region
 par(pin = c(3, 3 / asp_ratio))  # Set fixed width and height (in inches)
 
 
@@ -133,27 +134,34 @@ for (t in unique(tracking_kc_az$time)) {
   # football
   points(current_data$x[which(current_data$club == "football")],
          current_data$y[which(current_data$club == "football")],
+         pch = 18, col = "orange", cex = 1.2, lwd = 2)
+  points(current_data$x[which(current_data$club == "football")],
+         current_data$y[which(current_data$club == "football")],
          pch = 5, col = "black", cex = 1.2, lwd = 2)
   
   text(los + 3.5, ylim[1] + 4, "LOS", col = "#00000080", lwd = 2, cex = 0.9)
   
   # Box mit Spiel-Infos unten drunter
-  br = 25
-  rect(xlim[1], ylim[1] - 8, br, ylim[1], 
+  br = mean(xlim)
+  rect(xlim[1], ylim[1] - 8, xlim[2], ylim[1], 
        col = "#00000030", border = "black", lwd = 2)
-  rect(br, ylim[1] - 8, xlim[2], ylim[1], 
-       col = "#00000030", border = "black", lwd = 2)
-  # time
-  text(x = mean(c(xlim[1], br)), y = ylim[1] - 2.5, cex = 0.8,
-       labels = stringr::str_sub(t, 12, 19), col = "black", lwd = 2)
   # Spielstand
-  text(x = mean(c(xlim[1], br)), y = ylim[1] - 5.5, cex = 0.8,
+  text(x = mean(xlim), y = ylim[1] - 2.5, cex = 1,
        labels = "0:0", col = "black", lwd = 2)
   # teams
-  text(x = 4.5, y = ylim[1] - 4, 
-       labels = clubs[1], col = team_colors[1], lwd = 2, cex = 1.2)
-  text(x = 20.5, y = ylim[1] - 4, 
-       labels = clubs[2], col = team_colors[3], lwd = 2, cex = 1.2)
+  text(x = mean(c(xlim[1], br)) + 4, y = ylim[1] - 2.5, 
[…]

### 2025-01-04 submission.Rmd
.

@@
 In particular, we conducted the following feature engineering steps: We first consider all 11 players on each side of the field and compute features related to the convex hull of the positions of the players.
-In particular, for defense and offense, we compute the area spanned by the convex hull of all player such as well as the largest $y$ distance (i.e.\ the width of the hull) and the largest $x$ distance (i.e.
+In particular, for defense and offense, we compute the area spanned by the convex hull of all player such as well as the largest $y$ distance (i.e.
+the width of the hull) and the largest $x$ distance (i.e.
 the length of the hull).
@@
 inferring information on the guarding allocation based on the observations made.
-For this task, HMMs not only leverage information on the $y$-coordinates of the defender and all offenders potetially guarded at each time point, but also incorporate probabilistic information based on the previous and subsequent observations.
+For this task, HMMs not only leverage information on the $y$-coordinates of the defender and all offenders potentially guarded at each time point, but also incorporate probabilistic information based on the previous and subsequent observations.
 For practical details on the model fitting and state decoding, also see the Appendix.
@@
 However, a caveat of the aforemententioned state probabilities is that they constitute as high-dimensional time-series data which makes them hard to use them as features.
-An outlet of this problem is to employ suitable summary statistics (`r colorize("Robert hier spezifizieren welche du final genommen hast, ich gehe jetzt mal von summe an switches und entropy aus; Robert: Im moment nehmen ich alle 4 (sum, average, nr changes, average ent)...
+An outlet of this problem is to deduce suitable summary statistics that can serve as features in the existing model(`r colorize("Robert hier spezifizieren welche du final genommen hast, ich gehe jetzt mal von summe an switches und entropy aus; Robert: Im moment nehmen ich alle 4 (sum, average, nr changes, average ent)...
 Sollten wir nochmal besprechen.","orange")`).
-In particular, we calculate for each defender the most likely offensive player $k = 1, \ldots, 5$ to be guarded in each time point $t = 1, \ldots, T_n$.
-From this, we count the number of state switches and sum these numbers across defenders as a first, simple, summary statistics and further calculate across defenders $j = 1, \ldots, 5$ the mean entropy in offensive play $n = 1,\ldots, N$, i.e.:
+To overcome this, we first calculate for each defender the most likely offensive player $k = 1, \ldots, 5$ to be guarded in each time point $t = 1, \ldots, T_n$ and count the number of state switches, i.e.
+the number of times for which a defender's most likely guarded offensive players differs between two consecutive time points $s$ and $t$, $s>t$.
+From this, we can calculate some simple, summary statistics: 1) the sum of state switches, 2) the average number of state switches and 3) the number of defenders during a play that switch offensive players during a play.
+Besides, we use the aforementioned state decodings to calculate a more elaborated statistic, in particular the mean entropy across defenders $j = 1, \ldots, 5$ in each offensive play $n = 1,\ldots, N$, i.e.:
 
@@
 
-While we were clearly able to show that motion enables the defensive coverage scheme, a limitation of our approach lies in the imperfect prediction accuracy of the pre-motion model, primarily due to insufficient hyperparameter tuning and the relatively small number of plays involving motion.
+While we were clearly able to show that motion facilitates the defensive coverage scheme detection, a limitation of our approach lies in the imperfect prediction accuracy of the pre-motion model, primarily due to insufficient hyperparameter tuning and the relatively small number of plays involving motion.
 However, the primary focus of this project was on pre-snap motion, particularly on how to effectively translate this information into the hidden Markov model.
-Importantly, our pre-motion model is modular and can be seamlessly replaced by another model, such as the NFL Next Gen Stats model, which can then be integrated into our modeling extension to fully leverage the insights provided by our extention.
+Importantly, our pre-motion model is modular and can be seamlessly replaced by another model, such as the NFL Next Gen Stats model, which can then be integrated into our modeling extension to fully leverage the insights provided by our extension.
 
@@
 Following the approaches of Franks et al.
-(2015), we opt for a Gaussian distribution with a mean that is fully determined by the current $y$-coordinate of defender $j \in \{1, \dots, N\}$ and a standard deviation that is fixed accross all states but estimated from the data.
+(2015), we opt for a Gaussian distribution with a mean that is fully determined by the current $y$-coordinate of defender $j \in \{1, \dots, N\}$ and a standard deviation that is fixed across all states but estimated from the data.
 
@@
 The parameters to be estimated are only the transition probability matrix $\boldsymbol{\Gamma}$ and the standard deviation of the Gaussian distribution. 
-The initial distribution of each defender for each play is fixed and computed based on ... ROUVEN
+The initial distribution of each defender for each play comes from a deterministic assignment approach based on spatial proximity of the initial y-coordinates at the beginning of the time-series, i.e.
+the moment the line is set.
+Specifically, for each defender in a predefined set of potential defenders, we calculated their Euclidean distances to the five offensive players For each offense player, ranked in descending order by proximity, we assigned the closest defender to that player.
+The initial distribution was then set as a categorical distribution, where each defender was assigned a probability of 1 for their corresponding attacker and 0 for all others.
 

### 2025-01-04 submission.Rmd
test

@@
 While we similarly predict man- or zone coverage when the teams are set before snaps, we further leverage the additional information of pre-snap player movements.
-Using a hidden Markov model (HMM), we model defenders' trajectories based on hidden states, which represent the offensive players they may be guarding.
-Incorporating summary statistics of the probabilistic HMM results as features into the existing pre-motion model significantly improves both the AUC and detection accuracy and further allows for evaluating the effectiveness of pre-snap motion in uncovering defensive strategies, providing real-time tactical insights for coaches.
+For this, we first include rather naive post-motion features.
+Afterwards, we use a hidden Markov model (HMM) to model defenders' trajectories based on hidden states, which represent the offensive players they may be guarding.
+Incorporating summary statistics of the state decoding results as features into the existing models significantly improves both the AUC and log loss even more.
+This allows for evaluating the effectiveness of pre-snap motion in uncovering defensive strategies, providing real-time tactical insights for coaches.
 
@@
 Robert: ich denke das sollten wir noch besprechen.
-Ich habe ohnehin mehrere Modelle gefitted...","orange")`).
+Ich habe ohnehin mehrere Modelle gefitted, Rouven: Entscheide gerne selber","orange")`).
 Finally, we extract relevant information from play-by-play data, such as quarter, down, yards to go, home and away score and the game clock.
-
-Detailed information can be found in the Appendix.
 
@@
 
-#### 2. Hidden Markov model
+#### 2. Naive post-motion prediciton
 
-We describe the movements of the five defensive players during the phase of pre-snap motion by a hidden Markov model (see the Appendix for an in-depth description).
+We re-train the pre-motion model to predict whether the defense plays a man- or zone coverage scheme, but, in this step, we incorporate results from `r colorize("Robert hier du","orange")`
+
+#### 3. Post-motion prediction using a hidden Markov model
+
+In the last step, we aim to use results from a hidden Markov model as further features in the previously described models.
+For this, we describe the movements of the five defensive players during the phase of pre-snap motion by an HMM (see the Appendix for an in-depth description).
 In particular, we assume that each defender's $y$-coordinate at each time point $t$ is a realization from a particular Gaussian distribution with mean according to the $y$-coordinate of the offensive player that is currently guarded and an estimated standard deviation.
@@
 For practical details on the model fitting and state decoding, also see the Appendix.
+
+However, a caveat of the aforemententioned state probabilities is that they constitute as high-dimensional time-series data which makes them hard to use them as features.
+An outlet of this problem is to deduce suitable summary statistics that can serve as features in the existing model.
+To overcome this, we first calculate for each defender the most likely offensive player $k = 1, \ldots, 5$ to be guarded in each time point $t = 1, \ldots, T_n$ and count the number of state switches, i.e.
+the number of times for which a defender's most likely guarded offensive players differs between two consecutive time points $s$ and $t$, $s>t$.
+From this, we can calculate some simple, summary statistics: 1) the sum of state switches, 2) the average number of state switches and 3) the number of defenders during a play that switch offensive players during a play.
+Besides, we use the aforementioned state decodings to calculate a more elaborated statistic, in particular the mean entropy across defenders $j = 1, \ldots, 5$ in each offensive play $n = 1,\ldots, N$, i.e.:
+
+$$H(n) = - \frac{1}{5}\sum_{j = 1}^5 \sum_{k=1}^{5} \left( \frac{1}{T_n} \sum_{t=1}^{T_n} \mathbb{1}\left(\arg\max_{i=1,\ldots,5} X_{t,i} = k\right) \cdot \log\left(\frac{1}{T_n} \sum_{t=1}^{T_n} \mathbb{1}\left(\arg\max_{i=1,\ldots,5} X_{t,i} = k\right)\right) \right).$$
+Entropy is a measure of uncertainty or randomness in a probability distribution, here, the empirical distribution of the defenders guarding the respective offensive players.
+Higher entropy indicates greater unpredictability, while lower entropy signifies more predictability.
+While these four features are correlated, the models that we have chosen are able to cope with the problem of multicollinearity.
 
@@
 
-#### 3. Post-motion prediction
-We re-train the pre-motion model to predict whether the defense plays a man- or zone coverage scheme, but, in this step, we incorporate results from the HMM analysis as further features.
-However, a caveat of the aforemententioned state probabilities is that they constitute as high-dimensional time-series data which makes them hard to use them as features.
-An outlet of this problem is to deduce suitable summary statistics that can serve as features in the existing model(`r colorize("Robert hier spezifizieren welche du final genommen hast, ich gehe jetzt mal von summe an switches und entropy aus; Robert: Im moment nehmen ich alle 4 (sum, average, nr changes, average ent)...
-Sollten wir nochmal besprechen.","orange")`).
-To overcome this, we first calculate for each defender the most likely offensive player $k = 1, \ldots, 5$ to be guarded in each time point $t = 1, \ldots, T_n$ and count the number of state switches, i.e.
-the number of times for which a defender's most likely guarded offensive players differs between two consecutive time points $s$ and $t$, $s>t$.
-From this, we can calculate some simple, summary statistics: 1) the sum of state switches, 2) the average number of state switches and 3) the number of defenders during a play that switch offensive players during a play.
-Besides, we use the aforementioned state decodings to calculate a more elaborated statistic, in particular the mean entropy across defenders $j = 1, \ldots, 5$ in each offensive play $n = 1,\ldots, N$, i.e.:
+When performing state decoding, it becomes evident that the HMM effectively captures Marco Wilson's guarding assignment.
+The unique strength of HMMs is particularly apparent when Marco Wilson reaches the same height as the running back, Jerick McKinnon (KC #1).
+At this point, clustering algorithms that disregard the temporal component would briefly suggest a change in the guarding assignment, introducing noise into summary statistics such as the total number of switches predicted by the model.
+In contrast, the HMM consistently and accurately maintains the correct coverage assignment throughout the motion.
 
-$$H(n) = - \frac{1}{5}\sum_{j = 1}^5 \sum_{k=1}^{5} \left( \frac{1}{T_n} \sum_{t=1}^{T_n} \mathbb{1}\left(\arg\max_{i=1,\ldots,5} X_{t,i} = k\right) \cdot \log\left(\frac{1}{T_n} \sum_{t=1}^{T_n} \mathbb{1}\left(\arg\max_{i=1,\ldots,5} X_{t,i} = k\right)\right) \right).$$
-Entropy is a measure of uncertainty or randomness in a probability distribution, here, the empirical distribution of the defenders guarding the respective offensive players.
-Higher entropy indicates greater unpredictability, while lower entropy signifies more predictability.
+<center>
+  <img src="coverage.gif" width="500" height="500" alt="Coverage Prediction">
+</center>
 

### 2025-01-04 submission.Rmd
Sub

--- 2025-01-03 submission.Rmd
@@
 ##### Feature engineering
+
 To accurately forecast the defensive scheme (man- or zone defense) for every play, we need to create various features derived from the tracking data.
@@
 Ich habe ohnehin mehrere Modelle gefitted...","orange")`).
-Finally, we extract relevant information from play-by-play data, such as quarter, down, yards to go, home and away score and the game clock.
+Finally, we extract relevant information from play-by-play data, such as quarter, down, yards to go, home and away score and the remaining game time in the current half (in seconds).
 
@@
 
+<!-- NEW STRUCTURE -->
+
+`r colorize("New Structure still in developement","red")`
+
+We train different models to predict whether the defense plays a man- or zone coverage scheme.
+Since the aim of the project is to show the effectiveness of pre-snap motion, we follow a 3 step strategy:
+
+1. Pre-motion model
+2. Naive post-motion model
+3. Enhanced post-motion model
+
+In general, we have a limited dataset available (only 3985 plays) and therefore need to balance the complexity of the model, i.e.
+we want to control the number of features available.
+Using the previously described features, we obtain 67 variables basic features from tracking and play by play data, which can be used for all 3 models above.
+With a more profound dataset, using all of these features (and possibly more) certainly makes sense.
+However given the small dataset, we focus on only 32 basic features: 6 convex hull related features (3 for offense, 3 for defense), 20 player positions features (10 standardized $x$ and $y$ coordinates, 5 for offense, 5 for defense), and 6 play by play features.
+In the appendix, we provide more details and analyses using more features.
+Finally we need a suitable model class for predicting man vs zone coverage and chose the following two: On the one hand, we fit a `glmnet` model, which performs implicit feature selection and is able to handle multicollinearity.
+On the other hand we use an `xgboost` model, which is able to capture non linear effects (and interactions) and in general also handles collinearity, but needs careful hyperparameter tuning and generally performs better on bigger data sets.
+For all of the model we use careful 10-fold cross validation on a suitable hyperparameter grid.
+
+#### 1. Pre-motion prediction
+
+We fit a model with the previously described basic features, all of these features are derived at the time of line-set, so this model does not use motion information.
+This very basic model allows us to establish a baseline, that allows us to measure the effect of including motion features accurately.
+
+#### 2. Naive post-motion prediction
+
+We augment our basic pre-motion model with naive post-motion features.
+In order to maintain a manageable complexity, we only derive 6 additional post-motion features: for each team (offense and defense), we derive the maximum $y$ distance, the maximum $x$ distance, and the total distance traveled by this team up until the snap.
+
+#### 3. Enhancing post-motion predictions
+
+##### 3.1 Hidden Markov model
+
+We describe the movements of the five defensive players during the phase of pre-snap motion by a hidden Markov model (see the Appendix for an in-depth description).
+In particular, we assume that each defender's $y$-coordinate at each time point $t$ is a realization from a particular Gaussian distribution with mean according to the $y$-coordinate of the offensive player that is currently guarded and an estimated standard deviation.
+HMMs are particularly useful in this setting as there is no direct knowledge of the guarding allocation.
+This is only partially observed based on the reaction of the defender to the offenders' movement.
+Naturally, the model thus treats the information on who is currently guarded as a latent state variable, and hence each observation as a realization from a mixture of Gaussian distributions (see Franks et al.
+2015 for a similar approach in basketball).
+Ultimately, the goal is to use the fitted model for so-called *state-decoding*, i.e.
+inferring information on the guarding allocation based on the observations made.
+For this task, HMMs not only leverage information on the $y$-coordinates of the defender and all offenders potetially guarded at each time point, but also incorporate probabilistic information based on the previous and subsequent observations.
+For practical details on the model fitting and state decoding, also see the Appendix.
+
+The results of the HMM are exemplified using the following video.
+It displays a touchdown from the Kansas City Chiefs against the Arizona Cardinals in Week 1 of the 2022 NFL season.
+We can see that, pre-snap, Mecole Hardman (KC #17) is in motion.
+He is immediately followed by the defender Marco Wilson (AZ #20), which is a clear indication for man-coverage.
+
+<center>
+<video width="600" height="400" controls>
+  <source src="kc_az_td.mp4" type="video/mp4">
+</video>
+</center>
+
+##### 3.2 Enhanced post-motion model
+We re-train the post-motion model to predict whether the defense plays a man- or zone coverage scheme, but, in this step, we incorporate results from the HMM analysis as further features.
+However, a caveat of the aforemententioned state probabilities is that they constitute as high-dimensional time-series data which makes them hard to use them as features.
+An outlet of this problem is to employ suitable summary statistics (`r colorize("Robert hier spezifizieren welche du final genommen hast, ich gehe jetzt mal von summe an switches und entropy aus; Robert: Im moment nehme ich alle 4 (sum, average, nr changes, average ent)...
+Einfach alle bitte kurz beschreiben.","orange")`).
+In particular, we calculate for each defender the most likely offensive player $k = 1, \ldots, 5$ to be guarded in each time point $t = 1, \ldots, T_n$.
+From this, we count the number of state switches and sum these numbers across defenders as a first, simple, summary statistics and further calculate across defenders $j = 1, \ldots, 5$ the mean entropy in offensive play $n = 1,\ldots, N$, i.e.:
+
+$$H(n) = - \frac{1}{5}\sum_{j = 1}^5 \sum_{k=1}^{5} \left( \frac{1}{T_n} \sum_{t=1}^{T_n} \mathbb{1}\left(\arg\max_{i=1,\ldots,5} X_{t,i} = k\right) \cdot \log\left(\frac{1}{T_n} \sum_{t=1}^{T_n} \mathbb{1}\left(\arg\max_{i=1,\ldots,5} X_{t,i} = k\right)\right) \right).$$
+Entropy is a measure of uncertainty or randomness in a probability distribution, here, the empirical distribution of the defenders guarding the respective offensive players.
+Higher entropy indicates greater unpredictability, while lower entropy signifies more predictability.
+
+<!-- OLD STRUCTURE -->
+
+<!--
 Our analysis comprises different steps:
@@
 
-Since we have a limited dataset available (only 3985 plays), we want to balance the complexity of the model, i.e.
+Given the limited dataset available (only 3985 plays), we want to balance the complexity of the model, i.e.
 we control the number of features available.
@@
 `r colorize("6 convex hull related features (3 for offense, 3 for defense), 20 player positions features (10 standardized $x$ and $y$ coordinates, 5 for offense, 5 for defense), 30 distance features (10 Euclidian distances, 10 $x$ distances, and 10 $y$ distances to the football), 10 orientation features (5 offense, 5 defense) and 6 play by play features (probably better suited for appendix...)","orange")`.
-Given the small data set this is rather high number of features and therefore we consider two model classes.
+This is rather high number of features for such a small data set and therefore we consider two model classes.
 On the one hand, we fit an elastic net model, which performs implicit feature selection and is able to handle multicollinearity.
@@
 
+-->
+
+<!--
 #### 2. Hidden Markov model
@@
 </center>
+-->
 
+<!--
 #### 3. Post-motion prediction
@@
 Higher entropy indicates greater unpredictability, while lower entropy signifies more predictability.
+-->
 
@@
 
+#### Model and feature evaluation
+
+We compare the predictive performance of our 3 model to determine the effectiveness of player movements before snap for detecting the correct defensive scheme.
+We train our the models on a 85% of the data and evaluate performance on the remaining 15%.
+Since we use cross-validation for tuning of the models during the training phase, and this relies on randomly dividing the data into folds, we perform our evaluation repeatedly.
+We fix 50 different seeds (that is, for tuning, the data is split differently each of the 50 times), and for each we tune and train a model and then evaluate performance of this model on the test set.
+We do this in order to ensure that our results are meaningful and not due to chance (especially in our small data set).
+
+<!--
 By comparing the predictive performance of our model without motion and our post-motion model we can determine the effectiveness of player movements before the snap to detect the correct defensive scheme.
@@
 
-![](AUCs.png)
-![](loglosses.png)
+-->
 
-LEGENDE INFO:
+![](ACCs_bdb_main.png)
+![](AUCs_bdb_main.png)
 
-- Model af: alle features im model (pre motion mit distances und orientation, hmm features (average,sum,nr changes, average ent), naive postmotion, pbp)
-- Model af_ex_hmm: alle features außer hmm features (pre motion mit distances und orientation, naive postmotion, pbp)
-- Model af_npo: alle features außer naive postmotion features
-- Model af_npo_ex_hmm: alle features außer naive postmotion features und hmm features
-- Model xy_f: premotion features nur x,y coords der spieler (keine distances und orientation), sonst alle features
-- Model xy_f_ex_hmm: wie xy_f, nur ohne hmm features
-- Model xy_npo: wie xy_f, nur ohne naive postmotion features
-- Model xy_npo_ex_hmm: wie xy_npo, nur ohne hmm features
+#### Team analyses
 
+<!--
 robustheit checken mit simplen summarys nach der motion
@@
 team analysen
+-->
 
@@
 However, the primary focus of this project was on pre-snap motion, particularly on how to effectively translate this information into the hidden Markov model.
+`r colorize("Noch mehr Fokus auf Modelle und der Möglichkeit flexibler Modelle, wie die von AMAZON zu fitten, wenn mehr Daten dabei sind.","orange")`).
 Importantly, our pre-motion model is modular and can be seamlessly replaced by another model, such as the NFL Next Gen Stats model, which can then be integrated into our modeling extension to fully leverage the insights provided by our extention.
+`r colorize("Vergleiche unser Modell mit Results von Amazon: Unser Modell wesentlich weniger aufwändig, performt aber gut (Accuracy > 88 %.
+Accuracy von deren Modellen zwischen 88 und 96% obwohl sie 3 ganze Seasons an Daten haben (verglichen zu unseren 9 Wochen, wo wir 2/3 der plays entfernen müssen!)","orange")`).
 
@@
 ## Appendix
+
+#### Feature engineering and model features
+
+As frequently mentioned, we face the problem of having only a small amount of relevant data available.
+While we therefore focus on a small amount of features for our models for our main results, we provide more detailed results in this sections, using more thoroughly crafted features.
+
+`r colorize("Results go here...","orange")`
+
+We see that using the more pronounced set of features seems to provide more information.
+At least the `glmnet` model, which is more simple but known to be able to handle a high number of features well, the version including a bigger set of features (orientations and distances) performs better than the any other model.
+For the `xgboost` model, the model including most features performs on par with the one including only small set of basic features.
+This suggests, that with more data the model can still be improved.
+Finally, we mention that we also varied the number of players used in the analyses.
+Presently, we use only information of 5 offensive and 5 defensive players.
+However using more information resulted in worse performance of the models, and thus we refrain from showing the results here.
+`r colorize("Review this section!","orange")`
 

### 2025-01-04 submission.Rmd
Add changes from Rouven...

@@
 
-We describe the movements of the five defensive players during the phase of pre-snap motion by a hidden Markov model (see the Appendix for an in-depth description).
+In the last step, we aim to use results from a hidden Markov model as further features in the previously described models.
+For this, we describe the movements of the five defensive players during the phase of pre-snap motion by an HMM (see the Appendix for an in-depth description).
 In particular, we assume that each defender's $y$-coordinate at each time point $t$ is a realization from a particular Gaussian distribution with mean according to the $y$-coordinate of the offensive player that is currently guarded and an estimated standard deviation.
@@
 Einfach alle bitte kurz beschreiben.","orange")`).
-In particular, we calculate for each defender the most likely offensive player $k = 1, \ldots, 5$ to be guarded in each time point $t = 1, \ldots, T_n$.
-From this, we count the number of state switches and sum these numbers across defenders as a first, simple, summary statistics and further calculate across defenders $j = 1, \ldots, 5$ the mean entropy in offensive play $n = 1,\ldots, N$, i.e.:
+To overcome this, we first calculate for each defender the most likely offensive player $k = 1, \ldots, 5$ to be guarded in each time point $t = 1, \ldots, T_n$ and count the number of state switches, i.e.
+the number of times for which a defender's most likely guarded offensive players differs between two consecutive time points $s$ and $t$, $s>t$.
+From this, we can calculate some simple, summary statistics: 1) the sum of state switches, 2) the average number of state switches and 3) the number of defenders during a play that switch offensive players during a play.
+Besides, we use the aforementioned state decodings to calculate a more elaborated statistic, in particular the mean entropy across defenders $j = 1, \ldots, 5$ in each offensive play $n = 1,\ldots, N$, i.e.:
 
@@
 Higher entropy indicates greater unpredictability, while lower entropy signifies more predictability.
+
+When performing state decoding, it becomes evident that the HMM effectively captures Marco Wilson's guarding assignment.
+The unique strength of HMMs is particularly apparent when Marco Wilson reaches the same height as the running back, Jerick McKinnon (KC #1).
+At this point, clustering algorithms that disregard the temporal component would briefly suggest a change in the guarding assignment, introducing noise into summary statistics such as the total number of switches predicted by the model.
+In contrast, the HMM consistently and accurately maintains the correct coverage assignment throughout the motion.
+
+<center>
+  <img src="coverage.gif" width="500" height="500" alt="Coverage Prediction">
+</center>
 
@@
 
-While we were clearly able to show that motion enables the defensive coverage scheme, a limitation of our approach lies in the imperfect prediction accuracy of the pre-motion model, primarily due to insufficient hyperparameter tuning and the relatively small number of plays involving motion.
+While we were clearly able to show that motion facilitates the defensive coverage scheme detection, a limitation of our approach lies in the imperfect prediction accuracy of the pre-motion model, primarily due to insufficient hyperparameter tuning and the relatively small number of plays involving motion.
 However, the primary focus of this project was on pre-snap motion, particularly on how to effectively translate this information into the hidden Markov model.
 `r colorize("Noch mehr Fokus auf Modelle und der Möglichkeit flexibler Modelle, wie die von AMAZON zu fitten, wenn mehr Daten dabei sind.","orange")`).
-Importantly, our pre-motion model is modular and can be seamlessly replaced by another model, such as the NFL Next Gen Stats model, which can then be integrated into our modeling extension to fully leverage the insights provided by our extention.
+Importantly, our pre-motion model is modular and can be seamlessly replaced by another model, such as the NFL Next Gen Stats model, which can then be integrated into our modeling extension to fully leverage the insights provided by our extension.
 `r colorize("Vergleiche unser Modell mit Results von Amazon: Unser Modell wesentlich weniger aufwändig, performt aber gut (Accuracy > 88 %.

### 2025-01-05 submission.Rmd
Sub

@@
 Finally, we use their standardized $x$ and $y$ coordinates as features and order defensive and offensive players according to their $y$ coordinates.
-Additionally, for each of the relevant defenders, we compute distances to the football and their orientation with respect to the quarterback (`r colorize("values taken at event lineset.
-Rouven: Die Values haben wir doch rausgenommen aus dem finalen Modell oder?
-Robert: ich denke das sollten wir noch besprechen.
-Ich habe ohnehin mehrere Modelle gefitted...","orange")`).
+Additionally, for each of the relevant defenders, we compute distances to the football and their orientation with respect to the quarterback.
 Finally, we extract relevant information from play-by-play data, such as quarter, down, yards to go, home and away score and the remaining game time in the current half (in seconds).
@@
 In the appendix, we provide more details and analyses using more features.
-Finally we need a suitable model class for predicting man vs zone coverage and chose the following two: On the one hand, we fit a `glmnet` model, which performs implicit feature selection and is able to handle multicollinearity.
+Finally we need a suitable model class for predicting man vs zone coverage and chose the following two: On the one hand, we fit a `glmnet` (elastic net) model, which performs implicit feature selection and is able to handle multicollinearity.
 On the other hand we use an `xgboost` model, which is able to capture non linear effects (and interactions) and in general also handles collinearity, but needs careful hyperparameter tuning and generally performs better on bigger data sets.
@@
 
-We compare the predictive performance of our 3 model to determine the effectiveness of player movements before snap for detecting the correct defensive scheme.
-We train our the models on a 85% of the data and evaluate performance on the remaining 15%.
+We compare the predictive performance of our 3 models to determine the effectiveness of player movements before snap for detecting the correct defensive scheme.
+We train our the models on 85% of the data and evaluate performance on the remaining 15%.
 Since we use cross-validation for tuning of the models during the training phase, and this relies on randomly dividing the data into folds, we perform our evaluation repeatedly.
-We fix 50 different seeds (that is, for tuning, the data is split differently each of the 50 times), and for each we tune and train a model and then evaluate performance of this model on the test set.
+We fix 50 different random seeds (that is, for tuning, the data is split differently each of the 50 times), and for each we tune and train a model and then evaluate performance of this model on the test set.
 We do this in order to ensure that our results are meaningful and not due to chance (especially in our small data set).
@@
 ![](AUCs_bdb_main.png)
+
+The plots above show the results for our 50 repetitions of the experiment.
+First, we show the classification accuracy, when using a threshold of 50%.
+Since this is an arbitrary value and our data is slightly imbalanced (~ 75% zone coverage plays and ~25% man coverage plays), accuracy is not the best metric for analyzing the results.
+Therefore we also present the AUC values of the model (`r colorize("We also provide further metrics in the Appendix","orange")`).
+For both metrics, we notice a few thing.
+On the one hand, the more flexible `xgboost` model performs better than the `glmnet`.
+However, there is more variation (larger boxes) for the `xgboost` model, suggesting that more data would help improve the tuning of the models.
+On the other hand, the pre-motion model clearly performs worst.
+As soon as we incorporate motion information, there is a large increase in performance.
+While this is expected, we also see that the enhanced model outperforms the model with only naive post-motion information.
+This shows, that the HMM features capture information, which is distinct from simple motion features and provide another dimension of information, which needs to be incorporated.
 

### 2025-01-05 HMM_plot.R
adding animation

@@ -1,6 +1,8 @@
 library(dplyr)
 library(stringr)
-library(animation)
+# library(animation)
+library(magick)
+# dir.create("animation/frames")
 
 tracking_kc_az = read.csv("tracking_kc.csv")
 defender_data = read.csv("kc_defender_data.csv")
@@ -15,7 +17,6 @@ los = 19
 
 clubs = levels(as.factor(tracking_kc_az$club))[levels(as.factor(tracking_kc_az$club)) != "football"]
 
-
 xlim = round(c(min(tracking_kc_az$x) - 8, 
                max(tracking_kc_az$x)+5), - 1)
 # ylim = round(c(min(tracking_kc_az$y)-1, max(tracking_kc_az$y)+1), -1)
@@ -24,18 +25,19 @@ asp_ratio =  diff(xlim) / diff(ylim)
 
 xlabs = c("10", "20", "30", "40", "50", "40", "30", "20", "10")
 
-par(pty = "m",
-    mar = c(1,1,1,1))  # Allow custom plot region
-par(pin = c(3, 3 / asp_ratio))  # Set fixed width and height (in inches)
-
-
-ani.options(
-  res = 400
-)
+imInd = 1
 
 # saveGIF({
 # Animation: Spielerbewegungen
 for (t in unique(tracking_kc_az$time)) {
+  
+  pdf(file = sprintf("./animation/frames/frame%03d.pdf", imInd), width = 3, height = 3 / asp_ratio)
+  
+  par(pty = "m",
+      mar = c(1,1,1,1))  # Allow custom plot region
+  par(pin = c(3, 3 / asp_ratio))  # Set fixed width and height (in inches)
+  
+  
   # Daten für den aktuellen Zeitpunkt filtern
   current_data <- tracking_kc_az[tracking_kc_az$time == t, ]
   
@@ -144,7 +146,7 @@ for (t in unique(tracking_kc_az$time)) {
   # Box mit Spiel-Infos unten drunter
   br = mean(xlim)
   rect(xlim[1], ylim[1] - 8, xlim[2], ylim[1], 
-       col = "#00000030", border = "black", lwd = 2)
+       col = "gray80", border = "black", lwd = 2)
   # Spielstand
   text(x = mean(xlim), y = ylim[1] - 2.5, cex = 1,
        labels = "0:0", col = "black", lwd = 2)
@@ -164,6 +166,21 @@ for (t in unique(tracking_kc_az$time)) {
   
   
   # Pause für Animation
[…]

### 2025-01-05 HMM_plot.R
updating animation

@@ -59,14 +59,16 @@ for (t in unique(tracking_kc_az$time)) {
   # draw the endzones if necessary
   if(xlim[1] < 10){
     rect(xlim[1], ylim[1], 10, ylim[2], 
-         col = scales::alpha(team_colors[1], 0.7), 
+         # col = scales::alpha(team_colors[1], 0.7), 
+         col = colorspace::lighten(team_colors[1], 0.3),
          border = NA)
     text(5, mean(ylim), "ARIZONA", srt = 90, 
-         col = scales::alpha(team_colors[1], 0.4), cex = 2.25)
+         col = team_colors[1], cex = 2.25)
   }
   if(xlim[2] > 110){
     rect(110, ylim[1], xlim[2], ylim[2], 
-         col = scales::alpha(team_colors[3], 0.7), 
+         #col = scales::alpha(team_colors[3], 0.7), 
+         col = colorspace::lighten(team_colors[2], 0.3),
          border = NA)
   }
   
@@ -80,7 +82,8 @@ for (t in unique(tracking_kc_az$time)) {
       right = xlim[2]
     }
     rect(i, ylim[1], right, ylim[2], 
-         col = scales::alpha(pitch_colors[i %% 2 + 1], 0.5), 
+         # col = scales::alpha(pitch_colors[i %% 2 + 1], 0.5), 
+         col = colorspace::lighten(pitch_colors[i %% 2 + 1], 0.6),
          border = NA)
     
     if(i >= 20 & i %% 10 == 0 & i < 100){
@@ -180,7 +183,18 @@ for (t in unique(tracking_kc_az$time)) {
 # Convert PDF to GIF
 frames <- list.files("./animation/frames", full.names = TRUE, pattern = "*.pdf")
 
-# Use magick to combine frames into an animated GIF
-animation <- image_read(frames, density = "300x300")  # Read all PDF frames
-animation <- image_animate(animation, fps = 10)  # Set frame rate (10 FPS)
+# Convert PDF frames to PNG
+png_frames <- lapply(frames, function(frame) {
+  image <- image_read(frame, density = "250")
+  image <- image_convert(image, format = "png")
+  png_file <- sub(".pdf$", ".png", frame)
+  image_write(image, png_file)
+  return(png_file)
+})
+
+png_frames = unlist(png_frames)
+
+# Read PNG frames and create an animated GIF
+animation <- image_read(png_frames)
+animation <- image_animate(animation, fps = 10)
 image_write(animation, "./animation/animation.gif")

### 2025-01-05 HMM_plot.R
updating animation

@@ -19,6 +19,7 @@ clubs = levels(as.factor(tracking_kc_az$club))[levels(as.factor(tracking_kc_az$c
 
 xlim = round(c(min(tracking_kc_az$x) - 8, 
                max(tracking_kc_az$x)+5), - 1)
+xlim[2] = 50
 # ylim = round(c(min(tracking_kc_az$y)-1, max(tracking_kc_az$y)+1), -1)
 ylim = c(0, field_width)
 asp_ratio =  diff(xlim) / diff(ylim)
@@ -185,7 +186,7 @@ frames <- list.files("./animation/frames", full.names = TRUE, pattern = "*.pdf")
 
 # Convert PDF frames to PNG
 png_frames <- lapply(frames, function(frame) {
-  image <- image_read(frame, density = "250")
+  image <- image_read(frame, density = "300")
   image <- image_convert(image, format = "png")
   png_file <- sub(".pdf$", ".png", frame)
   image_write(image, png_file)
@@ -198,3 +199,4 @@ png_frames = unlist(png_frames)
 animation <- image_read(png_frames)
 animation <- image_animate(animation, fps = 10)
 image_write(animation, "./animation/animation.gif")
+

### 2025-01-05 submission.Rmd
updating animation
shortened

@@
 ## Introduction
-Deciding between man and zone coverage is one of the most critical strategic choices a defensive coordinator must make before each offensive play in American football.
+Deciding between man and zone coverage is one of the most critical strategic decisions a defensive coordinator must take before each offensive play in American football.
 While experienced offensive coordinators and quarterbacks often rely on visual cues to identify these defensive schemes, the increasing availability of player tracking data offers a new avenue to uncover and analyze these tactics.
-A notable example is Amazon's NFL Next Gen Stats model, which delivers coverage predictions during live broadcasts (see a snapshot of the 2024 Week 12 matchup between the Pittsburgh Steelers and Cleveland Browns).
-However, these models seem to be trained on plays without pre-snap motion, or at least to the situations before motions (see [Amazon](https://aws.amazon.com/de/blogs/machine-learning/identifying-defense-coverage-schemes-in-nfls-next-gen-stats/)), while motion is a crucial element of modern offensive strategies.
+A notable example is Amazon's NFL Next Gen Stats model, which delivers coverage predictions during live broadcasts (see a snapshot of the 2024 Week 12 match between the Pittsburgh Steelers and Cleveland Browns).
+However, in these pre-snap motion does not seem to play an accentuated role (see [Amazon](https://aws.amazon.com/de/blogs/machine-learning/identifying-defense-coverage-schemes-in-nfls-next-gen-stats/)), although it is a crucial element of modern offensive strategies and it is often used to detect the hidden defensive scheme.
 
-Our project takes this model a step further.
-While we similarly predict man- or zone coverage when the teams are set before snaps, we further leverage the additional information of pre-snap player movements.
+Hence, our project goes one step further.
+While we similarly predict man- or zone coverage when the teams are set before snaps, we further leverage the additional information of pre-snap player movements in a specific manner.
 For this, we first include rather naive post-motion features.
 Afterwards, we use a hidden Markov model (HMM) to model defenders' trajectories based on hidden states, which represent the offensive players they may be guarding.
-Incorporating summary statistics of the state decoding results as features into the existing models significantly improves both the AUC and log loss even more.
-This allows for evaluating the effectiveness of pre-snap motion in uncovering defensive strategies, providing real-time tactical insights for coaches.
+Incorporating summary statistics of the state decoding results as features into the existing models significantly improves the predictive ability.
+This opens up the avenue for further analysis such as the evaluation of the effectiveness of pre-snap motion in uncovering defensive strategies.
 
@@
 ## Data
-We analyze tracking data from nine weeks of the NFL 2022 season, provided by the NFL Big Data Bowl 2024.
-Beside the tracking data, we also use information on plays and players.
-We further considered the corresponding data from PFF that assigned the categories \textit{man}, \textit{zone} and \textit{others} representing the different schemes to each play.
+We analyze tracking data from the first nine weeks of the NFL 2022 season, provided by the NFL Big Data Bowl 2024.
+Moreover, we use play-by-play data and the corresponding data from PFF that assigned the categories \textit{man}, \textit{zone} and \textit{others} representing the different schemes to each play.
 As it is not properly described what \textit{others} means, we omit every play that is associated with this value.
-Moreover, we omit plays with more than five offensive linemen and with two quarterbacks.
-Since we are specifically interested in analyzing pre-snap player movements, we omit plays that did not contain any pre-snap motion.
+We also omit plays with more than five offensive linemen and with two quarterbacks.
+Since we are specifically interested in analyzing pre-snap player movements, we concentrate on plays that contain any pre-snap motion.
 Then, we end up with $3985$ offensive plays in total, from which the defense played $2973$ in zone and $1012$ in man coverage.
@@
 In addition, we select the five most relevant players on each side of the field.
-For offense, we omit the offensive line and the QB, while, for defense, we omit nose tackles, defensive tackles and defensive ends, and select the five defenders that were the closest to the five attackers corresponding to a weighted euclidean distance, putting much more emphasis on the y-axis.
+For offense, we omit the offensive line and the QB, thus coming up with five players that are a composition of running backs, tight ends and wide receivers.
+In contrast, for defense, we omit nose tackles, defensive tackles and defensive ends, and select the five defenders that were the closest to the five offensive players, corresponding to a weighted euclidean distance, putting much more emphasis on the y-axis.
 Finally, we use their standardized $x$ and $y$ coordinates as features and order defensive and offensive players according to their $y$ coordinates.
@@
 
-`r colorize("New Structure still in developement","red")`
+We train different models to predict whether the defense plays a man- or zone coverage scheme.
+Since the aim of the project is to show the effectiveness of pre-snap motion, we follow a three-step approach:
 
-We train different models to predict whether the defense plays a man- or zone coverage scheme.
-Since the aim of the project is to show the effectiveness of pre-snap motion, we follow a 3 step strategy:
+1. Pre-motion models
+2. Naive post-motion models
+3. HMM post-motion models
 
-1. Pre-motion model
-2. Naive post-motion model
-3. Enhanced post-motion model
+`r colorize("der erste Teil ist doch eigentlich eine Wiederholung des Feature Engeneerings?","red")`
+In general, we have a limited dataset available (only 3985 plays) and therefore need to balance the complexity of the model, thus aiming to control the number of features.
+Given the small dataset, we focus on 32 previously described basic features: 6 convex hull related features (3 for offense and defense, respectively), 20 player positions features (10 standardized $x$ and $y$ coordinates, 5 for offense, 5 for defense), and 6 play-by-play features.
+In the Appendix, we provide more details and analyses using more features.
 
-In general, we have a limited dataset available (only 3985 plays) and therefore need to balance the complexity of the model, i.e.
-we want to control the number of features available.
-Using the previously described features, we obtain 67 variables basic features from tracking and play by play data, which can be used for all 3 models above.
-With a more profound dataset, using all of these features (and possibly more) certainly makes sense.
-However given the small dataset, we focus on only 32 basic features: 6 convex hull related features (3 for offense, 3 for defense), 20 player positions features (10 standardized $x$ and $y$ coordinates, 5 for offense, 5 for defense), and 6 play by play features.
-In the appendix, we provide more details and analyses using more features.
-Finally we need a suitable model class for predicting man vs zone coverage and chose the following two: On the one hand, we fit a `glmnet` (elastic net) model, which performs implicit feature selection and is able to handle multicollinearity.
-On the other hand we use an `xgboost` model, which is able to capture non linear effects (and interactions) and in general also handles collinearity, but needs careful hyperparameter tuning and generally performs better on bigger data sets.
-For all of the model we use careful 10-fold cross validation on a suitable hyperparameter grid.
+#### 1. Pre-motion models
 
-#### 1. Pre-motion prediction
+First of all, we need a suitable basic model class for predicting man or zone coverage and opt for the following two: First, we fit a `glmnet` (elastic net) model, which performs implicit feature selection and is able to handle multicollinearity.
+Second, we use an `xgboost` model, which is able to capture non-linear effects (and interactions) and also handles multicollinearity.
+However, it necessitates careful hyperparameter tuning and generally performs better on bigger data sets.
+For all of the models, we use 10-fold cross validation on a suitable hyperparameter grid.
 
-We fit a model with the previously described basic features, all of these features are derived at the time of line-set, so this model does not use motion information.
-This very basic model allows us to establish a baseline, that allows us to measure the effect of including motion features accurately.
+We fit the models with the previously described basic features.
+All of these features are derived at the time of line-set, which is why these models do not use any pre-snap motion information.
+These very basic models serve as baseline models that allows to measure the effect of pre-snap motion (features) in the following.
 
-#### 2. Naive post-motion prediction
+#### 2. Naive post-motion models
 
-We augment our basic pre-motion model with naive post-motion features.
-In order to maintain a manageable complexity, we only derive 6 additional post-motion features: for each team (offense and defense), we derive the maximum $y$ distance, the maximum $x$ distance, and the total distance traveled by this team up until the snap.
+In a second step, we extend our basic pre-motion model with naive post-motion features.
+To keep the complexity manageable, we derive only six additional post-motion features: for each team (offense and defense), we infer the maximum $y$-distance, the maximum $x$-distance and the total distance traveled by both teams until the snap.
 
-#### 3. Enhancing post-motion predictions
+#### 3. HMM post-motion models
 
@@
 
-In the last step, we aim to use results from a hidden Markov model as further features in the previously described models.
+The last step is as the heart of our approach.
+In particular, we use results from a hidden Markov model as further features in the previously described models.
 For this, we describe the movements of the five defensive players during the phase of pre-snap motion by an HMM (see the Appendix for an in-depth description).
@@
 
-##### 3.2 Enhanced post-motion model
-We re-train the post-motion model to predict whether the defense plays a man- or zone coverage scheme, but, in this step, we incorporate results from the HMM analysis as further features.
-However, a caveat of the aforemententioned state probabilities is that they constitute as high-dimensional time-series data which makes them hard to use them as features.
-An outlet of this problem is to employ suitable summary statistics (`r colorize("Robert hier spezifizieren welche du final genommen hast, ich gehe jetzt mal von summe an switches und entropy aus; Robert: Im moment nehme ich alle 4 (sum, average, nr changes, average ent)...
-Einfach alle bitte kurz beschreiben.","orange")`).
-To overcome this, we first calculate for each defender the most likely offensive player $k = 1, \ldots, 5$ to be guarded in each time point $t = 1, \ldots, T_n$ and count the number of state switches, i.e.
-the number of times for which a defender's most likely guarded offensive players differs between two consecutive time points $s$ and $t$, $s>t$.
-From this, we can calculate some simple, summary statistics: 1) the sum of state switches, 2) the average number of state switches and 3) the number of defenders during a play that switch offensive players during a play.
-Besides, we use the aforementioned state decodings to calculate a more elaborated statistic, in particular the mean entropy across defenders $j = 1, \ldots, 5$ in each offensive play $n = 1,\ldots, N$, i.e.:
-
-$$H(n) = - \frac{1}{5}\sum_{j = 1}^5 \sum_{k=1}^{5} \left( \frac{1}{T_n} \sum_{t=1}^{T_n} \mathbb{1}\left(\arg\max_{i=1,\ldots,5} X_{t,i} = k\right) \cdot \log\left(\frac{1}{T_n} \sum_{t=1}^{T_n} \mathbb{1}\left(\arg\max_{i=1,\ldots,5} X_{t,i} = k\right)\right) \right).$$
-Entropy is a measure of uncertainty or randomness in a probability distribution, here, the empirical distribution of the defenders guarding the respective offensive players.
-Higher entropy indicates greater unpredictability, while lower entropy signifies more predictability.
-
 When performing state decoding, it becomes evident that the HMM effectively captures Marco Wilson's guarding assignment.
@@
 <center>
-  <img src="coverage.gif" width="500" height="500" alt="Coverage Prediction">
+  <img src="./animation/animation.gif" width="470" height="500" alt="Coverage Prediction according to a hidden Markov model.">
 </center>
 
-<!-- OLD STRUCTURE -->
+##### 3.2 Enhanced post-motion model
+We re-train the post-motion model to predict whether the defense plays a man- or zone coverage scheme, but, in this step, we incorporate results from the HMM analysis as further features.
+However, a caveat of the aforemententioned state probabilities is that they constitute as high-dimensional time-series data which makes them hard to use as features in the previously described models.
+An outlet of this problem is to employ suitable summary statistics.
+To overcome this, we first calculate for each defender the most likely offensive player $k = 1, \ldots, 5$ to be guarded in each time point $t = 1, \ldots, T_n$, where $T_n$ is the length of the $n$-th play, and count the number of state switches, i.e.
+the number of times for which a defender's most likely guarded offensive players differs between two consecutive time points $s$ and $t$, $s>t$.
+From this, we can calculate some simple, summary statistics: 1) the sum of state switches, 2) the average number of state switches and 3) the number of defenders that switch offensive players during a play.
+Fourth, we use the aforementioned state decodings to calculate a more elaborated statistic, the mean entropy across defenders $j = 1, \ldots, 5$ in each offensive play $n = 1,\ldots, N$, i.e.:
 
-<!--
-Our analysis comprises different steps:
+$$H(n) = - \frac{1}{5}\sum_{j = 1}^5 \sum_{k=1}^{5} \left( \frac{1}{T_n} \sum_{t=1}^{T_n} \mathbb{1}\left(\arg\max_{i=1,\ldots,5} X_{t,i} = k\right) \cdot \log\left(\frac{1}{T_n} \sum_{t=1}^{T_n} \mathbb{1}\left(\arg\max_{i=1,\ldots,5} X_{t,i} = k\right)\right) \right),$$
 
-#### 1. Pre-motion prediction
-We train different models to predict whether the defense plays a man- or zone coverage scheme.
-In particular, we start with fitting a model only using features available pre-motion.
-Since the aim of the project is to show the effectiveness of pre-snap motion, this allows us to establish a baseline model for detecting zone vs man coverage without motion.
-
-Given the limited dataset available (only 3985 plays), we want to balance the complexity of the model, i.e.
-we control the number of features available.
-Using the previously described features, we obtain 67 variables.
-`r colorize("6 convex hull related features (3 for offense, 3 for defense), 20 player positions features (10 standardized $x$ and $y$ coordinates, 5 for offense, 5 for defense), 30 distance features (10 Euclidian distances, 10 $x$ distances, and 10 $y$ distances to the football), 10 orientation features (5 offense, 5 defense) and 6 play by play features (probably better suited for appendix...)","orange")`.
-This is rather high number of features for such a small data set and therefore we consider two model classes.
-On the one hand, we fit an elastic net model, which performs implicit feature selection and is able to handle multicollinearity.
-On the other hand we use an xgboost model, which is able to capture non linear effects (and interactions) and in general also handles collinearity, but needs careful hyperparameter tuning and generally performs better on bigger data sets.
-
--->
-
-<!--
-#### 2. Hidden Markov model
-
-We describe the movements of the five defensive players during the phase of pre-snap motion by a hidden Markov model (see the Appendix for an in-depth description).
-In particular, we assume that each defender's $y$-coordinate at each time point $t$ is a realization from a particular Gaussian distribution with mean according to the $y$-coordinate of the offensive player that is currently guarded and an estimated standard deviation.
-HMMs are particularly useful in this setting as there is no direct knowledge of the guarding allocation.
-This is only partially observed based on the reaction of the defender to the offenders' movement.
-Naturally, the model thus treats the information on who is currently guarded as a latent state variable, and hence each observation as a realization from a mixture of Gaussian distributions (see Franks et al.
-2015 for a similar approach in basketball).
-Ultimately, the goal is to use the fitted model for so-called *state-decoding*, i.e.
-inferring information on the guarding allocation based on the observations made.
-For this task, HMMs not only leverage information on the $y$-coordinates of the defender and all offenders potetially guarded at each time point, but also incorporate probabilistic information based on the previous and subsequent observations.
-For practical details on the model fitting and state decoding, also see the Appendix.
-
-The results of the HMM are exemplified using the following video.
-It displays a touchdown from the Kansas City Chiefs against the Arizona Cardinals in Week 1 of the 2022 NFL season.
-We can see that, pre-snap, Mecole Hardman (KC #17) is in motion.
-He is immediately followed by the defender Marco Wilson (AZ #20), which is a clear indication for man-coverage.
-
-<center>
-<video width="600" height="400" controls>
-  <source src="kc_az_td.mp4" type="video/mp4">
-</video>
-</center>
--->
-
-<!--
-#### 3. Post-motion prediction
-We re-train the pre-motion model to predict whether the defense plays a man- or zone coverage scheme, but, in this step, we incorporate results from the HMM analysis as further features.
-However, a caveat of the aforemententioned state probabilities is that they constitute as high-dimensional time-series data which makes them hard to use them as features.
-An outlet of this problem is to employ suitable summary statistics (`r colorize("Robert hier spezifizieren welche du final genommen hast, ich gehe jetzt mal von summe an switches und entropy aus; Robert: Im moment nehmen ich alle 4 (sum, average, nr changes, average ent)...
-Sollten wir nochmal besprechen.","orange")`).
-In particular, we calculate for each defender the most likely offensive player $k = 1, \ldots, 5$ to be guarded in each time point $t = 1, \ldots, T_n$.
-From this, we count the number of state switches and sum these numbers across defenders as a first, simple, summary statistics and further calculate across defenders $j = 1, \ldots, 5$ the mean entropy in offensive play $n = 1,\ldots, N$, i.e.:
-
-$$H(n) = - \frac{1}{5}\sum_{j = 1}^5 \sum_{k=1}^{5} \left( \frac{1}{T_n} \sum_{t=1}^{T_n} \mathbb{1}\left(\arg\max_{i=1,\ldots,5} X_{t,i} = k\right) \cdot \log\left(\frac{1}{T_n} \sum_{t=1}^{T_n} \mathbb{1}\left(\arg\max_{i=1,\ldots,5} X_{t,i} = k\right)\right) \right).$$
-Entropy is a measure of uncertainty or randomness in a probability distribution, here, the empirical distribution of the defenders guarding the respective offensive players.
-Higher entropy indicates greater unpredictability, while lower entropy signifies more predictability.
--->
+as the entropy is a measure of uncertainty or randomness in a probability distribution. Higher entropy indicates greater unpredictability, while lower entropy signifies more predictability.
 
@@
 
-We compare the predictive performance of our 3 models to determine the effectiveness of player movements before snap for detecting the correct defensive scheme.
+We compare the predictive performance of our three models to determine the effectiveness of player movements before snap for detecting the correct defensive scheme.
 We train our the models on 85% of the data and evaluate performance on the remaining 15%.
 Since we use cross-validation for tuning of the models during the training phase, and this relies on randomly dividing the data into folds, we perform our evaluation repeatedly.
-We fix 50 different random seeds (that is, for tuning, the data is split differently each of the 50 times), and for each we tune and train a model and then evaluate performance of this model on the test set.
-We do this in order to ensure that our results are meaningful and not due to chance (especially in our small data set).
-
-<!--
-By comparing the predictive performance of our model without motion and our post-motion model we can determine the effectiveness of player movements before the snap to detect the correct defensive scheme.
-Moreover, we assess which teams predominantly apply pre-snap motions to increase the likelihood of correctly identifying the applied defensive strategy.
-
-auc da unbalanced data
-Hier die Animation rein mit den Verbindungen von den decodierten States
-
--->
+We fix 50 different random seeds (that is, for tuning, the data is split differently each of the 50 times), and for each we tune and train a model and then evaluate the performance of this model on the test set.
+We do this to ensure that our results are meaningful and not due to chance (especially in our small data set).
 
@@
 The plots above show the results for our 50 repetitions of the experiment.
-First, we show the classification accuracy, when using a threshold of 50%.
-Since this is an arbitrary value and our data is slightly imbalanced (~ 75% zone coverage plays and ~25% man coverage plays), accuracy is not the best metric for analyzing the results.
-Therefore we also present the AUC values of the model (`r colorize("We also provide further metrics in the Appendix","orange")`).
+First, we show the classification accuracy, when using a threshold of 50\%.
+Since this is an arbitrary value and our data is slightly imbalanced ($\approx$ 75\% zone coverage plays and $\approx$ 25\% man coverage plays), accuracy is not the best metric for analyzing the results.
+Therefore, we also present the AUC values of the model (`r colorize("We also provide further metrics in the Appendix","orange")`).
 For both metrics, we notice a few thing.
 On the one hand, the more flexible `xgboost` model performs better than the `glmnet`.
-However, there is more variation (larger boxes) for the `xgboost` model, suggesting that more data would help improve the tuning of the models.
+However, there is more variation for the `xgboost` model, suggesting that more data would help improve the tuning of the models.
 On the other hand, the pre-motion model clearly performs worst.
 As soon as we incorporate motion information, there is a large increase in performance.
-While this is expected, we also see that the enhanced model outperforms the model with only naive post-motion information.
-This shows, that the HMM features capture information, which is distinct from simple motion features and provide another dimension of information, which needs to be incorporated.
+While this is expected, we also see that the HMM post-motion model outperforms the model with only naive post-motion information.
+This shows that the HMM features capture information, which is distinct from simple motion features and provide another dimension of information, which needs to be incorporated.
 
@@
 
-While we were clearly able to show that motion facilitates the defensive coverage scheme detection, a limitation of our approach lies in the imperfect prediction accuracy of the pre-motion model, primarily due to insufficient hyperparameter tuning and the relatively small number of plays involving motion.
+While we were clearly able to show that pre-snap motion facilitates the defensive coverage scheme detection, a limitation of our approach lies in the imperfect prediction accuracy of the pre-motion model, primarily due to insufficient hyperparameter tuning and the relatively small number of plays involving motion.
 However, the primary focus of this project was on pre-snap motion, particularly on how to effectively translate this information into the hidden Markov model.
-`r colorize("Noch mehr Fokus auf Modelle und der Möglichkeit flexibler Modelle, wie die von AMAZON zu fitten, wenn mehr Daten dabei sind.","orange")`).
-Importantly, our pre-motion model is modular and can be seamlessly replaced by another model, such as the NFL Next Gen Stats model, which can then be integrated into our modeling extension to fully leverage the insights provided by our extension.
-`r colorize("Vergleiche unser Modell mit Results von Amazon: Unser Modell wesentlich weniger aufwändig, performt aber gut (Accuracy > 88 %.
-Accuracy von deren Modellen zwischen 88 und 96% obwohl sie 3 ganze Seasons an Daten haben (verglichen zu unseren 9 Wochen, wo wir 2/3 der plays entfernen müssen!)","orange")`).
+Importantly, our HMM approach is modular and can be seamlessly replaced by another model, such as the NFL Next Gen Stats model, in the case of a richer data set.
+Combining both of these worlds could fully leverage the insights provided by the present data.
 
@@
 
-A hidden Markov model consists of an observed time series $\{\boldsymbol{y}_t\}_{t=1}^T$ and an unobserved first-order Markov chain $\{ g_t\}_{t=1}^T$, with $g_t \in \{1,\ldots,N\}$.
-In this case, at every time point $t$, $\boldsymbol{y}_t$ is the y-coordinate of the defensive player and $g_t$ proxies the offensives player to be guarded, i.e.
+A hidden Markov model consists of an observed time series $\{y_t\}_{t=1}^T$ and an unobserved first-order Markov chain $\{ g_t\}_{t=1}^T$, with $g_t \in \{1,\ldots,N\}$.
+In this case, at every time point $t$, $y_t$ is the y-coordinate of the defensive player and $g_t$ proxies the offensives player to be guarded, i.e.
 the guarding allocation.
 The Markov chain is fully described by an initial distribution $\boldsymbol{\delta}^{(1)} = \bigl( \Pr(g_1=1), \ldots, \Pr(g_1=N) \bigr)$ and a transition probability matrix (t.p.m.) $\boldsymbol{\Gamma} = (\gamma_{ij})$, with $\gamma_{ij} = \Pr(g_t = j \mid g_{t-1} = i), \ i,j = 1, \ldots, N$.
-The connection of both stochastic processes arises from the assumption that the distribution of the observations $\boldsymbol{y}_t$ are fully determined by the state that is currently active.
+The connection of both stochastic processes arises from the assumption that the distribution of the observations $y_t$ are fully determined by the state that is currently active.
 More formally,
 \begin{equation*}
-f(\boldsymbol{y}_t \mid g_1, \ldots, g_T, \boldsymbol{y}_1, \ldots, \boldsymbol{y}_{t-1},\boldsymbol{y}_{t+1},\ldots,\boldsymbol{y}_T) = f(\boldsymbol{y}_t \mid g_t = j), \qquad j \in \{1, \ldots, N\},
+f(y_t \mid g_1, \ldots, g_T, y_1, \ldots, y_{t-1},y_{t+1},\ldots,y_T) = f(y_t \mid g_t = j), \qquad j \in \{1, \ldots, N\},
 \end{equation*}
-which we denote by $f_j(\boldsymbol{y}_t)$ in short.
+which we denote by $f_j(y_t)$ in short.
 In general, $f_j$ can be any density or probability mass function depending on the type of data and a typical choice is a parametric distribution with separate parameters for each latent state.
@@
 $$
-\Pr(g_t = j \mid \boldsymbol{y}_1, \dots, \boldsymbol{y}_T)
+\Pr(g_t = j \mid y_1, \dots, y_T)
 $$

### 2025-01-06 submission.Rmd
Sub sub sub

@@
 ## Data
-We analyze tracking data from the first nine weeks of the NFL 2022 season, provided by the NFL Big Data Bowl 2024.
+
+We aim to forecast the defensive scheme (man- or zone defense) using the `pff_passCoverage` indicator in play-by play data.
+We omit plays tagged as \textit{others} as well as plays with more than five offensive linemen and with two quarterbacks.
+Since we are specifically interested in analyzing pre-snap player movements, we concentrate on plays that contain any pre-snap motion.
+Then, we end up with $3985$ offensive plays in total, from which the defense played $2973$ in zone and $1012$ in man coverage.
+
+<!--We analyze tracking data from the first nine weeks of the NFL 2022 season, provided by the NFL Big Data Bowl 2024.
 Moreover, we use play-by-play data and the corresponding data from PFF that assigned the categories \textit{man}, \textit{zone} and \textit{others} representing the different schemes to each play.
@@
 Then, we end up with $3985$ offensive plays in total, from which the defense played $2973$ in zone and $1012$ in man coverage.
+-->
 
@@
 
-To accurately forecast the defensive scheme (man- or zone defense) for every play, we need to create various features derived from the tracking data.
-In particular, we conducted the following feature engineering steps: We first consider all 11 players on each side of the field and compute features related to the convex hull of the positions of the players.
+To accurately forecast the defensive scheme (man- or zone defense) for every play, we create various features derived from the tracking data.
+In particular, we conducted the following feature engineering steps: First, using all 11 players on each side, we compute the area spanned by the convex hull of a team as well as the largest $y$ distance (i.e.\ the width of the hull) and the largest $x$ distance (i.e.
+the length of the hull).
+Then, we select the five most relevant players on each side of the field.
+For offense, we omit the offensive line and the QB, thus coming up with five players that are a composition of running backs, tight ends and wide receivers.
+In contrast, for defense, disregarding defense liners (NT, DT, DE), we select the five defenders that were the closest to the five offensive players, corresponding to a weighted euclidean distance, putting much more emphasis on the y-axis.
+or these 10 players, we extract features related to their (standardized) position, distance and orientation.
+
+<!--
+For all 11 players on each side of the field we compute features related to the convex hull of the positions of the players.
 In particular, for defense and offense, we compute the area spanned by the convex hull of all player such as well as the largest $y$ distance (i.e.\ the width of the hull) and the largest $x$ distance (i.e.
@@
 In contrast, for defense, we omit nose tackles, defensive tackles and defensive ends, and select the five defenders that were the closest to the five offensive players, corresponding to a weighted euclidean distance, putting much more emphasis on the y-axis.
+For these 10 players, we extract features related to their (standardized) position, distance and orientation.
+<!--
 Finally, we use their standardized $x$ and $y$ coordinates as features and order defensive and offensive players according to their $y$ coordinates.
 Additionally, for each of the relevant defenders, we compute distances to the football and their orientation with respect to the quarterback.
+-->
 Finally, we extract relevant information from play-by-play data, such as quarter, down, yards to go, home and away score and the remaining game time in the current half (in seconds).
 
-Detailed information can be found in the Appendix.
+More detailed information can be found in the Appendix.
 
@@
 
+In general, we have a limited dataset available (only 3985 plays) and therefore need to balance the complexity of the model, thus aiming to control the number of features.
+Given the small dataset, we focus on 32 previously described basic features used in all 3 models: 6 convex hull related features, 20 player features, and 6 play-by-play features.
+In the Appendix, we provide a discussion on the choice of features.
+
+<!--
 `r colorize("der erste Teil ist doch eigentlich eine Wiederholung des Feature Engeneerings?","red")`
 In general, we have a limited dataset available (only 3985 plays) and therefore need to balance the complexity of the model, thus aiming to control the number of features.
-Given the small dataset, we focus on 32 previously described basic features: 6 convex hull related features (3 for offense and defense, respectively), 20 player positions features (10 standardized $x$ and $y$ coordinates, 5 for offense, 5 for defense), and 6 play-by-play features.
-In the Appendix, we provide more details and analyses using more features.
+Given the small dataset, we focus on 32 previously described basic features used in all 3 models: 6 convex hull related features (3 for offense and defense, respectively), 20 player positions features (10 standardized $x$ and $y$ coordinates, 5 for offense, 5 for defense), and 6 play-by-play features.
+In the Appendix, we provide a discussion on the choice of features.
+<!--we provide more details and analyses using more features.
+-->
 
@@
 For all of the models, we use 10-fold cross validation on a suitable hyperparameter grid.
+`r colorize("Ich finde der Teil ist besser in der Einleitung zur Analyse, da wir die zwei Modellklassen ja für alle 3 Modelle verwenden.
+Daher auch nochmal die Beschreibung der Features oben, weil die 32 features sind einfach für alle Modell gleich.","red")`
 
@@
 In a second step, we extend our basic pre-motion model with naive post-motion features.
-To keep the complexity manageable, we derive only six additional post-motion features: for each team (offense and defense), we infer the maximum $y$-distance, the maximum $x$-distance and the total distance traveled by both teams until the snap.
+To keep the complexity manageable, we derive only 6 additional post-motion features: for each team (offense and defense), we infer the maximum $y$-distance, the maximum $x$-distance and the total distance traveled by both teams until the snap.
 
@@
 We train our the models on 85% of the data and evaluate performance on the remaining 15%.
-Since we use cross-validation for tuning of the models during the training phase, and this relies on randomly dividing the data into folds, we perform our evaluation repeatedly.
-We fix 50 different random seeds (that is, for tuning, the data is split differently each of the 50 times), and for each we tune and train a model and then evaluate the performance of this model on the test set.
+Since we use cross-validation for tuning of the models during the training phase, we perform our evaluation repeatedly.
+We use 50 repetition each time splitting the data differently, and for each we tune and train a model and then evaluate the performance on the test set.
 We do this to ensure that our results are meaningful and not due to chance (especially in our small data set).
@@
 Since this is an arbitrary value and our data is slightly imbalanced ($\approx$ 75\% zone coverage plays and $\approx$ 25\% man coverage plays), accuracy is not the best metric for analyzing the results.
-Therefore, we also present the AUC values of the model (`r colorize("We also provide further metrics in the Appendix","orange")`).
-For both metrics, we notice a few thing.
+Therefore, we also present the AUC values of the model (see Appendix for a discussion on evaluation metrics).
+For both metrics, we notice a few things.
 On the one hand, the more flexible `xgboost` model performs better than the `glmnet`.
@@
 
-<!--
-robustheit checken mit simplen summarys nach der motion
+![](differences_teams.png)
 
-tests mit conditional independence (vllt in Anhang)
+Main findings:
 
-team analysen
--->
+Plot: In general, motion improves the correct detection of coverage (median of most teams positive).
+Table: Giants (Titans) --> low number of motions but very effective 49 out of 60 (82 %) of the motion plays increased coverage detections.
+Miami, Atlanta, SF --> Extensive use of motion but still very effective.
+Raider, Seattle, Rams --> Substantial use of motion but very ineffective.
+Bills (Bengals) -->  little use of motion and also ineffective.
+`r colorize("Rouven, wenn Du es schaffst einen netten text zu den Findings zu verfassen wäre es super!","red")`
+
+```{r team_tab, echo = FALSE, warning=FALSE, message=FALSE}
+
+library(tidyverse)
+library(reactable)
+library(reactablefmtr)
+library(scales)
+
+team_info <- readRDS("team_tab.rds")
+
+make_pal <- function(value,col_low = "red",col_middle = "skyblue",col_high = "darkgreen",nq = 10){
+  colourer <- col_quantile(
+    palette = c(col_low,col_middle,col_high),
+    domain = c(min(value), max(value)),
+    n = nq)
+  colourer(value)
+}
+
+team_pals <- team_info |> 
+  mutate(col1 = make_pal(n),
+         col2 = make_pal(n_improv),
+         col3 = make_pal(prop_improv))
+
+reactable(team_info |> select(url,n,n_improv,prop_improv),
+          bordered = TRUE,
+          defaultSorted = "prop_improv",
+          defaultSortOrder = "desc",
+
+          columns = list(
+            url = colDef(
+              name = "Team",
+              maxWidth = 70,
+              align = "center",
+              cell = embed_img(height = 25, width = 40)
+            ),
+            n = colDef(
+              name = "# Motions",
+              format = colFormat(digits = 0),
+              style = color_scales(team_pals, color_ref = "col1"),
+              maxWidth = 90
+            ),
+            n_improv = colDef(
+              name = "# Plays with improved probability of predicting correct defensive scheme in post motion model",
+              format = colFormat(digits = 0),
+              style = color_scales(team_pals, color_ref = "col2")
+            ),
+            prop_improv = colDef(
+              name = "Percentation of plays with improved probability of predicting correct defensive scheme in post motion model",
+              format = colFormat(digits = 2),
+              style = color_scales(team_pals, color_ref = "col3")
+            )
+          #   avg_per_tackle2 = colDef(
+          #     name = "Avg PEP[2]",
+          #     format = colFormat(digits = 2),
+          #     style = color_scales(best_tacklers_u, color_ref = "col4"),
+          #     maxWidth = 80
+          #   )
+          )
+
+)
+
+```
 
@@
 While we were clearly able to show that pre-snap motion facilitates the defensive coverage scheme detection, a limitation of our approach lies in the imperfect prediction accuracy of the pre-motion model, primarily due to insufficient hyperparameter tuning and the relatively small number of plays involving motion.
-However, the primary focus of this project was on pre-snap motion, particularly on how to effectively translate this information into the hidden Markov model.
+`r colorize("Ich würde nicht mit limitations anfangen sondern mit dem was wir gezeigt haben.
+Außerdem sollte man das insufficient hyperparameter tuning weglassen und nur auf den kleinen Datensatz fokussieren.
+Wir tunen gut genug und es liegt auch am Datensatz, dass wir nicht mehr machen.","red")` However, the primary focus of this project was on pre-snap motion, particularly on how to effectively translate this information into the hidden Markov model.
 Importantly, our HMM approach is modular and can be seamlessly replaced by another model, such as the NFL Next Gen Stats model, in the case of a richer data set.
@@
 
-#### Feature engineering and model features
+#### Feature engineering
+
+Prior to more involved feature engeneering steps, we transform the coordinate system by redefining the x-variable as the x-distance to the endzone (such that all play directions are from right to left and the relevant endzone is at zero), and changing the direction variable, such that zero degrees represents heading straight towards the corresponding endzone.
+As mentioned in the main text, we use player features from 5 defensive and 5 offensive player.
+First, we standardized their $x$ and $y$ coordinates with respect to the football and order the players according to their $y$ coordinates.
+That is the the first defender in our dataset is always the leftmost defensive player, while the first offensive player is always the rightmost one (offensive play direction is from left to right).
+Furthermore, for each player we compute distances to the football and their orientation with respect to the quarterback.
+
+#### Model features and comparison
 
 As frequently mentioned, we face the problem of having only a small amount of relevant data available.
-While we therefore focus on a small amount of features for our models for our main results, we provide more detailed results in this sections, using more thoroughly crafted features.
+In order to avoid overfitting problems, we therefore focus on a basic features set of 32 variables for our main results: 6 convex hull related features (3 for offense and defense, respectively), 20 player positions features (10 standardized $x$ and $y$ coordinates, 5 for offense, 5 for defense), and 6 play-by-play features.
+However, using more thoroughly crafted features such as distances and orientation as described above, we can enlarge the feature set to 67 total variables: 30 distance variables (for each of the 10 relevant players the total distance, $x$-distances, and $y$-distances to the football), and 5 orientation variables (for each defender the orientation with respect to the QB).
 
-`r colorize("Results go here...","orange")`
+In the following, we provide a comparison of these enlarged models and the smaller ones used for our main results.
+When comparing results of a binary outcome variable, accuracy can be a misleading metric, since it is dependent on the threshold selected for classification of the outcome.
+Usually, for probabilistic predictions as obtained by the two model classes considered, one uses the naive threshold of 0.5.
+However doing so is arbitrary and especially for imbalanced data, changing the threshold may change the results of the accuracy drastically.
+AUC (or area under the ROC curve) on the other hand, tries to evaluated the performance of a model over a suitable grid of thresholds, and is thus avoids the subjective choice of threshold.
+Another popularly used metric is the logloss (or negative log-likelihood loss), which is often a preferred choice due to being a proper scoring rule (mathematically speaking a scoring rule is proper, if it is optimal in expectation under the true probability).
+While using the logloss as evaluation metric is mathematically the best option, interpretation is not as intuitive.
+Specifically, whether a logloss of a model can be considered good, depends on the (im)balance of the classes (as well as the number of classes, which in a our binary case is only 2).
+We will therefore focus on these two metrics for the evaluation of the models.
 
-We see that using the more pronounced set of features seems to provide more information.
-At least the `glmnet` model, which is more simple but known to be able to handle a high number of features well, the version including a bigger set of features (orientations and distances) performs better than the any other model.
-For the `xgboost` model, the model including most features performs on par with the one including only small set of basic features.
-This suggests, that with more data the model can still be improved.
+![](AUCs_bdb_app.png)
+![](Loglosses_bdb_app.png)
+
+We see that using the more pronounced set of features (models contain "AF" in the name) seems to provide more information.
+At least the `glmnet` model, which is more simple but known to be able to handle a high number of features well, the versions including a bigger set of features (orientations and distances) performs better than the corresponding smaller models.
+For the `xgboost` model, the model including all features performs on par with the one including only small set of basic features.
+This suggests, that with more data the model may still be improved.
+
 Finally, we mention that we also varied the number of players used in the analyses.
 Presently, we use only information of 5 offensive and 5 defensive players.
-However using more information resulted in worse performance of the models, and thus we refrain from showing the results here.
-`r colorize("Review this section!","orange")`
+However, using more information resulted in worse performance of the models, thus we refrain from showing the results here.
 

### 2025-01-06 submission.Rmd
increasing caption size

@@
 ## Introduction
+
 Deciding between man and zone coverage is one of the most critical strategic decisions a defensive coordinator must take before each offensive play in American football.
@@
 
-#### 1. Pre-motion models
+### 1. Pre-motion models
 
@@
 
-#### 2. Naive post-motion models
+### 2. Naive post-motion models
 
@@
 
-#### 3. HMM post-motion models
+### 3. HMM post-motion models
 
-##### 3.1 Hidden Markov model
+#### 3.1 Hidden Markov model
 
@@
 
-##### 3.2 Enhanced post-motion model
+#### 3.2 Enhanced post-motion model
+
 We re-train the post-motion model to predict whether the defense plays a man- or zone coverage scheme, but, in this step, we incorporate results from the HMM analysis as further features.
@@
 
-#### Model and feature evaluation
+### Model and feature evaluation
 
@@
 
-#### Team analyses
+### Team analyses
 
@@
 
-#### Feature engineering
+### Feature engineering
 
@@
 
-#### Model features and comparison
+### Model features and comparison
 
@@
 
-#### Hidden Markov model
+### Hidden Markov model
 

### 2025-01-06 submission.Rmd
making animation smaller
updating submission file

@@
 
-Hence, our project goes one step further.
-While we similarly predict man- or zone coverage when the teams are set before snaps, we further leverage the additional information of pre-snap player movements in a specific manner.
-For this, we first include rather naive post-motion features.
-Afterwards, we use a hidden Markov model (HMM) to model defenders' trajectories based on hidden states, which represent the offensive players they may be guarding.
-Incorporating summary statistics of the state decoding results as features into the existing models significantly improves the predictive ability.
-This opens up the avenue for further analysis such as the evaluation of the effectiveness of pre-snap motion in uncovering defensive strategies.
+Hence, our contribution explores the potential of the information in pre-snap motion, previously omitted.
+While we similarly predict man- or zone coverage when the teams are set before snaps, we further leverage the additional information of pre-snap player movements.
+Specifically, in addition to including rather naive post-motion features, we use a hidden Markov model (HMM) to model defenders' trajectories based on hidden states, which represent the offensive players they may be guarding.
+Incorporating summary statistics of the state decoding results as features into the existing models substantially improves the predictive ability.
+This lays the groundwork for further analyses such as the evaluation of the effectiveness of pre-snap motion in uncovering defensive strategies.
 
@@
 Since we are specifically interested in analyzing pre-snap player movements, we concentrate on plays that contain any pre-snap motion.
-Then, we end up with $3985$ offensive plays in total, from which the defense played $2973$ in zone and $1012$ in man coverage.
+Ultimately, we end up with $3985$ offensive plays in total, from which the defense played $2973$ in zone and $1012$ in man coverage.
 
@@
 In contrast, for defense, disregarding defense liners (NT, DT, DE), we select the five defenders that were the closest to the five offensive players, corresponding to a weighted euclidean distance, putting much more emphasis on the y-axis.
-or these 10 players, we extract features related to their (standardized) position, distance and orientation.
-
+From these 10 players, we derive features related to their (standardized) position, distance and orientation.
 <!--
@@
 -->
-Finally, we extract relevant information from play-by-play data, such as quarter, down, yards to go, home and away score and the remaining game time in the current half (in seconds).
-
-More detailed information can be found in the Appendix.
+Additionally, we extract relevant information from the play-by-play data, such as quarter, down, yards to go, home and away score and the remaining game time in the current half (in seconds).
+See the Appendix for a more detailed description of the features.
 
@@
 
-In general, we have a limited dataset available (only 3985 plays) and therefore need to balance the complexity of the model, thus aiming to control the number of features.
+In general, we have a limited dataset available (only 3985 plays) and therefore need to manage model complexity by controlling the number of features.
 Given the small dataset, we focus on 32 previously described basic features used in all 3 models: 6 convex hull related features, 20 player features, and 6 play-by-play features.
@@
 
-The last step is as the heart of our approach.
-In particular, we use results from a hidden Markov model as further features in the previously described models.
-For this, we describe the movements of the five defensive players during the phase of pre-snap motion by an HMM (see the Appendix for an in-depth description).
-In particular, we assume that each defender's $y$-coordinate at each time point $t$ is a realization from a particular Gaussian distribution with mean according to the $y$-coordinate of the offensive player that is currently guarded and an estimated standard deviation.
-HMMs are particularly useful in this setting as there is no direct knowledge of the guarding allocation.
-This is only partially observed based on the reaction of the defender to the offenders' movement.
-Naturally, the model thus treats the information on who is currently guarded as a latent state variable, and hence each observation as a realization from a mixture of Gaussian distributions (see Franks et al.
+The previously described models primarily serve as baselines for evaluating the impact of incorporating pre-snap motion information into predicting defensive schemes.
+The main contribution of our project lies in the effective integration of this information using a hidden Markov model (HMM).
+
+Specifically, we use results derived from an HMM fitted to the pre-snap motion data as further features in the previously described models.
+To achieve this, we model the movements of the five defensive players during pre-snap motion by an HMM (see the Appendix for an in-depth description).
+In particular, we assume that each defender's $y$-coordinate at each time point $t$ is a realization from a Gaussian distribution with mean aligned to the $y$-coordinate of the offensive player they are guarding and an estimated standard deviation.
+HMMs are particularly well-suited for this task since the guarding assignments are not directly observed but need to be inferred from the defenders' responses to the offenders' movements.
+Hence, the model naturally treats this information as a latent state variable, making each observation a realization from a mixture of Gaussian distributions (see Franks et al.
 2015 for a similar approach in basketball).
-Ultimately, the goal is to use the fitted model for so-called *state-decoding*, i.e.
-inferring information on the guarding allocation based on the observations made.
-For this task, HMMs not only leverage information on the $y$-coordinates of the defender and all offenders potentially guarded at each time point, but also incorporate probabilistic information based on the previous and subsequent observations.
-For practical details on the model fitting and state decoding, also see the Appendix.
+The ultimate goal is to use the fitted model for *state-decoding*, i.e., inferring information on the guarding assignment based on the observations made.
+HMMs excel in this context as they leverage not only the defenders' $y$-coordiate as well as the $y$-coordiate of all offenders potentially guarded at the current time point, but also incorporate probabilistic information based on the previous and subsequent observations.
+State-decoding can either provide us with fully probabilistic information, i.e., the probability of each defender guarding each offensive player at each time point, or a single most likely guarding assignment for each defender at each time point.
+We will use the both to enhance our post-motion models and use the latter for visualization purposes.
+For details on model fitting and state decoding, also see the Appendix.
 
-The results of the HMM are exemplified using the following video.
-It displays a touchdown from the Kansas City Chiefs against the Arizona Cardinals in Week 1 of the 2022 NFL season.
+The results of the HMM are exemplified using the following video and animation.
+They display a touchdown of the Kansas City Chiefs against the Arizona Cardinals in Week 1 of the 2022 NFL season.
 We can see that, pre-snap, Mecole Hardman (KC #17) is in motion.
@@
 
-When performing state decoding, it becomes evident that the HMM effectively captures Marco Wilson's guarding assignment.
-The unique strength of HMMs is particularly apparent when Marco Wilson reaches the same height as the running back, Jerick McKinnon (KC #1).
+After state decoding using the fitted HMM, we can visualize the inferred guarding assignment as depicted below.
+It becomes evident that the HMM effectively captures Marco Wilson's guarding assignment.
+The unique strength of HMMs is apparent when Marco Wilson reaches the same $y$-coordinate as the running back, Jerick McKinnon (KC #1).
 At this point, clustering algorithms that disregard the temporal component would briefly suggest a change in the guarding assignment, introducing noise into summary statistics such as the total number of switches predicted by the model.
-In contrast, the HMM consistently and accurately maintains the correct coverage assignment throughout the motion.
+In contrast, the HMM consistently and accurately maintains the correct coverage assignment throughout the motion as a consequence of its temporal persistence.
 
 <center>
-  <img src="./animation/animation.gif" width="470" height="500" alt="Coverage Prediction according to a hidden Markov model.">
+  <img src="./animation/animation.gif" width="423" height="450" alt="Coverage Prediction according to a hidden Markov model.">
 </center>
@@
 
-We re-train the post-motion model to predict whether the defense plays a man- or zone coverage scheme, but, in this step, we incorporate results from the HMM analysis as further features.
-However, a caveat of the aforemententioned state probabilities is that they constitute as high-dimensional time-series data which makes them hard to use as features in the previously described models.
-An outlet of this problem is to employ suitable summary statistics.
-To overcome this, we first calculate for each defender the most likely offensive player $k = 1, \ldots, 5$ to be guarded in each time point $t = 1, \ldots, T_n$, where $T_n$ is the length of the $n$-th play, and count the number of state switches, i.e.
-the number of times for which a defender's most likely guarded offensive players differs between two consecutive time points $s$ and $t$, $s>t$.
-From this, we can calculate some simple, summary statistics: 1) the sum of state switches, 2) the average number of state switches and 3) the number of defenders that switch offensive players during a play.
-Fourth, we use the aforementioned state decodings to calculate a more elaborated statistic, the mean entropy across defenders $j = 1, \ldots, 5$ in each offensive play $n = 1,\ldots, N$, i.e.:
+To incorporate pre-snap information, we re-train the post-motion model to predict whether the defense employs a man- or zone-coverage scheme, now integrating results from the HMM analysis as additional features.
+A caveat of the aforementioned decoded state probabilities is that they form a multi-dimensional time series, complicating their direct use as features in the previously described models.
+This challenge can however be remedied by employing suitable summary statistics.
+To achieve this, we first calculate for each defender the most likely offensive player $k = 1, \ldots, 5$ to be guarded in each time point $t = 1, \ldots, T_n$, where $T_n$ is the length of the $n$-th play, and count the number of state switches, i.e.
+the number of times for which a defender's most likely guarded offensive players differs between two consecutive time points $t$ and $t+1$.
+From this, we can calculate some simple summary statistics, specifically 1) the sum of state switches, 2) the average number of state switches, and 3) the number of defenders that switch offensive players during a play.
+Additionally, we use the aforementioned decoded state probabilities to calculate a more elaborated statistic, the mean entropy across defenders $j = 1, \ldots, 5$ in each offensive play $n = 1,\ldots, N$:
 
-$$H(n) = - \frac{1}{5}\sum_{j = 1}^5 \sum_{k=1}^{5} \left( \frac{1}{T_n} \sum_{t=1}^{T_n} \mathbb{1}\left(\arg\max_{i=1,\ldots,5} X_{t,i} = k\right) \cdot \log\left(\frac{1}{T_n} \sum_{t=1}^{T_n} \mathbb{1}\left(\arg\max_{i=1,\ldots,5} X_{t,i} = k\right)\right) \right),$$
+$$H(n) = - \frac{1}{5}\sum_{j = 1}^5 \sum_{k=1}^{5} \left( \frac{1}{T_n} \sum_{t=1}^{T_n} \mathbb{1}\left(\arg\max_{i=1,\ldots,5} X_{t,i} = k\right) \cdot \log\left(\frac{1}{T_n} \sum_{t=1}^{T_n} \mathbb{1}\left(\arg\max_{i=1,\ldots,5} X_{t,i} = k\right)\right) \right)$$
 
-as the entropy is a measure of uncertainty or randomness in a probability distribution. Higher entropy indicates greater unpredictability, while lower entropy signifies more predictability.
+As the entropy is a measure of uncertainty or randomness in a probability distribution, higher entropy indicates greater unpredictability, while lower entropy signifies more predictability.
+In our setting, we suspect that higher entropy values are associated with less persistent guarding allocations, while lower entropy values are indicative of more stable guarding assignments.
 
@@
 
+`r colorize("Die plots sind noch sehr verpixelt","red")`
+
+<div style="text-align: center;">
   ![](ACCs_bdb_main.png)
   ![](AUCs_bdb_main.png)
+</div>
+
+<!-- <div style="text-align: center;"> -->
+<!--   ![](AUCs_bdb_main.png) -->
+<!-- </div> -->
+
+<!-- ![](ACCs_bdb_main.png) -->
+<!-- ![](AUCs_bdb_main.png) -->
 

### 2025-01-06 submission.Rmd
.

@@
 
+`r colorize("Den Text finde ich irgendwie unrund und nicht so leicht verständlich.","red")`
+
 We compare the predictive performance of our three models to determine the effectiveness of player movements before snap for detecting the correct defensive scheme.
@@
 
-<!-- <div style="text-align: center;"> -->
-<!--   ![](AUCs_bdb_main.png) -->
-<!-- </div> -->
-
 <!-- ![](ACCs_bdb_main.png) -->
@@
 
-The plots above show the results for our 50 repetitions of the experiment.
-First, we show the classification accuracy, when using a threshold of 50\%.
-Since this is an arbitrary value and our data is slightly imbalanced ($\approx$ 75\% zone coverage plays and $\approx$ 25\% man coverage plays), accuracy is not the best metric for analyzing the results.
-Therefore, we also present the AUC values of the model (see Appendix for a discussion on evaluation metrics).
-For both metrics, we notice a few things.
-On the one hand, the more flexible `xgboost` model performs better than the `glmnet`.
-However, there is more variation for the `xgboost` model, suggesting that more data would help improve the tuning of the models.
-On the other hand, the pre-motion model clearly performs worst.
-As soon as we incorporate motion information, there is a large increase in performance.
-While this is expected, we also see that the HMM post-motion model outperforms the model with only naive post-motion information.
-This shows that the HMM features capture information, which is distinct from simple motion features and provide another dimension of information, which needs to be incorporated.
+The plots above show the results from our 50 repetitions of the experiment.
+The first figure illustrates classification accuracy, using a threshold of 50\%.
+Since this value is somewhat arbitrary and our dataset is slightly imbalanced ($\approx$ 75\% zone coverage plays and $\approx$ 25\% man coverage plays), accuracy is a suboptimal metric for evaluating the results.
+Hence, we also present the AUC values of the three models (see Appendix for a discussion on evaluation metrics).
+Several observations emerge from these metrics.
+On the one hand, the more flexible `xgboost` model outperforms `glmnet`, but it also exhibits greater variation, indicating that additional data could enhance the tuning of this model.
+On the other hand, the pre-motion model performs the worst. However, once motion information is incorporated, we observe a substantial improvement in performance.
+While this is to be expected, we also see that the HMM post-motion model outperforms the model with only naive post-motion information.
+This indicates that the HMM features capture distinct information beyond simple motion features, providing an additional dimension that enhances the overall analysis.
 
 ### Team analyses
+
+`r colorize("Die figure fehlt irgendwie noch","red")`
 
@@
 Combining both of these worlds could fully leverage the insights provided by the present data.
+
+`r colorize("Von chatGPT:","red")`
+
+This project explored the use of hidden Markov models (HMMs) to enhance the prediction of defensive schemes in football by incorporating pre-snap player movements.
+Models augmented with HMM-derived features demonstrated improved predictive performance compared to those relying solely on naive motion data.
+The HMM effectively captured latent defensive behaviors, offering deeper insights into the defenders' actions.
+
+The HMM features allowed for more effective inference of guarding assignments, improving the distinction between man and zone coverage.
+This improvement underscores the HMM's capacity to model the dynamic and often subtle nature of defensive strategies, capturing patterns that simple motion features cannot.
+
+Looking ahead, the potential of HMMs goes beyond predicting man or zone defense.
+They could be applied to decode complex guarding assignments and analyze defensive adjustments in real-time, providing a powerful tool for deeper strategic insights in sports analytics.
 

### 2025-01-06 submission.Rmd
.

@@
 Prior to more involved feature engeneering steps, we transform the coordinate system by redefining the x-variable as the x-distance to the endzone (such that all play directions are from right to left and the relevant endzone is at zero), and changing the direction variable, such that zero degrees represents heading straight towards the corresponding endzone.
-As mentioned in the main text, we use player features from 5 defensive and 5 offensive player.
-First, we standardized their $x$ and $y$ coordinates with respect to the football and order the players according to their $y$ coordinates.
-That is the the first defender in our dataset is always the leftmost defensive player, while the first offensive player is always the rightmost one (offensive play direction is from left to right).
+As mentioned in the main text, we use player features from 5 defensive and 5 offensive players.
+First, we standardized their $x$- and $y$-coordinates with respect to the football and ordered the players according to their $y$-coordinates, i.e., the the first defender in our dataset is always the leftmost defensive player, while the first offensive player is always the rightmost one (offensive play direction is from left to right).
 Furthermore, for each player we compute distances to the football and their orientation with respect to the quarterback.
@@
 
-As frequently mentioned, we face the problem of having only a small amount of relevant data available.
-In order to avoid overfitting problems, we therefore focus on a basic features set of 32 variables for our main results: 6 convex hull related features (3 for offense and defense, respectively), 20 player positions features (10 standardized $x$ and $y$ coordinates, 5 for offense, 5 for defense), and 6 play-by-play features.
+As frequently mentioned, we face the problem of only having a small amount of relevant data available.
+In order to avoid overfitting problems, we therefore focus on a basic feature set of 32 variables for our main results: 6 convex hull related features (3 for offense and defense, respectively), 20 player positions features (10 standardized $x$ and $y$ coordinates, 5 for offense, 5 for defense), and 6 play-by-play features.
 However, using more thoroughly crafted features such as distances and orientation as described above, we can enlarge the feature set to 67 total variables: 30 distance variables (for each of the 10 relevant players the total distance, $x$-distances, and $y$-distances to the football), and 5 orientation variables (for each defender the orientation with respect to the QB).
@@
 In this case, at every time point $t$, $y_t$ is the y-coordinate of the defensive player and $g_t$ proxies the offensives player to be guarded, i.e.
-the guarding allocation.
+the guarding assignment.
 The Markov chain is fully described by an initial distribution $\boldsymbol{\delta}^{(1)} = \bigl( \Pr(g_1=1), \ldots, \Pr(g_1=N) \bigr)$ and a transition probability matrix (t.p.m.) $\boldsymbol{\Gamma} = (\gamma_{ij})$, with $\gamma_{ij} = \Pr(g_t = j \mid g_{t-1} = i), \ i,j = 1, \ldots, N$.
@@
 It effectively performs a summation over all possible latent state sequences in an efficient manner, rendering the computational complexity linear in the number of observations.
-Time series of different defenders within the same play, and of different plays are treated as independent, hence their likelihood contributions are summed to obtain the full likelihood of the training data.
+Time series of different defenders within the same play and of different plays are treated as independent, hence their likelihood contributions are summed to obtain the full likelihood of the training data.
 For practical implementation, we wrote a custom likelihood function in R, using the function `forward()` and other convenience functions from the R package `LaMa` (Koslik, 2024) to speed up computations.
@@
 The parameters to be estimated are only the transition probability matrix $\boldsymbol{\Gamma}$ and the standard deviation of the Gaussian distribution. 
-The initial distribution of each defender for each play comes from a deterministic assignment approach based on spatial proximity of the initial y-coordinates at the beginning of the time-series, i.e.
-the moment the line is set.
-Specifically, for each defender in a predefined set of potential defenders, we calculated their Euclidean distances to the five offensive players For each offense player, ranked in descending order by proximity, we assigned the closest defender to that player.
-The initial distribution was then set as a categorical distribution, where each defender was assigned a probability of 1 for their corresponding attacker and 0 for all others.
+The initial distribution of the guarding assignment for each defender in each play arises from a deterministic assignment approach, based on spatial proximity at the beginning of the time-series, i.e., the moment the line is set.
+Specifically, for each defender in the predefined set of defenders considered, we compute a weighted Euclidean distance (giving more weight to the $y$-coordinate) to the five offensive players that could potentially be guarded.
+The initial distribution is then set to 1 for the closest defender under this metric and 0 otherwise.
+While this approach proved to be sufficiently reliable in this application, it may be finetuned in future iterations.
 
 Having fitted an HMM to the data, we can use the model to predict the underlying state sequence based on the observations.
-This process is called state decoding and two main approaches exist.
+This process is called *state decoding* and two main approaches exist.
 So-called *local decoding* constructs the conditional distributions
@@
 while *global decoding* using the *Viterbi-algorithm* finds the state sequence that maximizes the joint probability of the state sequence given the observations.
-Local decoding retains more probabilistic information as it provides a categorical state distribution for each time point while global decoding is more suitable for visualization purposes as it provides a single state sequence that is most likely to have generated the observations.
+Local decoding retains more probabilistic information as it provides a categorical state distribution for each time point, while global decoding is more suitable for visualization purposes as it provides a single state sequence that is most likely to have generated the observations.
 To obtain both the local state probabilities and the global state sequence, we used the functions `stateprobs()` and `viterbi()` that are also contained in the R package `LaMa`.

### 2025-01-06 submission.Rmd
.

@@
 We aim to forecast the defensive scheme (man- or zone defense) using the `pff_passCoverage` indicator in play-by play data.
-We omit plays tagged as \textit{others} as well as plays with more than five offensive linemen and with two quarterbacks.
+We omit plays tagged as *others* as well as plays with more than five offensive linemen and with two quarterbacks.
 Since we are specifically interested in analyzing pre-snap player movements, we concentrate on plays that contain any pre-snap motion.
@@
 We compare the predictive performance of our three models to determine the effectiveness of player movements before snap for detecting the correct defensive scheme.
-We train our the models on 85% of the data and evaluate performance on the remaining 15%.
+We train the models on 85% of the data and evaluate performance on the remaining 15%.
 Since we use cross-validation for tuning of the models during the training phase, we perform our evaluation repeatedly.
@@
 As frequently mentioned, we face the problem of only having a small amount of relevant data available.
-In order to avoid overfitting problems, we therefore focus on a basic feature set of 32 variables for our main results: 6 convex hull related features (3 for offense and defense, respectively), 20 player positions features (10 standardized $x$ and $y$ coordinates, 5 for offense, 5 for defense), and 6 play-by-play features.
+In order to avoid overfitting problems, we therefore focus on a basic feature set of 32 variables for our main results: 6 convex hull related features (3 for offense and defense, respectively), 20 player positions features (10 standardized $x$ and 10 standardized $y$ coordinates, 5 for offense, 5 for defense), and 6 play-by-play features.
 However, using more thoroughly crafted features such as distances and orientation as described above, we can enlarge the feature set to 67 total variables: 30 distance variables (for each of the 10 relevant players the total distance, $x$-distances, and $y$-distances to the football), and 5 orientation variables (for each defender the orientation with respect to the QB).
@@
 However doing so is arbitrary and especially for imbalanced data, changing the threshold may change the results of the accuracy drastically.
-AUC (or area under the ROC curve) on the other hand, tries to evaluated the performance of a model over a suitable grid of thresholds, and is thus avoids the subjective choice of threshold.
-Another popularly used metric is the logloss (or negative log-likelihood loss), which is often a preferred choice due to being a proper scoring rule (mathematically speaking a scoring rule is proper, if it is optimal in expectation under the true probability).
+AUC (or area under the ROC curve) on the other hand, tries to evaluated the performance of a model over a suitable grid of thresholds, and thus avoids the subjective choice of threshold.
+Another popularly used metric is the logloss (or negative log-likelihood loss), which is often a preferred choice due to being a proper scoring rule.
 While using the logloss as evaluation metric is mathematically the best option, interpretation is not as intuitive.
 Specifically, whether a logloss of a model can be considered good, depends on the (im)balance of the classes (as well as the number of classes, which in a our binary case is only 2).
-We will therefore focus on these two metrics for the evaluation of the models.
+Without going into further detail, we therefore focus on these two metrics for the evaluation of the models.
 

### 2025-01-06 submission.Rmd

@@
 
+<!--
 We compare the predictive performance of our three models to determine the effectiveness of player movements before snap for detecting the correct defensive scheme.
@@
 We do this to ensure that our results are meaningful and not due to chance (especially in our small data set).
+-->
+
+`r colorize("Vorschlag (Robert):","red")`
+
+We evaluate the predictive performance of our three models to assess how well pre-snap player movements help identify the correct defensive scheme.
+To do this, we split the data into two sets: 85% for training and 15% for testing.
+During the training phase, we use cross-validation to fine-tune the models.
+
+To ensure robust and reliable results, we repeat the evaluation process 50 times, using different cv splits for each iteration.
+For each split, we train and fine-tune the model on the training data and then evaluate its performance on the test set.
+This approach helps minimize the risk of our findings being influenced by randomness, which is particularly important given the limited size of our dataset.
 
@@
 
+`r colorize("Robert: komisch, bei mir sind sie eig recht scharf... hab versucht es bisschen zu ändern. Wenns das nicht besser macht, weiß ich leider auch nich was hilft...","red")`
+
 <div style="text-align: center;">
-  ![](ACCs_bdb_main.png)
-  ![](AUCs_bdb_main.png)
+  ![](ACCs_bdb_main2.png)
+  ![](AUCs_bdb_main2.png)
 </div>
@@
 
+`r colorize("Vorschlag Robert (mit Teilen von oben und chatGPT Hilfe von unten):","red")`
+
+This project explored the use of hidden Markov models (HMMs) to enhance the prediction of defensive schemes in football by incorporating pre-snap player movements.
+Models augmented with HMM-derived features demonstrated improved predictive performance compared to those relying solely on naive motion data.
+
+Comparing predictions from pre-motion and post-motion models provides insights into an offense's ability to use motion effectively to identify defensive coverage schemes.
+Similarly, it allows for evaluating a defense's skill in disguising their coverage and confusing the offense during pre-snap motion.
+
+While our pre- and post-motion models are limited due to small sample size available, our HMM approach is modular and therefore our models can be replaced by more another model in the case of a richer data set.
+Combining both of these worlds could fully leverage the insights provided by the present data.
+
 `r colorize("Von chatGPT:","red")`

### 2025-01-06 submission.Rmd
.

@@
 
-We evaluate the predictive performance of our three models to assess how well pre-snap player movements help identify the correct defensive scheme.
+To assess whether the incorporation of pre-snap player movements can improve the prediction of defensive schemes, we evaluate the predictive performance of our three models.
+Therefore, we perform a train-test split using 85% for training and 15% for testing.
+To ensure robust and reliable results, we repeat the evaluation process 50 times, using different CV splits for each iteration.
+For each split, we train and fine-tune the model on the training data and then evaluate its performance on the test set.
+This approach helps minimize the risk of our findings being influenced by randomness, which is particularly important given the limited size of our dataset.
+The results of the evaluation are shown in the figures below.
+
+<!-- We evaluate the predictive performance of our three models to assess how well pre-snap player movements help identify the correct defensive scheme.
 To do this, we split the data into two sets: 85% for training and 15% for testing.
 During the training phase, we use cross-validation to fine-tune the models.
+-->
 
-To ensure robust and reliable results, we repeat the evaluation process 50 times, using different cv splits for each iteration.
+<!-- To ensure robust and reliable results, we repeat the evaluation process 50 times, using different cv splits for each iteration.
 For each split, we train and fine-tune the model on the training data and then evaluate its performance on the test set.
 This approach helps minimize the risk of our findings being influenced by randomness, which is particularly important given the limited size of our dataset.
+-->
 
@@
 
-The plots above show the results from our 50 repetitions of the experiment.
-The first figure illustrates classification accuracy, using a threshold of 50\%.
+The two figures above show the results from our 50 repetitions of the experiment.
+The first figure illustrates the classification accuracy, computed using a threshold of 50\%.
 Since this value is somewhat arbitrary and our dataset is slightly imbalanced ($\approx$ 75\% zone coverage plays and $\approx$ 25\% man coverage plays), accuracy is a suboptimal metric for evaluating the results.
-Hence, we also present the AUC values of the three models (see Appendix for a discussion on evaluation metrics).
+Hence, we also present the AUC values of the three models (see Appendix for a discussion on evaluation metrics) in the second figure.
 Several observations emerge from these metrics.
 On the one hand, the more flexible `xgboost` model outperforms `glmnet`, but it also exhibits greater variation, indicating that additional data could enhance the tuning of this model.
-On the other hand, the pre-motion model performs the worst. However, once motion information is incorporated, we observe a substantial improvement in performance.
-While this is to be expected, we also see that the HMM post-motion model outperforms the model with only naive post-motion information.
-This indicates that the HMM features capture distinct information beyond simple motion features, providing an additional dimension that enhances the overall analysis.
+On the other hand, the pre-motion model clearly performs worst. Once motion information is incorporated, we observe a substantial improvement in performance.
+While this is to be expected, importantly, we also see that the HMM post-motion model outperforms the model only using naive post-motion information.
+This indicates that the features derived from the HMM's state decoding capture distinct information beyond simple motion features, providing an additional dimension that enhances the overall analysis.
 

### 2025-01-06 submission.Rmd
.

@@
 
-<!--We analyze tracking data from the first nine weeks of the NFL 2022 season, provided by the NFL Big Data Bowl 2024.
-Moreover, we use play-by-play data and the corresponding data from PFF that assigned the categories \textit{man}, \textit{zone} and \textit{others} representing the different schemes to each play.
-As it is not properly described what \textit{others} means, we omit every play that is associated with this value.
-We also omit plays with more than five offensive linemen and with two quarterbacks.
-Since we are specifically interested in analyzing pre-snap player movements, we concentrate on plays that contain any pre-snap motion.
-Then, we end up with $3985$ offensive plays in total, from which the defense played $2973$ in zone and $1012$ in man coverage.
--->
-
 ##### Feature engineering
 
-To accurately forecast the defensive scheme (man- or zone defense) for every play, we create various features derived from the tracking data.
+To accurately forecast the defensive scheme, we create various features derived from the tracking data.
 In particular, we conducted the following feature engineering steps: First, using all 11 players on each side, we compute the area spanned by the convex hull of a team as well as the largest $y$ distance (i.e.\ the width of the hull) and the largest $x$ distance (i.e.
@@
 From these 10 players, we derive features related to their (standardized) position, distance and orientation.
-<!--
-For all 11 players on each side of the field we compute features related to the convex hull of the positions of the players.
-In particular, for defense and offense, we compute the area spanned by the convex hull of all player such as well as the largest $y$ distance (i.e.\ the width of the hull) and the largest $x$ distance (i.e.
-the length of the hull).
-In addition, we select the five most relevant players on each side of the field.
-For offense, we omit the offensive line and the QB, thus coming up with five players that are a composition of running backs, tight ends and wide receivers.
-In contrast, for defense, we omit nose tackles, defensive tackles and defensive ends, and select the five defenders that were the closest to the five offensive players, corresponding to a weighted euclidean distance, putting much more emphasis on the y-axis.
-For these 10 players, we extract features related to their (standardized) position, distance and orientation.
-<!--
-Finally, we use their standardized $x$ and $y$ coordinates as features and order defensive and offensive players according to their $y$ coordinates.
-Additionally, for each of the relevant defenders, we compute distances to the football and their orientation with respect to the quarterback.
--->
+
 Additionally, we extract relevant information from the play-by-play data, such as quarter, down, yards to go, home and away score and the remaining game time in the current half (in seconds).
@@
 
-<!--
-`r colorize("der erste Teil ist doch eigentlich eine Wiederholung des Feature Engeneerings?","red")`
-In general, we have a limited dataset available (only 3985 plays) and therefore need to balance the complexity of the model, thus aiming to control the number of features.
-Given the small dataset, we focus on 32 previously described basic features used in all 3 models: 6 convex hull related features (3 for offense and defense, respectively), 20 player positions features (10 standardized $x$ and $y$ coordinates, 5 for offense, 5 for defense), and 6 play-by-play features.
-In the Appendix, we provide a discussion on the choice of features.
-<!--we provide more details and analyses using more features.
--->
-
-### 1. Pre-motion models
-
 First of all, we need a suitable basic model class for predicting man or zone coverage and opt for the following two: First, we fit a `glmnet` (elastic net) model, which performs implicit feature selection and is able to handle multicollinearity.
@@
 For all of the models, we use 10-fold cross validation on a suitable hyperparameter grid.
-`r colorize("Ich finde der Teil ist besser in der Einleitung zur Analyse, da wir die zwei Modellklassen ja für alle 3 Modelle verwenden.
-Daher auch nochmal die Beschreibung der Features oben, weil die 32 features sind einfach für alle Modell gleich.","red")`
 
-We fit the models with the previously described basic features.
-All of these features are derived at the time of line-set, which is why these models do not use any pre-snap motion information.
+### 1. Pre-motion models
+
+At first, we fit the aforementioned models with the previously described basic pre-motion features as all of these features are derived at the time of line-set.
 These very basic models serve as baseline models that allows to measure the effect of pre-snap motion (features) in the following.
@@
 In a second step, we extend our basic pre-motion model with naive post-motion features.
-To keep the complexity manageable, we derive only 6 additional post-motion features: for each team (offense and defense), we infer the maximum $y$-distance, the maximum $x$-distance and the total distance traveled by both teams until the snap.
+To keep the complexity manageable, we derive only six additional post-motion features: for each team (offense and defense), we infer the maximum $y$-distance, the maximum $x$-distance and the total distance traveled by both teams until the snap.
 
@@
 
-Specifically, we use results derived from an HMM fitted to the pre-snap motion data as further features in the previously described models.
+Specifically, we use results derived from an HMM fitted to the pre-snap motion tracking data as further features in the previously described models.
 To achieve this, we model the movements of the five defensive players during pre-snap motion by an HMM (see the Appendix for an in-depth description).
@@
 The ultimate goal is to use the fitted model for *state-decoding*, i.e., inferring information on the guarding assignment based on the observations made.
-HMMs excel in this context as they leverage not only the defenders' $y$-coordiate as well as the $y$-coordiate of all offenders potentially guarded at the current time point, but also incorporate probabilistic information based on the previous and subsequent observations.
+HMMs excel in this context as they leverage not only the defenders' $y$-coordiate and the $y$-coordiate of all offenders potentially guarded at the current time point, but also incorporate probabilistic information based on the previous and subsequent observations.
 State-decoding can either provide us with fully probabilistic information, i.e., the probability of each defender guarding each offensive player at each time point, or a single most likely guarding assignment for each defender at each time point.
-We will use the both to enhance our post-motion models and use the latter for visualization purposes.
-For details on model fitting and state decoding, also see the Appendix.
+We will use the former to enhance our post-motion models and the latter for visualization purposes.
+For details on model fitting and state decoding, we also refer to the Appendix.
 
@@
 
-To incorporate pre-snap information, we re-train the post-motion model to predict whether the defense employs a man- or zone-coverage scheme, now integrating results from the HMM analysis as additional features.
-A caveat of the aforementioned decoded state probabilities is that they form a multi-dimensional time series, complicating their direct use as features in the previously described models.
+To incorporate pre-snap motion information, we re-train the post-motion model to predict whether the defense employs a man- or zone-coverage scheme, now integrating results from the HMM analysis as additional features.
+A caveat of the aforementioned decoded state probabilities is that they form a multi-dimensional time series, complicating their direct inclusion as features in the previously described models.
 This challenge can however be remedied by employing suitable summary statistics.
@@
 <div style="text-align: center;">
-  ![](ACCs_bdb_main2.png)
-  ![](AUCs_bdb_main2.png)
+  ![](ACCs_bdb_main.png)
+  ![](AUCs_bdb_main.png)
 </div>
@@
 
-![](differences_teams.png)
+![](differences_teams.png){width=50%}
 
@@
             n_improv = colDef(
-              name = "# Plays with improved probability of predicting correct defensive scheme in post motion model",
+              name = "# Plays with improved probability of predicting correct defensive scheme in HMM post-motion model",
               format = colFormat(digits = 0),
@@
             prop_improv = colDef(
-              name = "Percentation of plays with improved probability of predicting correct defensive scheme in post motion model",
+              name = "Percentation of plays with improved probability of predicting correct defensive scheme in HMM post-motion model",
               format = colFormat(digits = 2),
@@
 
-While we were clearly able to show that pre-snap motion facilitates the defensive coverage scheme detection, a limitation of our approach lies in the imperfect prediction accuracy of the pre-motion model, primarily due to insufficient hyperparameter tuning and the relatively small number of plays involving motion.
-`r colorize("Ich würde nicht mit limitations anfangen sondern mit dem was wir gezeigt haben.
-Außerdem sollte man das insufficient hyperparameter tuning weglassen und nur auf den kleinen Datensatz fokussieren.
-Wir tunen gut genug und es liegt auch am Datensatz, dass wir nicht mehr machen.","red")` However, the primary focus of this project was on pre-snap motion, particularly on how to effectively translate this information into the hidden Markov model.
-Importantly, our HMM approach is modular and can be seamlessly replaced by another model, such as the NFL Next Gen Stats model, in the case of a richer data set.
-Combining both of these worlds could fully leverage the insights provided by the present data.
-
-`r colorize("Vorschlag Robert (mit Teilen von oben und chatGPT Hilfe von unten):","red")`
-
 This project explored the use of hidden Markov models (HMMs) to enhance the prediction of defensive schemes in football by incorporating pre-snap player movements.
@@
 Comparing predictions from pre-motion and post-motion models provides insights into an offense's ability to use motion effectively to identify defensive coverage schemes.
-Similarly, it allows for evaluating a defense's skill in disguising their coverage and confusing the offense during pre-snap motion.
+Similarly, future analysis could also relate to evaluating a defense's skill in disguising their coverage and confusing the offense during pre-snap motion.
 
-While our pre- and post-motion models are limited due to small sample size available, our HMM approach is modular and therefore our models can be replaced by more another model in the case of a richer data set.
+While our pre- and post-motion models are limited due to small sample size available, our HMM approach is modular and therefore our models can be seamlessly replaced by other models, potentially more complex ones in the case of a richer data set.
 Combining both of these worlds could fully leverage the insights provided by the present data.
-
-`r colorize("Von chatGPT:","red")`
-
-This project explored the use of hidden Markov models (HMMs) to enhance the prediction of defensive schemes in football by incorporating pre-snap player movements.
-Models augmented with HMM-derived features demonstrated improved predictive performance compared to those relying solely on naive motion data.
-The HMM effectively captured latent defensive behaviors, offering deeper insights into the defenders' actions.
-
-The HMM features allowed for more effective inference of guarding assignments, improving the distinction between man and zone coverage.
-This improvement underscores the HMM's capacity to model the dynamic and often subtle nature of defensive strategies, capturing patterns that simple motion features cannot.
-
-Looking ahead, the potential of HMMs goes beyond predicting man or zone defense.
-They could be applied to decode complex guarding assignments and analyze defensive adjustments in real-time, providing a powerful tool for deeper strategic insights in sports analytics.
 
@@
 As mentioned in the main text, we use player features from 5 defensive and 5 offensive players.
-First, we standardized their $x$- and $y$-coordinates with respect to the football and ordered the players according to their $y$-coordinates, i.e., the the first defender in our dataset is always the leftmost defensive player, while the first offensive player is always the rightmost one (offensive play direction is from left to right).
+First, we standardized their $x$- and $y$-coordinates with respect to the football and ordered the players according to their $y$-coordinates, i.e., the first defender in our dataset is always the leftmost defensive player, while the first offensive player is always the rightmost one (offensive play direction is from left to right).
 Furthermore, for each player we compute distances to the football and their orientation with respect to the quarterback.
@@
 As frequently mentioned, we face the problem of only having a small amount of relevant data available.
-In order to avoid overfitting problems, we therefore focus on a basic feature set of 32 variables for our main results: 6 convex hull related features (3 for offense and defense, respectively), 20 player positions features (10 standardized $x$ and 10 standardized $y$ coordinates, 5 for offense, 5 for defense), and 6 play-by-play features.
+To avoid overfitting problems, we therefore focus on a basic feature set of 32 variables for our main results: 6 convex hull related features (3 for offense and defense, respectively), 20 player positions features (10 standardized $x$ and 10 standardized $y$ coordinates, 5 for offense, 5 for defense), and 6 play-by-play features.
 However, using more thoroughly crafted features such as distances and orientation as described above, we can enlarge the feature set to 67 total variables: 30 distance variables (for each of the 10 relevant players the total distance, $x$-distances, and $y$-distances to the football), and 5 orientation variables (for each defender the orientation with respect to the QB).
@@
 When comparing results of a binary outcome variable, accuracy can be a misleading metric, since it is dependent on the threshold selected for classification of the outcome.
-Usually, for probabilistic predictions as obtained by the two model classes considered, one uses the naive threshold of 0.5.
+Usually, for probabilistic predictions as obtained by the two model classes considered, one uses the naive threshold of $0.5$.
 However doing so is arbitrary and especially for imbalanced data, changing the threshold may change the results of the accuracy drastically.

### 2025-01-06 submission_61.Rmd
Final?

--- 2025-01-06 submission.Rmd
@@
 Deciding between man and zone coverage is one of the most critical strategic decisions a defensive coordinator must take before each offensive play in American football.
-While experienced offensive coordinators and quarterbacks often rely on visual cues to identify these defensive schemes, the increasing availability of player tracking data offers a new avenue to uncover and analyze these tactics.
-A notable example is Amazon's NFL Next Gen Stats model, which delivers coverage predictions during live broadcasts (see a snapshot of the 2024 Week 12 match between the Pittsburgh Steelers and Cleveland Browns).
-However, in these pre-snap motion does not seem to play an accentuated role (see [Amazon](https://aws.amazon.com/de/blogs/machine-learning/identifying-defense-coverage-schemes-in-nfls-next-gen-stats/)), although it is a crucial element of modern offensive strategies and it is often used to detect the hidden defensive scheme.
+While experienced offensive coordinators and quarterbacks often rely on visual cues to identify these defensive schemes, the increasing availability of player tracking data offers a new avenue to uncover these tactics.
+A notable example is Amazon's NFL Next Gen Stats model, which delivers coverage predictions during live broadcasts (see snapshot below).
+However, pre-snap motion does not seem to play an accentuated role in this model (see [Amazon](https://aws.amazon.com/de/blogs/machine-learning/identifying-defense-coverage-schemes-in-nfls-next-gen-stats/)), although it is a crucial element of modern offensive strategies.
 
-Hence, our contribution explores the potential of the information in pre-snap motion, previously omitted.
-While we similarly predict man- or zone coverage when the teams are set before snaps, we further leverage the additional information of pre-snap player movements.
-Specifically, in addition to including rather naive post-motion features, we use a hidden Markov model (HMM) to model defenders' trajectories based on hidden states, which represent the offensive players they may be guarding.
+Hence, our contribution explores the potential of incorporating pre-snap motion.
+While we similarly predict man- or zone coverage before motion, we further leverage the additional information of pre-snap player movements.
+Specifically, in addition to including rather naive post-motion features, we use an HMM to model defenders' trajectories based on hidden states, which represent the offensive players they may be guarding.
 Incorporating summary statistics of the state decoding results as features into the existing models substantially improves the predictive ability.
@@
 
-We aim to forecast the defensive scheme (man- or zone defense) using the `pff_passCoverage` indicator in play-by play data.
-We omit plays tagged as *others* as well as plays with more than five offensive linemen and with two quarterbacks.
+We aim to forecast the man- or zone coverage using the `pff_passCoverage` indicator in play-by play data.
+We omit plays tagged as *others* and plays with more than five offensive linemen and with two quarterbacks.
 Since we are specifically interested in analyzing pre-snap player movements, we concentrate on plays that contain any pre-snap motion.
@@
 To accurately forecast the defensive scheme, we create various features derived from the tracking data.
-In particular, we conducted the following feature engineering steps: First, using all 11 players on each side, we compute the area spanned by the convex hull of a team as well as the largest $y$ distance (i.e.\ the width of the hull) and the largest $x$ distance (i.e.
-the length of the hull).
+In particular, we conducted the following feature engineering steps: First, using all 11 players on each side, we compute the area spanned by the convex hull of a team and the largest $y$ distance (i.e., the width of the hull) and the largest $x$ distance (i.e., the length of the hull), i.e.
+6 features.
 Then, we select the five most relevant players on each side of the field.
-For offense, we omit the offensive line and the QB, thus coming up with five players that are a composition of running backs, tight ends and wide receivers.
-In contrast, for defense, disregarding defense liners (NT, DT, DE), we select the five defenders that were the closest to the five offensive players, corresponding to a weighted euclidean distance, putting much more emphasis on the y-axis.
-From these 10 players, we derive features related to their (standardized) position, distance and orientation.
+For offense, we omit the offensive line and the QB and for defense, we disregard defense lineman (NT, DT, DE) and select the five defenders that were the closest to the five offensive players, corresponding to a weighted euclidean distance, putting much more emphasis on the y-axis.
+From these 10 players, we derive 20 features related to their (standardized) position.
 
-Additionally, we extract relevant information from the play-by-play data, such as quarter, down, yards to go, home and away score and the remaining game time in the current half (in seconds).
-See the Appendix for a more detailed description of the features.
+Additionally, we extract information from the play-by-play data, such as quarter, down, yards to go, home and away score and the remaining seconds in the current half.
+See the Appendix for a more detailed description and a discussion on the choice of features.
 
@@
 In general, we have a limited dataset available (only 3985 plays) and therefore need to manage model complexity by controlling the number of features.
-Given the small dataset, we focus on 32 previously described basic features used in all 3 models: 6 convex hull related features, 20 player features, and 6 play-by-play features.
-In the Appendix, we provide a discussion on the choice of features.
+Given the small dataset, we focus on the 32 previously described basic features.
 
-First of all, we need a suitable basic model class for predicting man or zone coverage and opt for the following two: First, we fit a `glmnet` (elastic net) model, which performs implicit feature selection and is able to handle multicollinearity.
-Second, we use an `xgboost` model, which is able to capture non-linear effects (and interactions) and also handles multicollinearity.
-However, it necessitates careful hyperparameter tuning and generally performs better on bigger data sets.
-For all of the models, we use 10-fold cross validation on a suitable hyperparameter grid.
+Regarding a suitable basic model class for predicting man or zone coverage, we opt for the following two: First, we fit a `glmnet` (elastic net) model, which performs implicit feature selection and can handle multicollinearity.
+Second, we use an `xgboost` model, which additionally captures non-linear effects (and interactions).
+For these models, we use 10-fold cross validation on a suitable hyperparameter grid.
 
@@
 
-At first, we fit the aforementioned models with the previously described basic pre-motion features as all of these features are derived at the time of line-set.
+First, we fit the aforementioned models with the previously described basic pre-motion features.
 These very basic models serve as baseline models that allows to measure the effect of pre-snap motion (features) in the following.
@@
 
-In a second step, we extend our basic pre-motion model with naive post-motion features.
+Second, we extend our basic pre-motion model with naive post-motion features.
 To keep the complexity manageable, we derive only six additional post-motion features: for each team (offense and defense), we infer the maximum $y$-distance, the maximum $x$-distance and the total distance traveled by both teams until the snap.
@@
 The previously described models primarily serve as baselines for evaluating the impact of incorporating pre-snap motion information into predicting defensive schemes.
-The main contribution of our project lies in the effective integration of this information using a hidden Markov model (HMM).
+The main contribution of our project lies in the effective integration of this information using an HMM.
 
-Specifically, we use results derived from an HMM fitted to the pre-snap motion tracking data as further features in the previously described models.
+Specifically, we use results derived from an HMM fitted to the pre-snap motion tracking data as additional features in the previously described models.
 To achieve this, we model the movements of the five defensive players during pre-snap motion by an HMM (see the Appendix for an in-depth description).
@@
 <center>
-  <img src="./animation/animation.gif" width="423" height="450" alt="Coverage Prediction according to a hidden Markov model.">
+  <img src="./animation/animation.gif" width="423" height="450" alt="Coverage Prediction according to the HMM.">
 </center>
@@
 
-To incorporate pre-snap motion information, we re-train the post-motion model to predict whether the defense employs a man- or zone-coverage scheme, now integrating results from the HMM analysis as additional features.
+Third, we re-train the models from above, now integrating HMM results as additional features.
 A caveat of the aforementioned decoded state probabilities is that they form a multi-dimensional time series, complicating their direct inclusion as features in the previously described models.
@@
 
-As the entropy is a measure of uncertainty or randomness in a probability distribution, higher entropy indicates greater unpredictability, while lower entropy signifies more predictability.
-In our setting, we suspect that higher entropy values are associated with less persistent guarding allocations, while lower entropy values are indicative of more stable guarding assignments.
+As the entropy is a measure of uncertainty in a probability distribution, lower (higher) entropy indicates greater (un)predictability.
+Here, we suspect that higher (lower) entropy values are associated with less (more stable) persistent guarding allocations.
 
@@
 
-`r colorize("Den Text finde ich irgendwie unrund und nicht so leicht verständlich.","red")`
-
-<!--
-We compare the predictive performance of our three models to determine the effectiveness of player movements before snap for detecting the correct defensive scheme.
-We train the models on 85% of the data and evaluate performance on the remaining 15%.
-Since we use cross-validation for tuning of the models during the training phase, we perform our evaluation repeatedly.
-We use 50 repetition each time splitting the data differently, and for each we tune and train a model and then evaluate the performance on the test set.
-We do this to ensure that our results are meaningful and not due to chance (especially in our small data set).
--->
-
-`r colorize("Vorschlag (Robert):","red")`
-
-To assess whether the incorporation of pre-snap player movements can improve the prediction of defensive schemes, we evaluate the predictive performance of our three models.
-Therefore, we perform a train-test split using 85% for training and 15% for testing.
+To assess whether the incorporation of pre-snap player movements can improve the prediction of defensive schemes, we evaluate the predictive performance of our three models using a train-test split of 85/15.
 To ensure robust and reliable results, we repeat the evaluation process 50 times, using different CV splits for each iteration.
@@
 
-<!-- We evaluate the predictive performance of our three models to assess how well pre-snap player movements help identify the correct defensive scheme.
-To do this, we split the data into two sets: 85% for training and 15% for testing.
-During the training phase, we use cross-validation to fine-tune the models.
--->
-
-<!-- To ensure robust and reliable results, we repeat the evaluation process 50 times, using different cv splits for each iteration.
-For each split, we train and fine-tune the model on the training data and then evaluate its performance on the test set.
-This approach helps minimize the risk of our findings being influenced by randomness, which is particularly important given the limited size of our dataset.
--->
-
-`r colorize("Die plots sind noch sehr verpixelt","red")`
-
-`r colorize("Robert: komisch, bei mir sind sie eig recht scharf... hab versucht es bisschen zu ändern. Wenns das nicht besser macht, weiß ich leider auch nich was hilft...","red")`
-
 <div style="text-align: center;">
-  ![](ACCs_bdb_main.png)
-  ![](AUCs_bdb_main.png)
+  ![](ACCs_bdb_main2.png){width=50%}
+  ![](AUCs_bdb_main2.png){width=50%}
 </div>
 
-<!-- ![](ACCs_bdb_main.png) -->
-<!-- ![](AUCs_bdb_main.png) -->
-
-The two figures above show the results from our 50 repetitions of the experiment.
 The first figure illustrates the classification accuracy, computed using a threshold of 50\%.
-Since this value is somewhat arbitrary and our dataset is slightly imbalanced ($\approx$ 75\% zone coverage plays and $\approx$ 25\% man coverage plays), accuracy is a suboptimal metric for evaluating the results.
-Hence, we also present the AUC values of the three models (see Appendix for a discussion on evaluation metrics) in the second figure.
+Since this value is somewhat arbitrary and our dataset is slightly imbalanced ($\approx$ 75\% zone and $\approx$ 25\% man coverage plays), accuracy is a suboptimal metric (see Appendix for a discussion on evaluation metrics).
+Hence, we also present the AUC values of the three models in the second figure.
 Several observations emerge from these metrics.
-On the one hand, the more flexible `xgboost` model outperforms `glmnet`, but it also exhibits greater variation, indicating that additional data could enhance the tuning of this model.
-On the other hand, the pre-motion model clearly performs worst. Once motion information is incorporated, we observe a substantial improvement in performance.
+First, the more flexible `xgboost` model outperforms `glmnet`, but also exhibits greater variation, indicating that additional data could enhance the tuning of this model.
+Second, the pre-motion model clearly performs worst. Once motion information are incorporated, we observe a substantial improvement in performance.
 While this is to be expected, importantly, we also see that the HMM post-motion model outperforms the model only using naive post-motion information.
@@
 
-`r colorize("Die figure fehlt irgendwie noch","red")`
+<!-- ![](differences_teams.png){width=50%} -->
 
-![](differences_teams.png){width=50%}
+Our approach is not limited to the mere prediction of coverages; it can also be applied to a variety of analyses.
+As an illustrative example, we focuse on examining how pre-snap motion has helped teams, according to our model, identify the correct defensive coverage.
+Specifically, we calculate pre- and HMM post-motion probabilities for each team and evaluated the extent to which these models accurately predicted the actual coverages.
 
-Main findings:
+As shown in the following table, teams that are well-known for extensive use of pre-snap motion --- such as the Niners and Dolphins --- are also among the ones that most frequently increased the likelihood of correctly decoding the defensive coverage.
+While some teams, such as the Giants or Titans, succeed in accurately identifying coverages despite employing limited motion, the general trend indicates that teams utilizing extensive motion are more effective at leveraging it to recognize defensive schemes.
 
-Plot: In general, motion improves the correct detection of coverage (median of most teams positive).
-Table: Giants (Titans) --> low number of motions but very effective 49 out of 60 (82 %) of the motion plays increased coverage detections.
-Miami, Atlanta, SF --> Extensive use of motion but still very effective.
-Raider, Seattle, Rams --> Substantial use of motion but very ineffective.
-Bills (Bengals) -->  little use of motion and also ineffective.
-`r colorize("Rouven, wenn Du es schaffst einen netten text zu den Findings zu verfassen wäre es super!","red")`
+<!-- Plot: In general, motion improves the correct detection of coverage (median of most teams positive). Table: Giants (Titans) --> 
+<!-- low number of motions but very effective 49 out of 60 (82 %) of the motion plays increased coverage detections. Miami, Atlanta, SF --> 
+<!-- Extensive use of motion but still very effective. Raider, Seattle, Rams --> 
+<!-- Substantial use of motion but very ineffective. Bills (Bengals) -->  
+<!-- little use of motion and also ineffective. -->
+<!-- `r colorize("Rouven, wenn Du es schaffst einen netten text zu den Findings zu verfassen wäre es super!","red")` -->
 
@@
 
-This project explored the use of hidden Markov models (HMMs) to enhance the prediction of defensive schemes in football by incorporating pre-snap player movements.
+This project explored the use of HMMs to enhance the prediction of defensive schemes by incorporating pre-snap player movements.
 Models augmented with HMM-derived features demonstrated improved predictive performance compared to those relying solely on naive motion data.
-
-Comparing predictions from pre-motion and post-motion models provides insights into an offense's ability to use motion effectively to identify defensive coverage schemes.
-Similarly, future analysis could also relate to evaluating a defense's skill in disguising their coverage and confusing the offense during pre-snap motion.
 

### 2025-01-06 submission_61.Rmd
final updates

@@
 While we similarly predict man- or zone coverage before motion, we further leverage the additional information of pre-snap player movements.
-Specifically, in addition to including rather naive post-motion features, we use an HMM to model defenders' trajectories based on hidden states, which represent the offensive players they may be guarding.
+Specifically, in addition to including rather naive post-motion features, we use a hidden Markov model (HMM) to model defenders' trajectories based on hidden states, which represent the offensive players they may be guarding.
 Incorporating summary statistics of the state decoding results as features into the existing models substantially improves the predictive ability.
@@
 
-We aim to forecast the man- or zone coverage using the `pff_passCoverage` indicator in play-by play data.
+We aim to forecast the man- or zone coverage using the `pff_passCoverage` indicator in the play-by play data.
 We omit plays tagged as *others* and plays with more than five offensive linemen and with two quarterbacks.
@@
 
-To accurately forecast the defensive scheme, we create various features derived from the tracking data.
-In particular, we conducted the following feature engineering steps: First, using all 11 players on each side, we compute the area spanned by the convex hull of a team and the largest $y$ distance (i.e., the width of the hull) and the largest $x$ distance (i.e., the length of the hull), i.e.
+To accurately forecast the defensive scheme, we created various features derived from the tracking data.
+In particular, we conducted the following feature engineering steps: First, using all 11 players on each side, we computed the area spanned by the convex hull of a team and the largest $y$ distance (i.e., the width of the hull) and the largest $x$ distance (i.e., the length of the hull), i.e.
 6 features.
-Then, we select the five most relevant players on each side of the field.
-For offense, we omit the offensive line and the QB and for defense, we disregard defense lineman (NT, DT, DE) and select the five defenders that were the closest to the five offensive players, corresponding to a weighted euclidean distance, putting much more emphasis on the y-axis.
-From these 10 players, we derive 20 features related to their (standardized) position.
+Then, we selected the five most relevant players on each side of the field.
+For offense, we omitted the offensive line and the QB and for defense, we disregarded defense lineman (NT, DT, DE) and selected the five defenders that were the closest to the five offensive players, corresponding to a weighted euclidean distance, putting much more emphasis on the y-axis.
+From these 10 players, we derived 20 features related to their (standardized) position.
 
-Additionally, we extract information from the play-by-play data, such as quarter, down, yards to go, home and away score and the remaining seconds in the current half.
+Additionally, we extracted information from the play-by-play data, such as quarter, down, yards to go, home and away score and the remaining seconds in the current half.
 See the Appendix for a more detailed description and a discussion on the choice of features.
@@
 We train different models to predict whether the defense plays a man- or zone coverage scheme.
-Since the aim of the project is to show the effectiveness of pre-snap motion, we follow a three-step approach:
+Since the aim of the project is to demonstrate the effectiveness of pre-snap motion for prediction, we consider the following three models:
 
@@
 First, we fit the aforementioned models with the previously described basic pre-motion features.
-These very basic models serve as baseline models that allows to measure the effect of pre-snap motion (features) in the following.
+These very basic models serve as a baseline that allows to measure the effect of pre-snap motion (features) in the following.
 
@@
 
-After state decoding using the fitted HMM, we can visualize the inferred guarding assignment as depicted below.
+After state decoding using the fitted HMM, we can visualize the inferred guarding assignment as shown below.
 It becomes evident that the HMM effectively captures Marco Wilson's guarding assignment.
@@
 ## References
-*Franks A, Miller A, Bornn L, Goldsberry K (2015). Characterizing the Spatial Structure of Defensive Skill in Professional Basketball. The Annals of Applied Statistics, 9(1), DOI:10.1214/14-AOAS799
 
-*Koslik J (2024). LaMa: Fast Numerical Maximum Likelihood Estimation for Latent Markov Models. R package version 2.0.2, <https://CRAN.R-project.org/package=LaMa>.
+* Franks A, Miller A, Bornn L, Goldsberry K (2015). Characterizing the Spatial Structure of Defensive Skill in Professional Basketball. The Annals of Applied Statistics, 9(1), DOI:10.1214/14-AOAS799
 
-*Zucchini W, MacDonald I, Langrock R (2016). Hidden Markov Models for Time Series - An Introduction Using R. CRC Press
+* Koslik J (2024). LaMa: Fast Numerical Maximum Likelihood Estimation for Latent Markov Models. R package version 2.0.2, <https://CRAN.R-project.org/package=LaMa>.
+
+* Zucchini W, MacDonald I, Langrock R (2016). Hidden Markov Models for Time Series - An Introduction Using R. CRC Press
 

### 2025-01-07 README.Rmd
updating README

---
output: github_document
---

```{r setup, echo = FALSE, warning=FALSE, message=FALSE, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE
)
```

# HMMotion: Using tracking data to predict coverage strategies in American Football <img src="plots/bdb_logo.png" align="right" height=170>

This is the github repository for the corresponding submission to the NFL Big Data Bowl 2024 competition on Kaggle. The full notebook can be viewed [here](https://www.kaggle.com/code/rouvenmichels/hmmotion-using-tracking-data-to-predict-coverage).

## Quick Summary

<center>
<img src = "./animation/animation.gif" width = 500>
</center>

## Code Information

### 2025-01-07 README.Rmd
updating README

@@
 
-This is the github repository for the corresponding submission to the NFL Big Data Bowl 2024 competition on Kaggle.
+This is the github repository for the corresponding submission to the NFL Big Data Bowl 2025 competition on Kaggle.
 The full notebook can be viewed [here](https://www.kaggle.com/code/rouvenmichels/hmmotion-using-tracking-data-to-predict-coverage).
@@
 ## Quick Summary
+
+In this project, we focused on incorporating pre-snap motion information into the prediction of the coverage scheme, i.e.
+man or zone-coverage.
+Specifically, we developed a hidden Markov model (HMM) to infer the latent guarding assignments based on the movement of the defenders in response to all offenders.
+We show that incorporating the informaton obtained from decoding the hidden guarding assignments can improve the predictive power of the coverage scheme.
 

### 2025-01-07 zvm_cv_runs.R
zvm models

@@ -1,0 +1,602 @@
+######################################################################################
+######################################################################################
+#### BDB 2025 modeling: zone vs man cross validation tuning study
+######################################################################################
+######################################################################################
+
+# Libraries ---------------------------------------------------------------
+
+library(tidyverse)
+library(pROC)
+library(glmnet)
+library(glmnetUtils)
+library(ranger)
+library(xgboost)
+
+# ZVM modeling ---------------------------------------------------------------
+
+
+### relevant link for xgboost: https://stats.stackexchange.com/questions/342552/what-is-the-binarylogistic-objective-function-in-xgboost
+
+zvm_models <- function(model_data,
+                       test_data,
+                       ml_type = c("glmnet","rf","xgboost"),
+                       full_cv = TRUE,
+                       excl_vars = NULL,
+                       param_grid = NULL,
+                       seed = 123,
+                       return_preds = FALSE,
+                       ...){
+  dta <- model_data |>
+    select(-gameId,-playId,
+           -gameplayId,-pff_passCoverage,-c({{excl_vars}}))
+  
+  dta_test <- test_data |>
+    select(-gameId,-playId,
+           -gameplayId,-pff_passCoverage,-c({{excl_vars}}))
+  
+  ml_type = match.arg(ml_type)
+  
+  if(ml_type == "glmnet"){
+    set.seed(seed)
+    if(full_cv){
+      mods <- cva.glmnet(x = as.matrix(dta |> select(-pff_manZone_num)),y = dta$pff_manZone_num,family = binomial(),...)
+      mins <- sapply(mods$modlist,\(x) min(x$cvm))
+      modu <- mods$modlist[[which.min(mins)]]
+    }else{
+      modu <- cv.glmnet(x = as.matrix(dta |> select(-pff_manZone_num)),y = dta$pff_manZone_num,family = binomial(),...)
+    }
+    coefs <- coef(modu)
+    newxu <- as.matrix(dta_test |> select(-pff_manZone_num))
+    preds <- predict(modu,newx = newxu, s = "lambda.min",type = "response")
+    out <- preds > 0.5 
+    acc <- mean(out == dta_test$pff_manZone_num)
+    
[…]

### 2025-01-07 zvm_team_analyses.R

@@ -1,0 +1,496 @@
+######################################################################################
+######################################################################################
+#### BDB 2025 preprocessing: team analyses
+######################################################################################
+######################################################################################
+
+# Libraries ---------------------------------------------------------------
+
+library(tidyverse)
+
+
+# Load data ---------------------------------------------------------------
+
+games = read.csv("data/games.csv")
+plays = read.csv("data/plays.csv") |>
+  mutate(gameplayId = as.numeric(paste0(gameId, playId)))
+full_data = readRDS("data/zvm_motion_mod_data.rds")
+
+fm_add <- full_data |>
+  left_join(games |> select(gameId,week),by = "gameId") |>
+  left_join(plays |> select(gameplayId,possessionTeam))
+
+quarter_seconds <- function(timestr){
+  st <- str_split(timestr,":")
+  sapply(st,function(s){
+    as.numeric(s[1])*60+as.numeric(s[2])
+  })
+}
+
+model_data_full <- fm_add |>
+  filter(pff_manZone != "Other") |>
+  mutate(half_seconds_remaining = if_else(quarter %in% c(1,3),
+                                          quarter_seconds(gameClock)+15*60,
+                                          quarter_seconds(gameClock))) |>
+  select(-gameClock,-starts_with("nflId")) |>
+  mutate(pff_manZone_num = if_else(pff_manZone == "Man",1,0)) |>
+  select(-pff_manZone) 
+
+
+# Team models ---------------------------------------------------------------
+
+##  CIN model -----------------------------------------
+
+CIN_plays <- model_data_full |> 
+  filter(possessionTeam == "CIN") |>
+  select(-week,-possessionTeam)
+
+prop.table(table(CIN_plays$pff_manZone_num))
+
+NCIN_plays <- model_data_full |> 
+  filter(possessionTeam != "CIN") |>
+  select(-week,-possessionTeam)
+
+prop.table(table(NCIN_plays$pff_manZone_num))
+
+seedu <- 42
+
+CIN_xy_npo <- zvm_models(NCIN_plays,CIN_plays,
+                         ml_type = "xgboost",
+                         excl_vars = c("average","sum","nr_player_changes","average_ent",
[…]
